# Review of All Data Sets

We have now processed all of the data sets for this study following the methods outlined in the [previous section](#primer). Here, we'll review the `cloud2trees` output data generated by each data set where each data set represents a unique UAS flight plan. Remember, the objective of the study is to compare the tree detection and forest structure outcomes across different UAS-lidar flight plan settings.

## Tracking

get a list of the data available and parse the folder names to create a data set with all relevant variables for tracking the different flight setting parameters

```{r}
tracking_df <-
  list.dirs("../data/processed", recursive = F) %>% 
  dplyr::tibble() %>%
  dplyr::rename(path = 1) %>% 
  dplyr::mutate(
    folder = basename(path) %>% tolower()
    # , path = normalizePath(path)
  ) %>%
  dplyr::filter(stringr::str_starts(folder,"processed_")) %>% 
  tidyr::extract(
    col = folder
    , into = c("agl_ft", "mph")
    , regex = "([0-9]+)agl_([0-9]+)mph"
    , remove = F
    , convert = T
  ) %>% 
  dplyr::arrange(agl_ft, mph) %>% 
  dplyr::mutate(
    # convert to metric
    altitude_m = round(agl_ft/3.281)
    , meters_per_second = round(mph/2.237,1)
    # factor
    , dataset_factor = 
      stringr::str_glue("Altitude (m): {altitude_m} | Speed (m/s): {meters_per_second}") %>%
      forcats::as_factor() %>%
      forcats::fct_inorder(ordered = T)
    # , tf = tf=="on"
  ) %>% 
  dplyr::select(-c(agl_ft,mph)) %>% 
  dplyr::rename_with(.cols = -c(path,folder,dataset_factor), .fn = ~paste0("flight_",.x,recycle0 = T) )
# huh?
dplyr::glimpse(tracking_df)
# View(tracking_df)
```

## Raster Data and Study Boundary

now, we're going to load in the CHM data for each flight

```{r}
chm_rast <-
  tracking_df$path %>% 
  purrr::map(
    \(x)
    terra::rast(
      file.path(x, "point_cloud_processing_delivery", "chm_0.25m.tif")
    )
  )
names(chm_rast) <- tracking_df$dataset_factor
# huh?
chm_rast[1:min(length(chm_rast),2)]
```

and we'll load in the stand boundary to get an idea of how our UAS data relates

```{r}
stand_boundary <-
  sf::st_read(
    "../data/N1_Boundary/N1_Boundary.shp"
    # "c:/data/usfs/uas_sfm_tree_detection/data/field_validation/field_data_Boundary.shp"
    , quiet = T
  ) %>% 
  dplyr::rename_with(tolower) %>% 
  dplyr::filter(
    name=="N1"
    # site=="N1"
  ) %>% 
  sf::st_transform(terra::crs(chm_rast[[1]])) %>% 
  dplyr::mutate(stand_area_m2 = sf::st_area(.) %>% as.numeric())
# what
stand_boundary %>% dplyr::glimpse()
```

plot CHMs

```{r, fig.height=10.5, fig.width=10, results=F, fig.show='asis'}
graphics::par(mfrow = c(ceiling(length(chm_rast)/3),3)) # mfrow = c(nr, nc)
# use purrr::walk() because we only want it to draw the plot
purrr::iwalk(
  chm_rast
  , function(x,nm){
    terra::plot(
      x
      # , mar = c(b,l,t,r)
      , mar = c(0.5,0.5,2,1.3)
      , col = viridis::plasma(n=100)
      , axes = F
      , main = nm
      , plg = list(
        # title = "CHM (m)"
        cex = 0.8, title.cex = 0.9, shrink = 0.6
      )
    )
    # add stand
    terra::plot(terra::vect(stand_boundary), lwd = 3, col = NA, border = "blue", axes = F, add = T)
})
# reset graphics::par
graphics::par(mfrow = c(1, 1))
```

what about the DTMs?

```{r}
dtm_rast <- 
  tracking_df$path %>% 
  purrr::map(
    \(x)
    terra::rast(
      file.path(x, "point_cloud_processing_delivery", "dtm_0.5m.tif")
    )
  )
names(dtm_rast) <- tracking_df$dataset_factor
# huh?
dtm_rast[1:min(length(dtm_rast),2)]
```

plot DTMs

```{r, fig.height=10.5, fig.width=10, results=F, fig.show='asis'}
graphics::par(mfrow = c(ceiling(length(dtm_rast)/3),3)) # mfrow = c(nr, nc)
# use purrr::walk() because we only want it to draw the plot
purrr::iwalk(
  dtm_rast
  , function(x,nm){
    terra::plot(
      x
      # , mar = c(b,l,t,r)
      , mar = c(0.5,0.5,2,1.3)
      , col = harrypotter::hp(n=100, option = "mischief", direction = -1)
      , axes = F
      , main = nm
      , plg = list(
        # title = "DTM (m)"
        cex = 0.8, title.cex = 0.9, shrink = 0.6
      )
    )
    # add stand
    terra::plot(terra::vect(stand_boundary), lwd = 3, col = NA, border = "blue", axes = F, add = T)
})
# reset graphics::par
graphics::par(mfrow = c(1, 1))
```

let's see what data we have for evaluation

```{r}
# quality
tracking_df %>% 
  dplyr::mutate(
    dplyr::across(tidyselect::starts_with("flight_"), as.factor)
    , flight_meters_per_second = forcats::fct_rev(flight_meters_per_second)
    , flight_tf = "✓"
  ) %>% 
  dplyr::arrange(flight_tf,desc(flight_meters_per_second),desc(flight_altitude_m)) %>% 
  ggplot2::ggplot(
    mapping = ggplot2::aes(x = flight_altitude_m, y = flight_meters_per_second, label = flight_tf, color = flight_tf, fontface = "bold")
  ) +
    ggplot2::geom_tile(fill = NA, color = "black") +
    ggplot2::geom_text(size = 20, color = "forestgreen") +
    ggplot2::labs() +
    ggplot2::scale_x_discrete(position = "top") +
    ggplot2::coord_cartesian(expand = F) +
    ggplot2::scale_color_manual(values = c("gray55","gray11"), guide = "none") +
    ggplot2::theme_light() +
    ggplot2::theme(
      panel.grid = ggplot2::element_blank()
      , axis.text = ggplot2::element_text(size = 11, color = "black")
      , axis.title = ggplot2::element_text(size = 12, face = "bold", color = "black")
      , panel.border = ggplot2::element_rect(color = "black")
    )
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# cloud2trees:::search_dir_final_detected(c2t_process_dir)
remove(dtm_rast)
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Field Trees

let's load the field data of stem mapped trees with individual tree measurements for use in validation

```{r}
# filter for height to match UAS filtering
my_tree_ht_m <- 2
# read and filter
validation_trees <-
  # sf::st_read("c:/data/usfs/uas_sfm_tree_detection/data/field_validation/N1.gpkg", quiet = T) %>%
  readxl::read_excel("../data/N1_2024_Field_Trees.xlsx") %>% 
  dplyr::rename_with(
    ~ tolower(.x) %>% 
      stringr::str_squish() %>% 
      make.names() %>% 
      stringr::str_replace_all( "\\.{2,}", ".") %>% 
      stringr::str_remove( "\\.$") %>% 
      stringr::str_replace_all( "\\.", "_")
  ) %>% 
  sf::st_as_sf(
    coords = c("easting", "northing")
    , crs = 26913
    , remove=F
  ) %>% 
  dplyr::rename(
    # field_dbh_cm = dbh_cm
    # , field_tree_height_m = ht_m
    field_dbh_in = dbh_24
    , field_tree_height_ft = ht_24
  ) %>% 
  # dplyr::select(field_dbh_in, field_tree_height_ft) %>% summary()
  dplyr::mutate(
    field_dbh_cm = field_dbh_in/0.394
    , field_tree_height_m = field_tree_height_ft/3.281
  ) %>% 
  # dplyr::select(field_dbh_cm, field_tree_height_m) %>% summary()
  sf::st_set_geometry("geometry") %>% 
  dplyr::filter(
    !is.na(field_dbh_cm)
    & !is.na(field_tree_height_m)
    & sf::st_is_valid(geometry)
    # only keep trees that are above height threshold used for uas processing
    & field_tree_height_m >= my_tree_ht_m
    # & field_dbh_cm >= min_tree_dbh_cm # if know min field dbh for field sampling
  ) %>% 
  sf::st_transform(terra::crs(chm_rast[[1]])) %>% 
  sf::st_intersection(
    stand_boundary %>%
      dplyr::select(stand_area_m2)
  ) %>%
  dplyr::mutate(
    field_tree_id = dplyr::row_number()
    , tree_utm_x = sf::st_coordinates(geometry)[,1] #lon
    , tree_utm_y = sf::st_coordinates(geometry)[,2] #lat
    # basal area
    , basal_area_m2 = pi * field_dbh_cm^2 / (4*10000)
  ) %>% 
  dplyr::relocate(field_tree_id)
# huh?
validation_trees %>% dplyr::glimpse()
```

look at a map of the validation trees over the study boundary

```{r, message = F}
mapview::mapview(
  stand_boundary
  , color = "blue"
  , lwd = 1.5
  , alpha.regions = 0
  , label = F
  , legend = F
  , popup = F
  , layer.name = "study boundary"
  , leaflet.options = list(maxZoom = 16)
) + 
mapview::mapview(
  validation_trees
  , zcol = "field_tree_height_m"
  , col.regions = viridis::plasma(n=100)
  , cex = 3
  , layer.name = "field tree ht. (m)"
)
```

let's check out very high-level stand summary metrics

```{r}
validation_trees %>% 
  sf::st_drop_geometry() %>% 
  dplyr::group_by(stand_area_m2) %>% 
  dplyr::summarise(
    dplyr::across(
      tidyselect::starts_with("field_")
      , .fns = list(mean = mean, sd = sd)
    )
    , n_trees = dplyr::n()
    , basal_area_m2 = sum(basal_area_m2)
  ) %>% 
  # add area
  dplyr::mutate(stand_area_ha = stand_area_m2/10000) %>% 
  dplyr::mutate(
    ht = paste0(
      field_tree_height_m_mean %>% 
        round(1) %>% 
        scales::comma(accuracy = 0.1)
      , "<br/>("
      , field_tree_height_m_sd %>% 
          round(1) %>% 
          scales::comma(accuracy = 0.1)
      , ")"
    )
    , dbh = paste0(
      field_dbh_cm_mean %>% round(1) %>% scales::comma(accuracy = 0.1)
      , "<br/>("
      , field_dbh_cm_sd %>% round(1) %>% scales::comma(accuracy = 0.1)
      , ")"
    )
    , trees_ha = n_trees/stand_area_ha
    , basal_area_m2_per_ha = basal_area_m2/stand_area_ha
    , stand_area_ha = stand_area_ha %>% round(1) %>% scales::comma(accuracy = 0.1)
    , n_trees = scales::comma(n_trees,accuracy=1)
  ) %>% 
  dplyr::select(
    stand_area_ha, n_trees, trees_ha, basal_area_m2_per_ha, ht, dbh
  ) %>% 
  kableExtra::kbl(
    caption = "Stand Summary of Field Data<br/>mean and SD of height and DBH are reported"
    , escape = F
    , digits = 1
    , col.names = c(
      "Hectares", "# trees"
      , "Trees<br/>ha<sup>-1</sup>"
      , "Basal Area<br/>m<sup>2</sup> ha<sup>-1</sup>"
      , "Height (m)", "DBH (cm)"
    )
  ) %>% 
  kableExtra::kable_styling()
```

let's quickly look at the summary stats on tree size

```{r, echo=FALSE, message=FALSE}
validation_trees %>% 
  sf::st_drop_geometry() %>% 
  dplyr::select(field_tree_height_m,field_dbh_cm) %>% 
  dplyr::mutate(dplyr::across(dplyr::everything(), ~round(.x,1))) %>% 
  summary() %>% 
  knitr::kable(format = "markdown", digits = 2)
```

the size distribution too

```{r}
validation_trees %>% 
  sf::st_drop_geometry() %>% 
  dplyr::select(field_tree_height_m,field_dbh_cm) %>% 
  tidyr::pivot_longer(
    cols = dplyr::everything()
    , names_to = "metric"
    , values_to = "value"
    , values_drop_na = T
  ) %>% 
  dplyr::mutate(
    tree_metric = metric %>% 
      stringr::str_remove("(_rmse|_rrmse|_mean|_mape)$") %>% 
      stringr::str_extract("(height|dbh)") %>% 
      factor(
        ordered = T
        , levels = c(
          "height"
          , "dbh"
        )
        , labels = c(
          "Height (m)"
          , "DBH (cm)"
        )
      )
  ) %>% 
  # dplyr::glimpse()
# plot dist
  ggplot2::ggplot(mapping = ggplot2::aes(x = value)) +
  ggplot2::geom_density(
    mapping = ggplot2::aes(y=ggplot2::after_stat(scaled))
    , alpha = 0.7
    , fill = "maroon4", color = "maroon4"
  ) +
  ggplot2::facet_grid(
    cols = dplyr::vars(tree_metric)
    , scales = "free_x", axes = "all_x"
  ) +
  ggplot2::scale_y_continuous(NULL,breaks=NULL) +
  ggplot2::scale_x_continuous(breaks=scales::breaks_extended(n=8)) +
  ggplot2::labs(
    x=""
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "top"
    , strip.text.x = ggplot2::element_text(color = "black", face = "bold")
    , panel.grid = ggplot2::element_blank()
  )
```

height and DBH relationship

```{r}
validation_trees %>% 
  sf::st_drop_geometry() %>% 
  dplyr::slice_sample(prop = 0.77) %>% 
  ggplot2::ggplot(mapping = ggplot2::aes(x = field_tree_height_m, y = field_dbh_cm)) + 
  ggplot2::geom_point(color = "maroon4", alpha = 0.6) +
  ggplot2::labs(x = "field tree ht. (m)", y = "field tree DBH (cm)") +
  ggplot2::scale_x_continuous(limits = c(0,NA), breaks = scales::breaks_extended(n=8)) +
  ggplot2::scale_y_continuous(limits = c(0,NA), breaks = scales::breaks_extended(n=8)) +
  ggplot2::theme_light()
```

a distribution of trees by size class (bin) is another thing to look at

```{r}
# make ggplot2::cut_* bins pretty
make_bins_pretty <- function(bin_factor, suffix = "") {
  # convert factor to character and remove brackets and parens
  clean_labels <- bin_factor %>%
    as.character() %>%
    stringr::str_replace_all(c("\\[" = "", "\\(" = "", "\\]" = "", "\\)" = "")) %>%
    stringr::str_replace(",", "-")
  
  # suffix
  if(stringr::str_squish(suffix) != "" && inherits(suffix,"character")){
    clean_labels <- stringr::str_c(clean_labels, " ", suffix) %>% stringr::str_squish()
  }
  
  # reorder based on the original factor levels 
  forcats::fct_reorder(factor(clean_labels), as.numeric(bin_factor))
}

validation_trees %>% 
  sf::st_drop_geometry() %>% 
  # metric cuts/bins
  dplyr::mutate(
    tree_height_m_bin = ggplot2::cut_width(field_tree_height_m, width = 5, boundary = 0, closed = "left") %>% 
      make_bins_pretty(suffix = "m") %>% 
      forcats::fct_relabel(~ stringr::str_replace(.x, "^0-", paste0(my_tree_ht_m,"-")) )
  ) %>% 
  dplyr::count(tree_height_m_bin) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(pct = n/sum(n,na.rm = T)) %>% 
  dplyr::mutate(
    lab = ifelse(
      pct>0.049
      , paste0(
        scales::percent(pct,accuracy=1) 
        ,"\n("
        , scales::comma(n,accuracy=1) 
        , ")"
      )
      , ""
    )
  ) %>% 
  dplyr::ungroup() %>% 
  # dplyr::glimpse()
  ggplot2::ggplot(
    mapping = ggplot2::aes(y = pct, x = tree_height_m_bin)
  ) +
  ggplot2::geom_col(
    width = 0.6
    , color = NA, alpha = 0.9, fill = "maroon4"
  ) +
  ggplot2::geom_text(
    mapping = ggplot2::aes(label = lab, fontface = "bold")
    # , color = "black"
     , color = "black", size = 2.5, vjust = -0.2
  ) +
  ggplot2::scale_x_discrete(drop = F) +
  ggplot2::scale_y_continuous(
    breaks = seq(0,1,by=0.2)
    , labels = scales::percent
    , expand = ggplot2::expansion(mult = c(0,0.1))
  ) +
  ggplot2::labs(
    x = "tree height", y = ""
    , subtitle = "field-measured trees by tree height bin"
  ) +
  ggplot2::theme_light()+
  ggplot2::theme(
    legend.position = "top"
    , strip.text = ggplot2::element_text(size = 11, color = "black", face = "bold")
    , axis.text.x = ggplot2::element_text(size = 7, angle = 45, hjust = 1, vjust = 1)
    , axis.text.y = ggplot2::element_blank()
    , axis.ticks.y = ggplot2::element_blank()
  )
```

and DBH bins

```{r}
validation_trees %>% 
  sf::st_drop_geometry() %>% 
  # metric cuts/bins
  dplyr::mutate(
    tree_dbh_cm_bin = ggplot2::cut_width(field_dbh_cm, width = 12, boundary = 0, closed = "left") %>% 
      make_bins_pretty(suffix = " cm") %>% 
      forcats::fct_relabel(~ stringr::str_replace(.x, "^0-", paste0(floor(min(validation_trees$field_dbh_cm)),"-")) )
  ) %>% 
  dplyr::count(tree_dbh_cm_bin) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(pct = n/sum(n,na.rm = T)) %>% 
  dplyr::mutate(
    lab = ifelse(
      pct>0.049
      , paste0(
        scales::percent(pct,accuracy=1) 
        ,"\n("
        , scales::comma(n,accuracy=1) 
        , ")"
      )
      , ""
    )
  ) %>% 
  dplyr::ungroup() %>% 
  # dplyr::glimpse()
  ggplot2::ggplot(
    mapping = ggplot2::aes(y = pct, x = tree_dbh_cm_bin)
  ) +
  ggplot2::geom_col(
    width = 0.6
    , color = NA, alpha = 0.9, fill = "maroon4"
  ) +
  ggplot2::geom_text(
    mapping = ggplot2::aes(label = lab, fontface = "bold")
    # , color = "black"
     , color = "black", size = 2.5, vjust = -0.2
  ) +
  ggplot2::scale_x_discrete(drop = F) +
  ggplot2::scale_y_continuous(
    breaks = seq(0,1,by=0.2)
    , labels = scales::percent
    , expand = ggplot2::expansion(mult = c(0,0.1))
  ) +
  ggplot2::labs(
    x = "tree DBH", y = ""
    , subtitle = "field-measured trees by tree DBH bin"
  ) +
  ggplot2::theme_light()+
  ggplot2::theme(
    legend.position = "top"
    , strip.text = ggplot2::element_text(size = 11, color = "black", face = "bold")
    , axis.text.x = ggplot2::element_text(size = 7, angle = 45, hjust = 1, vjust = 1)
    , axis.text.y = ggplot2::element_blank()
    , axis.ticks.y = ggplot2::element_blank()
  )
```

let's get a lot more detail on the tree measurements for DBH and height

```{r}
validation_trees %>% 
  sf::st_drop_geometry() %>% 
  dplyr::rename(
    tree_height_m = field_tree_height_m
    , dbh_cm = field_dbh_cm
  ) %>% 
  dplyr::select(tree_height_m, dbh_cm) %>% 
  dplyr::summarise(
    dplyr::across(
      dplyr::where(is.numeric)
      , .fns = list(
        mean = ~mean(.x,na.rm=T)
        , sd = ~sd(.x,na.rm=T)
        # , q10 = ~quantile(.x,na.rm=T,probs=0.1)
        , median = ~quantile(.x,na.rm=T,probs=0.5)
        # , q90 = ~quantile(.x,na.rm=T,probs=0.9)
        # , min = ~min(.x,na.rm=T)
        # , max = ~max(.x,na.rm=T)
        , range = ~paste0(
          scales::comma(min(.x,na.rm=T), accuracy = 0.1)
          ,"-"
          , scales::comma(max(.x,na.rm=T), accuracy = 0.1)
        )
      )
    )
    , n = dplyr::n()
  ) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(
    n = scales::comma(n,accuracy=1)
    , dplyr::across(
      dplyr::where(is.numeric)
      , ~scales::comma(.x,accuracy=0.1)
    )
    , `data` = "field-measured trees"
  ) %>% 
  dplyr::relocate(data,n) %>% 
  # ncol()
  kableExtra::kbl(
    caption = "height and diameter distributions <br />field-measured trees"
    , col.names = c(
      "data","# trees"
      , rep(c("Mean","Std Dev","Median","Range"), times = 2)
    )
    , escape = F
  ) %>% 
  kableExtra::kable_styling(font_size = 12) %>% 
  kableExtra::add_header_above(
    c(
      " "=2
      , "Height (m)" = 4
      , "DBH (cm)" = 4
    )
    , escape = F
  ) %>% 
  kableExtra::column_spec(seq(2,10,by=4), border_right = TRUE, include_thead = TRUE) %>% 
  kableExtra::collapse_rows(columns = 1, valign = "top")  
```

```{r, include=F, eval=F}
cloud2trees:::search_dir_final_detected(dir = file.path(tracking_df$path[1], "point_cloud_processing_delivery")) %>% 
  purrr::pluck("crowns_flist") %>% 
  purrr::map(
    \(x)
    sf::st_read(
      dsn = x
      , quiet = T
    )
  ) %>% 
  dplyr::bind_rows() %>% 
  dplyr::glimpse()

tracking_df$path[1]
ctg_temp <- 
  file.path(tracking_df$path[6], "point_cloud_processing_delivery", "norm_las") %>% 
  lidR::readLAScatalog()
ctg_temp
ctg_temp@data$geometry
mapview::mapview(
  stand_boundary
  , color = "blue"
  , lwd = 1.5
  , alpha.regions = 0
  , label = F
  , legend = F
  , popup = F
  , layer.name = "study boundary"
  , leaflet.options = list(maxZoom = 16)
) + 
mapview::mapview(ctg_temp@data$geometry, layer.name = "ptcld")

```

## UAS Predicted Trees

```{r, include=FALSE, eval=FALSE, results=F}
# chm_rast %>% 
tracking_df$folder %>% 
  # .[1] %>% 
  purrr::map(function(nm){
    of_temp <- tracking_df %>% 
      dplyr::filter(folder==nm) %>% 
      dplyr::pull(path) %>% 
      file.path("point_cloud_processing_delivery")
    # raster2trees
    ans_temp <- cloud2trees::raster2trees(
      chm_rast = chm_rast[[nm]]
      , outfolder = of_temp
      , ws = my_linear
      , min_height = 1.37
    )
    # ans_temp %>% dplyr::glimpse()
    # dbh
    dbh_ans_temp <- cloud2trees::trees_dbh(ans_temp, outfolder = of_temp)
    # dbh_ans_temp %>% dplyr::glimpse()
    nm_temp <- base::setdiff(names(dbh_ans_temp), names(ans_temp))
    # join
    ans_temp <- 
      ans_temp %>% 
      dplyr::left_join(
        dbh_ans_temp %>% 
          sf::st_drop_geometry() %>% 
          dplyr::select(dplyr::all_of(c(
            "treeID"
            , nm_temp
          )))
        , by = "treeID"
      )
    # write
    cloud2trees:::write_raster2trees_ans(ans_temp,dir = of_temp)
    # sf::st_read(file.path(of_temp,"final_detected_crowns.gpkg")) %>% dplyr::glimpse()
    # sf::st_read(file.path(of_temp,"final_detected_tree_tops.gpkg")) %>% dplyr::glimpse()

  })
```

let's load the UAS predicted trees

```{r, results=F}
predicted_trees <- 
  tracking_df$path %>% 
  # .[1:2] %>%
  purrr::map(function(i){
    dta <- 
      file.path(i, "point_cloud_processing_delivery", "final_detected_tree_tops.gpkg") %>% 
        sf::st_read(quiet=F) %>% 
        dplyr::filter(tree_height_m >= my_tree_ht_m) %>% 
        # clip to study boundary
        sf::st_intersection(
          stand_boundary %>% 
            dplyr::select(stand_area_m2) %>% 
            sf::st_buffer(3/2) # 0.5*max_dist_m where max_dist_m=allowable search distance for TP match
        ) %>% 
        dplyr::mutate(
          basal_area_m2 = pi * dbh_cm^2 / (4*10000)
          , las_overlap_stand_area_m2 = 
            file.path(i, "point_cloud_processing_delivery", "raw_las_ctg_info.gpkg") %>% 
              sf::st_read(quiet=T) %>% 
              sf::st_union() %>% 
              # sf::st_area() %>% as.numeric()
              # clip to study boundary
              sf::st_intersection(stand_boundary) %>% 
              sf::st_area() %>% as.numeric() %>% 
              sum()
        )
    # flag if in stand...do this to allow for trees outside of stand but might still match with reference due to lean, gps error, etc
    dta <- 
      dta %>% 
        dplyr::left_join(
          dta %>% 
            # clip to study boundary
            sf::st_intersection(
              stand_boundary %>% 
                dplyr::select(stand_area_m2) %>% 
                dplyr::mutate(is_in_stand = T)
            ) %>% 
            sf::st_drop_geometry() %>% 
            dplyr::select(treeID, is_in_stand)
          , by = "treeID"
        ) %>% 
        dplyr::mutate(
          is_in_stand = dplyr::coalesce(is_in_stand,F)
        )
    return (dta)
  })
names(predicted_trees) <- tracking_df$dataset_factor
```

let's check the number of predicted trees for each data set

```{r}
predicted_trees %>% 
  purrr::map_dfr(\(x) x %>% dplyr::filter(is_in_stand) %>% nrow()) %>% 
  tidyr::pivot_longer(dplyr::everything()) %>% 
  dplyr::rename(dataset_factor=name) %>% 
  dplyr::inner_join(tracking_df, by="dataset_factor") %>% 
  dplyr::select(tidyselect::starts_with("flight_"), value) %>% 
  dplyr::arrange(flight_altitude_m) %>% 
  dplyr::mutate(value=scales::comma(value)) %>% 
  kableExtra::kbl(
    caption = "Predicted trees by flight"
    , col.names = c(
      "altitude (m)", "speed (m sec<sup>-1</sup>)", "predicted trees"
    )
    , escape = F
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::collapse_rows(columns = c(1:2), valign = "top")
```

check the DBH and height distribution for each data set

```{r}
predicted_trees %>% 
  dplyr::bind_rows(.id = "dataset_factor") %>% 
  sf::st_drop_geometry() %>% 
  dplyr::filter(is_in_stand) %>% 
  dplyr::select(dataset_factor, tree_height_m, dbh_cm) %>% 
  tidyr::pivot_longer(cols = -c(dataset_factor)) %>% 
  dplyr::mutate(
    name = dplyr::recode_values(
      name
      , "tree_height_m" ~ "Height (m)"
      , "dbh_cm" ~ "DBH (cm)"
    )
  ) %>% 
  dplyr::inner_join(tracking_df, by="dataset_factor") %>% 
  dplyr::mutate(
    dplyr::across(tidyselect::starts_with("flight_"), as.factor)
    , dataset_factor = forcats::fct_relevel(dataset_factor, levels(tracking_df$dataset_factor))
  ) %>% 
  ggplot2::ggplot(mapping = ggplot2::aes(x=value,y=dataset_factor,fill=flight_altitude_m)) +
  ggplot2::geom_boxplot(width=0.6, outliers = F, color="gray55",lwd=0.7) +
  ggplot2::facet_grid(
    cols = dplyr::vars(name)
    , scales = "free_x"
    # , axes = "all"
  ) +
  ggplot2::scale_fill_viridis_d(option = "mako", begin = 0.5, end = 0.1) +
  ggplot2::scale_x_continuous(labels = scales::comma, breaks = scales::breaks_extended(7)) +
  ggplot2::theme_light() +
  ggplot2::labs(x = "", y = "", subtitle = "Predicted tree height and DBH by flight") +
  ggplot2::theme(
    legend.position = "none"
    , axis.text.y = ggplot2::element_text(size = 9, face = "bold")
    , axis.text.x = ggplot2::element_text(size = 8)
    , strip.text = ggplot2::element_text(size = 10, color = "black", face = "bold")
  )
```

both the count of trees and the height and DBH distributions look very similar across the different flight settings...this may be an early indication of limited influence of the flight settings on tree detection and forest structure quantification

let's get a lot more detail on the tree measurements for DBH and height

```{r}
predicted_trees %>% 
  dplyr::bind_rows(.id = "dataset_factor") %>% 
  sf::st_drop_geometry() %>% 
  dplyr::filter(is_in_stand) %>% 
  dplyr::select(dataset_factor, tree_height_m, dbh_cm) %>% 
  dplyr::inner_join(tracking_df, by="dataset_factor") %>% 
  dplyr::mutate(
    dplyr::across(tidyselect::starts_with("flight_"), as.factor)
    , dataset_factor = forcats::fct_relevel(dataset_factor, levels(tracking_df$dataset_factor))
  ) %>%
  dplyr::select(tidyselect::starts_with("flight_"), tree_height_m, dbh_cm) %>% 
  dplyr::group_by(dplyr::across(
    tidyselect::starts_with("flight_")
  )) %>% 
  dplyr::summarise(
    dplyr::across(
      c(tree_height_m, dbh_cm)
      , .fns = list(
        mean = ~mean(.x,na.rm=T)
        , sd = ~sd(.x,na.rm=T)
        # , q10 = ~quantile(.x,na.rm=T,probs=0.1)
        , median = ~quantile(.x,na.rm=T,probs=0.5)
        # , q90 = ~quantile(.x,na.rm=T,probs=0.9)
        # , min = ~min(.x,na.rm=T)
        # , max = ~max(.x,na.rm=T)
        , range = ~paste0(
          scales::comma(min(.x,na.rm=T), accuracy = 0.1)
          ,"-"
          , scales::comma(max(.x,na.rm=T), accuracy = 0.1)
        )
      )
    )
    , n = dplyr::n()
    , .groups = "drop_last"
  ) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(
    n = scales::comma(n,accuracy=1)
    , dplyr::across(
      dplyr::where(is.numeric)
      , ~scales::comma(.x,accuracy=0.1)
    )
  ) %>% 
  dplyr::relocate(tidyselect::starts_with("flight_"),n) %>% 
  # ncol()
  kableExtra::kbl(
    caption = "height and diameter distributions <br />UAS-detected trees"
    , col.names = c(
      "altitude (m)", "speed (m sec<sup>-1</sup>)","# trees"
      , rep(c("Mean","Std Dev","Median","Range"), times = 2)
    )
    , escape = F
  ) %>% 
  kableExtra::kable_styling(font_size = 11) %>% 
  kableExtra::add_header_above(
    c(
      " "=3
      , "Height (m)" = 4
      , "DBH (cm)" = 4
    )
    , escape = F
  ) %>% 
  kableExtra::column_spec(seq(3,11,by=4), border_right = TRUE, include_thead = TRUE) %>% 
  kableExtra::collapse_rows(columns = 1, valign = "top")  
```

let's check out very high-level stand summary metrics

```{r}
predicted_trees %>% 
  purrr::map_dfr(
    \(x)
    x %>% 
      sf::st_drop_geometry() %>% 
      dplyr::filter(is_in_stand) %>% 
      dplyr::group_by(las_overlap_stand_area_m2) %>% 
      dplyr::summarise(
        dplyr::across(
          c(tree_height_m, dbh_cm)
          , .fns = list(mean = mean, sd = sd)
        )
        , n_trees = dplyr::n()
        , basal_area_m2 = sum(basal_area_m2)
      ) %>% 
      # add area
      dplyr::mutate(stand_area_ha = las_overlap_stand_area_m2/10000) %>% 
      dplyr::mutate(
        ht = paste0(
          tree_height_m_mean %>% 
            round(1) %>% 
            scales::comma(accuracy = 0.1)
          , "<br/>("
          , tree_height_m_sd %>% 
              round(1) %>% 
              scales::comma(accuracy = 0.1)
          , ")"
        )
        , dbh = paste0(
          dbh_cm_mean %>% round(1) %>% scales::comma(accuracy = 0.1)
          , "<br/>("
          , dbh_cm_sd %>% round(1) %>% scales::comma(accuracy = 0.1)
          , ")"
        )
        , trees_ha = n_trees/stand_area_ha
        , basal_area_m2_per_ha = basal_area_m2/stand_area_ha
        , stand_area_ha = stand_area_ha %>% round(1) %>% scales::comma(accuracy = 0.1)
      ) %>% 
      dplyr::select(
        stand_area_ha, n_trees, trees_ha, basal_area_m2_per_ha, ht, dbh
      )
    , .id = "dataset_factor"
  ) %>% 
  # dplyr::glimpse()
  dplyr::inner_join(
    tracking_df %>% dplyr::select(dataset_factor, tidyselect::starts_with("flight_"))
    , by="dataset_factor"
  ) %>% 
  dplyr::select(-dataset_factor) %>% 
  dplyr::arrange(flight_altitude_m) %>% 
  dplyr::relocate(
    c(stand_area_ha, n_trees, trees_ha, basal_area_m2_per_ha, ht, dbh)
    , .after = dplyr::last_col()
  ) %>% 
  # dplyr::glimpse() %>% 
  kableExtra::kbl(
    caption = "Stand Summary of Predicted Data"
    , col.names = c(
      "altitude (m)", "speed (m sec<sup>-1</sup>)"
      , "Coverage<br/>Hectares", "# trees"
      , "Trees<br/>ha<sup>-1</sup>"
      , "Basal Area<br/>m<sup>2</sup> ha<sup>-1</sup>"
      , "Height (m)", "DBH (cm)"
    )
    , escape = F
    , digits = 1
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::collapse_rows(columns = c(1), valign = "top")
  
```


```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# cloud2trees:::search_dir_final_detected(c2t_process_dir)
remove(dtm_rast)
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Point Cloud Data and Processing Time

An important consideration when comparing enhanced forest inventories using UAS-lidar with traditional field-based surveys is the data processing time required to generate the tree list from the raw point cloud. Let's explore the point cloud data and processing time for each flight which `cloud2trees::cloud2trees()` automatically tracks for us and writes to the "processed_tracking_data.csv" file.

load in the tracking data

```{r}
timer_df <- 
  tracking_df$path %>% 
  purrr::set_names() %>% 
  # .[1:2] %>%
  purrr::map(function(i){
    i %>% 
      file.path("point_cloud_processing_delivery", "processed_tracking_data.csv") %>% 
      readr::read_csv(show_col_types = F, progress = F) %>% 
      dplyr::mutate(
        timer_total_time_mins = timer_cloud2raster_mins + timer_raster2trees_mins +
          timer_trees_dbh_mins + timer_trees_cbh_mins + timer_trees_type_mins +
          timer_trees_hmd_mins + timer_trees_biomass_mins
        , timer_total_dbhonly_time_mins = timer_cloud2raster_mins + timer_raster2trees_mins +
          timer_trees_dbh_mins
        , timer_tree_extraction_mins = timer_cloud2raster_mins + timer_raster2trees_mins
        , points_per_m2 = number_of_points/las_area_m2
        , timer_total_time_mins_per_ha = timer_total_time_mins/(las_area_m2/10000)
        , timer_total_dbhonly_time_mins_per_ha = timer_total_dbhonly_time_mins/(las_area_m2/10000)
      ) %>% 
      dplyr::mutate(
        dplyr::across(
         .cols = c(timer_tree_extraction_mins,
          timer_trees_dbh_mins, timer_trees_cbh_mins, timer_trees_type_mins,
          timer_trees_hmd_mins, timer_trees_biomass_mins)
         , .fns = ~ .x/timer_total_time_mins
         , .names = "{.col}_pct"
        )
        , dplyr::across(
         .cols = c(timer_tree_extraction_mins,
          timer_trees_dbh_mins, timer_trees_cbh_mins, timer_trees_type_mins,
          timer_trees_hmd_mins, timer_trees_biomass_mins)
         , .fns = ~ paste0(scales::comma( (.x/(las_area_m2/10000)),accuracy=0.01), "<br/>(", scales::percent( (.x/timer_total_time_mins), accuracy = 0.1 ), ")")
         , .names = "{.col}_full_lab"
        )
      )
  }) %>% 
  purrr::list_rbind(names_to = "path") %>% 
  dplyr::inner_join(
    tracking_df
    , by = "path"
  )
# timer_df %>% dplyr::glimpse()
```

table the processing stats

```{r}
timer_df %>% 
  dplyr::select(
    dataset_factor, tidyselect::starts_with("flight_")
    # data desc
    , number_of_points
    , points_per_m2
    # timer
    , timer_total_time_mins_per_ha
    , tidyselect::ends_with("_full_lab")
      # , timer_tree_extraction_mins,
      #   timer_trees_dbh_mins, timer_trees_cbh_mins, timer_trees_type_mins,
      #   timer_trees_hmd_mins, timer_trees_biomass_mins
  ) %>% 
  dplyr::mutate(
    number_of_points = scales::comma((number_of_points/1000000),accuracy=0.1,suffix = " M")
    , timer_total_time_mins_per_ha = scales::comma(timer_total_time_mins_per_ha,accuracy=0.1)
    , dplyr::across(tidyselect::ends_with("_per_m2"), ~scales::comma(.x,accuracy = 1))
  ) %>%
  dplyr::arrange(dataset_factor) %>% 
  dplyr::select(
    -c(
      dataset_factor
    )
  ) %>% 
  # ncol()
  # dplyr::glimpse()
  kableExtra::kbl(
    caption = "Data Processing Summary"
    , col.names = c(
      "altitude (m)", "speed (m sec<sup>-1</sup>)"
      # data desc
      , "# points", "points m<sup>-2</sup>"
      # timer
      , "TOTAL"
      , "tree<br/>extraction", "DBH", "CBH", "forest<br/>type", "HMD", "biomass"
    )
    , escape = F
    # , digits = 2
  ) %>% 
  kableExtra::kable_styling(font_size = 10) %>% 
  kableExtra::add_header_above(
    c(
      " "=2
      , "Input Data<br/>Description" = 2
      , "Processing Time<br/>minutes ha<sup>-1</sup>" = 7
    )
    , escape = F
  ) %>% 
  kableExtra::column_spec(c(2,4,11), border_right = TRUE, include_thead = TRUE) %>% 
  kableExtra::collapse_rows(columns = c(1), valign = "top")
  
```

for this forest inventory and our comparison to the field-measured data we only needed tree position, height, and DBH. let's make a condensed table with only those attributes

```{r}
timer_df %>% 
  dplyr::select(
    dataset_factor, tidyselect::starts_with("flight_")
    # data desc
    , number_of_points
    , points_per_m2
    # timer
    , timer_total_dbhonly_time_mins_per_ha
    , timer_total_dbhonly_time_mins
    , timer_tree_extraction_mins
    , timer_trees_dbh_mins
    , las_area_m2
  ) %>% 
  dplyr::mutate(
    dplyr::across(
     .cols = c(timer_tree_extraction_mins,timer_trees_dbh_mins)
     , .fns = ~ paste0(scales::comma( (.x/(las_area_m2/10000)),accuracy=0.01), "<br/>(", scales::percent( (.x/timer_total_dbhonly_time_mins), accuracy = 0.1 ), ")")
     , .names = "{.col}_full_lab"
    )
  ) %>% 
  dplyr::select(-c(
    timer_total_dbhonly_time_mins
    , timer_tree_extraction_mins
    , timer_trees_dbh_mins
    , las_area_m2
  )) %>% 
  dplyr::mutate(
    number_of_points = scales::comma((number_of_points/1000000),accuracy=0.1,suffix = " M")
    , timer_total_dbhonly_time_mins_per_ha = scales::comma(timer_total_dbhonly_time_mins_per_ha,accuracy=0.01)
    , dplyr::across(tidyselect::ends_with("_per_m2"), ~scales::comma(.x,accuracy = 1))
  ) %>%
  dplyr::arrange(dataset_factor) %>% 
  dplyr::select(
    -c(
      dataset_factor
    )
  ) %>% 
  # ncol()
  # dplyr::glimpse()
  kableExtra::kbl(
    caption = "Data Processing Summary (DBH only)"
    , col.names = c(
      "altitude (m)", "speed (m sec<sup>-1</sup>)"
      # data desc
      , "# points", "points m<sup>-2</sup>"
      # timer
      , "TOTAL"
      , "tree<br/>extraction", "DBH"
    )
    , escape = F
    # , digits = 2
  ) %>% 
  kableExtra::kable_styling(font_size = 11) %>% 
  kableExtra::add_header_above(
    c(
      " "=2
      , "Input Data<br/>Description" = 2
      , "Processing Time<br/>minutes ha<sup>-1</sup>" = 3
    )
    , escape = F
  ) %>% 
  kableExtra::column_spec(c(2,4,7), border_right = TRUE, include_thead = TRUE) %>% 
  kableExtra::collapse_rows(columns = c(1), valign = "top")

```


how is the point density related to speed and altitude?

```{r}
timer_df %>% 
  ggplot2::ggplot(
    mapping = aes(x = flight_altitude_m, y = points_per_m2, color = factor(flight_meters_per_second))
  ) +
  ggplot2::geom_point(size = 5.5, alpha = 0.9) +
  ggplot2::scale_color_manual(
    values = viridis::rocket(n = 3, begin = 0.9, end = 0.5)[c(1,3)]
  ) +
  ggplot2::scale_x_continuous(
    labels=scales::comma
    , breaks = scales::breaks_extended(n=8)
  ) +
  ggplot2::labs(
    color = latex2exp::TeX("Speed  $m \\cdot sec^{-1}$ ")
    , x = "Altitude (m)"
    , y = latex2exp::TeX("point cloud points  $m^{-2}$ ")
    , subtitle = "point cloud point density"
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "top"
    , strip.text.x = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , strip.text.y = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , plot.title = ggplot2::element_text(size = 10)
    , plot.subtitle = ggplot2::element_text(size = 10)
  ) +
  ggplot2::guides(
    color = ggplot2::guide_legend(override.aes = list(shape = 15, size = 6))
  )
```

flight speed has a greater impact on point density than altitude with decreasing point density as altitude increases at the same flight speed

how is the processing time related to speed and altitude?

```{r}
timer_df %>% 
  ggplot2::ggplot(
    mapping = aes(x = flight_altitude_m, y = timer_total_dbhonly_time_mins_per_ha, color = factor(flight_meters_per_second))
  ) +
  ggplot2::geom_point(size = 5.5, alpha = 0.9) +
  ggplot2::scale_color_manual(
    values = viridis::rocket(n = 3, begin = 0.9, end = 0.5)[c(1,3)]
  ) +
  ggplot2::scale_x_continuous(
    labels=scales::comma
    , breaks = scales::breaks_extended(n=8)
  ) +
  ggplot2::labs(
    color = latex2exp::TeX("Speed  $m \\cdot sec^{-1}$ ")
    , x = "Altitude (m)"
    , y = latex2exp::TeX("processing minutes  $ha^{-1}$")
    , subtitle = "point cloud point processing time"
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "top"
    , strip.text.x = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , strip.text.y = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , plot.title = ggplot2::element_text(size = 10)
    , plot.subtitle = ggplot2::element_text(size = 10)
  ) +
  ggplot2::guides(
    color = ggplot2::guide_legend(override.aes = list(shape = 15, size = 6))
  )
```

same relationship as with point density

how is point cloud processing time related to point density?

```{r}
timer_df %>% 
  ggplot2::ggplot(
    mapping = aes(
      x = points_per_m2, y = timer_total_dbhonly_time_mins_per_ha
      , shape = factor(flight_meters_per_second)
      , color = factor(flight_altitude_m)
    )
  ) +
  ggplot2::geom_point(size = 5.5, alpha = 0.9) +
  ggplot2::scale_color_viridis_d(option = "mako", begin = 0.5, end = 0.1) +
  ggplot2::scale_x_continuous(
    labels=scales::comma
    , breaks = scales::breaks_extended(n=8)
  ) +
  ggplot2::labs(
    shape = latex2exp::TeX("Speed  $m \\cdot sec^{-1}$ ")
    , x = latex2exp::TeX("points  $m^{-2}$ ")
    , color = "Altitude (m)"
    , y = latex2exp::TeX("processing minutes  $ha^{-1}$")
    , subtitle = "point cloud point processing time"
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "top"
    , strip.text.x = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , strip.text.y = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , plot.title = ggplot2::element_text(size = 10)
    , plot.subtitle = ggplot2::element_text(size = 10)
  ) +
  ggplot2::guides(
    color = ggplot2::guide_legend(override.aes = list(shape = 15, size = 6))
  )
```

point cloud processing section time

```{r}
timer_df %>% 
  dplyr::select(
    dataset_factor
    # timer
    , c( tidyselect::ends_with("_pct") & tidyselect::starts_with("timer_") )
  ) %>% 
  tidyr::pivot_longer(
    cols = -c(dataset_factor)
  ) %>% 
  dplyr::mutate(
    name = name %>% 
      stringr::str_remove_all("^timer_trees_") %>% 
      stringr::str_remove_all("^timer_") %>% 
      stringr::str_remove_all("timer_") %>% 
      stringr::str_remove_all("_mins_pct$") %>% 
      factor(
        ordered = T
        , levels = c(
          "tree_extraction"
          , "dbh"
          , "cbh"
          , "hmd"
          , "type"
          , "biomass"
        )
        , labels = c(
          "Tree Extraction"
          , "DBH", "CBH", "HMD", "forest type", "biomass"
        )
      ) %>% forcats::fct_rev()
    # , value_lab = scales::percent(value,accuracy=0.1)
    , dataset_factor = forcats::fct_rev(dataset_factor)
  ) %>%
ggplot2::ggplot(
  mapping = ggplot2::aes(x = value, y = dataset_factor, fill=name, group=name)
) +
  ggplot2::geom_col(
    width = 0.7, alpha=0.95
  ) +
  ggplot2::geom_text(
    mapping = aes(
        label = scales::percent(ifelse(value>=0.08,value,NA), accuracy = 1)
        , fontface = "bold"
      )
    , position = ggplot2::position_stack(vjust = 0.5)
    , color = "black", size = 4
  ) +
  ggplot2::scale_fill_brewer(palette = "Set2") +
  ggplot2::scale_x_continuous(labels = scales::percent_format()) +
  ggplot2::labs(
    fill = "processing\nsection"
    , y = ""
    , x = "% Point Cloud Total Processing Time"
    , subtitle = "point cloud total processing time by section"
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "top"
    , legend.direction  = "horizontal"
    , legend.title = element_text(size=7)
    , axis.title.x = element_text(size=10, face = "bold")
    , axis.text.x = element_blank()
    , axis.ticks.x = element_blank()
    , axis.text.y = ggplot2::element_text(size = 9, face = "bold")
  ) +
  ggplot2::guides(
    fill = guide_legend(nrow = 2, byrow = T, reverse = T, override.aes = list(alpha = 0.9))
  )  
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# cloud2trees:::search_dir_final_detected(c2t_process_dir)
remove(list = ls()[grep("_temp",ls())])
gc()
```