# Accuracy Results

We have finally arrived to the part where we review the results of the accuracy of the UAS-derived forest inventory based on:

* [**Detection Accuracy**](#detect_acc): tree-level instance segmentation matching pairing UAS-detected trees with stem-mapped field trees. The resulting True Positive, False Positive, and False Negative counts are used to calculate precision, recall, and F-score.
* [**Quantification Accuracy**](#quant_acc): tree-level accuracy of physical attributes of tree height and DBH for the successfully matched tree pairs. We use RMSE to capture typical error magnitude, MAPE to express relative precision as a percentage, and ME to quantify overestimation or underestimation of tree form.
* [**Stand-Level Accuracy**](#stand_acc): assessment of stand-level measurement accuracy that compares aggregated forest inventory metrics derived from the entire field inventory against all UAS-predicted trees. We calculate absolute bias and relative percent differences for stand basal area and trees per hectare across the different flight configurations to evaluate total structural summary performance.

aggregate the instance matching to get detection and quantification accuracy for each data set

```{r}
# aggregate field trees for stand-level comp
ref_temp <-
  validation_trees %>% 
  sf::st_drop_geometry() %>% 
  dplyr::ungroup() %>% 
  dplyr::summarise(
    # dplyr::across(
    #   c(field_tree_height_m, field_dbh_cm)
    #   , .fns = list(
    #     mean = ~mean(.x,na.rm=T)
    #     , sd = ~sd(.x,na.rm=T)
    #     # , q10 = ~quantile(.x,na.rm=T,probs=0.1)
    #     , median = ~quantile(.x,na.rm=T,probs=0.5)
    #     # , q90 = ~quantile(.x,na.rm=T,probs=0.9)
    #     , min = ~min(.x,na.rm=T)
    #     , max = ~max(.x,na.rm=T)
    #   )
    # )
    ref_n_trees = dplyr::n()
    , ref_basal_area_m2 = sum(basal_area_m2)
  )
  # dplyr::rename_with(.fn = ~stringr::str_remove(.x,"^field_"))  

# agg_ground_truth_match for all and add stand-level comp
agg_ground_truth_match_ans <-
  ground_truth_prediction_match_ans %>% 
  purrr::compact() %>% 
  purrr::imap(
    function(x,nm){
      # agg stand level for matching
      p <- 
        predicted_trees[[nm]] %>% 
        sf::st_drop_geometry() %>% 
        dplyr::filter(is_in_stand) %>% 
        dplyr::ungroup() %>% 
        dplyr::summarise(
          # dplyr::across(
          #   c(tree_height_m, dbh_cm)
          #   , .fns = list(
          #     mean = ~mean(.x,na.rm=T)
          #     , sd = ~sd(.x,na.rm=T)
          #     # , q10 = ~quantile(.x,na.rm=T,probs=0.1)
          #     , median = ~quantile(.x,na.rm=T,probs=0.5)
          #     # , q90 = ~quantile(.x,na.rm=T,probs=0.9)
          #     , min = ~min(.x,na.rm=T)
          #     , max = ~max(.x,na.rm=T)
          #   )
          # )
          pred_n_trees = dplyr::n()
          , pred_basal_area_m2 = sum(basal_area_m2)
        )
      # agg_ground_truth_match
        agg_ground_truth_match(x) %>% 
        dplyr::bind_cols(p,ref_temp) %>% 
        # comps not done in agg_ground_truth_match()
        dplyr::mutate(
          stand_area_ha = stand_boundary$stand_area_m2[1]/10000
          , dplyr::across(
            c(ref_basal_area_m2,pred_basal_area_m2)
            , ~ .x/stand_area_ha
            , .names = "{.col}_per_ha"
          )
          , ref_trees_per_ha = ref_n_trees/stand_area_ha
          , pred_trees_per_ha = pred_n_trees/stand_area_ha
          # 'diff_' columns are calculated as the predicted value minus the actual value
          , diff_trees_per_ha = pred_trees_per_ha-ref_trees_per_ha
          , diff_basal_area_m2 = pred_basal_area_m2-ref_basal_area_m2
          , diff_basal_area_m2_per_ha = pred_basal_area_m2_per_ha-ref_basal_area_m2_per_ha
          # 'abs_diff_' columns are calculated as the predicted value minus the actual value
          , abs_diff_trees_per_ha = abs(diff_trees_per_ha)
          , abs_diff_basal_area_m2 = abs(diff_basal_area_m2)
          , abs_diff_basal_area_m2_per_ha = abs(diff_basal_area_m2_per_ha)
          # 'pct_diff_' columns are calculated as the predicted value minus the actual value divided by the actual value
          , pct_diff_trees_per_ha = (pred_trees_per_ha-ref_trees_per_ha)/ref_trees_per_ha
          , pct_diff_basal_area_m2 = (pred_basal_area_m2-ref_basal_area_m2)/ref_basal_area_m2
          , pct_diff_basal_area_m2_per_ha = (pred_basal_area_m2_per_ha-ref_basal_area_m2_per_ha)/ref_basal_area_m2_per_ha
        )
  }) %>% 
  purrr::list_rbind(names_to = "dataset_factor") 


# add tracking data cols
agg_ground_truth_match_ans <- tracking_df %>% 
  dplyr::left_join(
    agg_ground_truth_match_ans
    , by = "dataset_factor"
  ) %>% 
  dplyr::mutate(
    dataset_factor = ordered(dataset_factor, levels(tracking_df$dataset_factor))
  )
# # huh?
# agg_ground_truth_match_ans %>% dplyr::glimpse()
# levels(agg_ground_truth_match_ans$dataset_factor)
```

## Detection Accuracy Results{#detect_acc}

let's table that

```{r}
agg_ground_truth_match_ans %>% 
  dplyr::arrange(dataset_factor) %>% 
  dplyr::select(
    tidyselect::starts_with("flight_")
    , tidyselect::ends_with("_n")
    , precision, recall, f_score
  ) %>% 
  dplyr::mutate(
    dplyr::across(
      tidyselect::starts_with("flight_")
      , as.factor
    )
    , dplyr::across(
      tidyselect::ends_with("_n")
      , ~scales::comma(.x,accuracy=1)
    )
    , dplyr::across(
      dplyr::where(is.numeric)
      , ~scales::percent(.x,accuracy=0.1)
    )
  ) %>% 
  kableExtra::kbl(
    caption = "Detection Accuracy"
    , col.names = c(
      "altitude (m)", "speed (m sec<sup>-1</sup>)"
      , "TP predictions", "FP predictions", "FN predictions"
      , "Precision", "Recall", "F-score"
    )
    , escape = F
    # , digits = 2
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::column_spec(c(2), border_right = TRUE, include_thead = TRUE) %>% 
  kableExtra::collapse_rows(columns = c(1), valign = "top")
```

and plot it

```{r}
#pal
pal_eval_metric <- c(
  "F-score" = RColorBrewer::brewer.pal(3,"Oranges")[3]
  , "Precision" = RColorBrewer::brewer.pal(3,"Purples")[3]
  , "Recall" = RColorBrewer::brewer.pal(3,"Greys")[3]
)
#plt
agg_ground_truth_match_ans %>% 
  dplyr::arrange(dataset_factor) %>% 
  dplyr::select(
    dataset_factor
    , precision, recall, f_score
  ) %>% 
  tidyr::pivot_longer(
    cols = c(precision,recall,f_score)
    , names_to = "metric"
    , values_to = "value"
  ) %>% 
  dplyr::mutate(
    metric = metric %>%
      forcats::fct_relevel("f_score", "recall", "precision") %>%
      forcats::fct_recode("F-score" = "f_score", "Recall" = "recall", "Precision" = "precision")
  ) %>% 
  dplyr::mutate(
    dplyr::across(
      dplyr::where(is.numeric)
      , list(lab = ~scales::percent(.x,accuracy=0.1))
    )
  ) %>% 
  dplyr::mutate(dataset_factor = forcats::fct_rev(dataset_factor)) %>% 
  dataset_factor_relabel_fn() %>% 
  # dplyr::glimpse()
  # dplyr::pull(metric) %>% levels()
  ggplot2::ggplot(
    mapping = ggplot2::aes(
      x = value, y = dataset_factor
      , fill = metric
    )
  ) +
  ggplot2::geom_col(width = 0.4) +
  ggplot2::geom_text(
    mapping = ggplot2::aes(label = value_lab, fontface = "bold")
    , color = "black"
    # , size = 2.3
    , hjust = -0.1
  ) +
  ggplot2::scale_fill_manual(values=pal_eval_metric) +
  ggplot2::scale_x_continuous(
    limits = c(0,1)
    , labels=scales::percent
    , breaks = scales::breaks_extended(n=6)
    , expand = ggplot2::expansion(mult = c(0,0.03))
  ) +
  ggplot2::facet_grid(
    cols = dplyr::vars(metric)
    , scales = "free_x"
    # , axes = "all"
  ) +
  ggplot2::labs(
    x="",y=""
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "none"
    , axis.text.y = ggplot2::element_text(size = 9, face = "bold")
    , axis.text.x = ggplot2::element_blank()
    , strip.text = ggplot2::element_text(size = 10, color = "black", face = "bold")
  ) +
  ggplot2::guides(
    color = ggplot2::guide_legend(override.aes = list(shape = 15, linetype = 0, size = 6, alpha = 1))
  )
```

It doesn't look like there is much variation in detection accuracy across the different flight settings. Furthermore, since we have exactly one observation per flight (e.g. no repeated measures with these flight settings such as from different study sites) we do not have any variation within the flight groups and have measured the "truth" for the particular flight settings. As such, we cannot statistically test one flight (with those exact settings) against another. However, we can calculate the raw differences.

```{r}
# calc pairwise diffs
# dplyr::glimpse(agg_ground_truth_match_ans)
pairwise_diffs_temp <-
  tidyr::crossing(
    agg_ground_truth_match_ans %>% 
      dplyr::select(dataset_factor,f_score) %>% 
      dataset_factor_relabel_fn() %>% 
      dplyr::rename_with(~paste0(.x,"_1",recycle0 = T))
    , agg_ground_truth_match_ans %>% 
      dplyr::select(dataset_factor,f_score) %>% 
      dataset_factor_relabel_fn() %>% 
      dplyr::rename_with(~paste0(.x,"_2",recycle0 = T))
  ) %>% 
  # filter to remove self-joins and duplicates
  dplyr::filter(dataset_factor_1 < dataset_factor_2) %>%
  dplyr::mutate(
    diff = f_score_1 - f_score_2
    , abs_diff = abs(diff)
    , comp = paste(dataset_factor_1, "vs", dataset_factor_2)
  )
```

pairwise difference table

```{r}
pairwise_diffs_temp %>%
  dplyr::select(dataset_factor_1, dataset_factor_2, f_score_1, f_score_2, diff) %>%
  dplyr::mutate(dplyr::across(dplyr::where(is.numeric), ~scales::percent(.x,accuracy=0.1) )) %>% 
  kableExtra::kbl(
    caption = "Pairwise Differences in F-Score between Flights"
    , col.names = c(
      "Flight"
      , "Comparison Flight"
      , "F-score"
      , "Comparison F-score"
      , "F-score Difference"
    )
    , digits = 3
  ) %>%
  kableExtra::kable_styling() %>% 
  kableExtra::collapse_rows(columns = 1, valign = "top")
```

pairwise difference plot

```{r,fig.height=9.5}
ggplot2::ggplot(
  data = pairwise_diffs_temp
  , mapping = ggplot2::aes(y = reorder(comp, diff), x = diff)
) +
  ggplot2::geom_vline(xintercept = 0, color = "black") +
  ggplot2::geom_segment(
    mapping = ggplot2::aes(yend = comp, xend = 0)
    , color = "gray44"
  ) +
  ggplot2::geom_point(
    mapping = ggplot2::aes(color = diff > 0)
    , size = 3
  ) +
  ggplot2::scale_color_manual(values = c("firebrick","navy")) +
  ggplot2::scale_x_continuous(
    limits = c(-0.05,0.05)
    , labels = scales::percent
    , breaks = scales::breaks_extended(n=10)
    , sec.axis = ggplot2::dup_axis()
  ) +
  ggplot2::labs(
    y = "" # "comparison"
    , x = "difference F-score"
    , subtitle = "positive = first flight in pair scored higher"
  ) +
  ggplot2::theme_minimal() +
  ggplot2::theme(legend.position = "none", plot.subtitle = ggplot2::element_text(hjust = 1))

```

so, the difference in F-score is no more than `r scales::percent( max( abs(pairwise_diffs_temp$diff) ), accuracy = 0.01)` across all flights. yes, that is less than `r paste0( ceiling( max( abs(pairwise_diffs_temp$diff) )*100 ), "%")` on a percentage point basis

```{r}
agg_ground_truth_match_ans %>% 
  ggplot2::ggplot(
    mapping = ggplot2::aes(x = flight_meters_per_second, y = f_score, color = as.factor(flight_altitude_m))
  ) +
  ggplot2::geom_point(size = 5,alpha=0.9) +
  ggplot2::geom_line(
    mapping = ggplot2::aes(group = flight_altitude_m)
  ) +
  ggplot2::geom_text(
    mapping = ggplot2::aes(label = scales::percent(f_score,accuracy=0.1))
    , color = "black"
    , vjust = -1
  ) +
  ggplot2::scale_y_continuous(
    # limits = c(0,1)
    # , breaks = seq(0,1,by=0.2)
    labels = scales::percent
    , expand = ggplot2::expansion(mult = 0.4)
  ) +
  ggplot2::scale_color_viridis_d(option = "mako", begin = 0.5, end = 0.1) +
  ggplot2::labs(
    subtitle = "comparison of speed and altitude"
    , x = latex2exp::TeX("speed  $m \\cdot sec^{-1}$ ")
    , y = "F-score"
    , color = "altitude (m)"
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "top"
    , strip.text = ggplot2::element_text(color = "black", size = 9, face = "bold")
  ) +
  ggplot2::guides(
    color = ggplot2::guide_legend(override.aes = list(shape = 15, size = 6))
  )

```

note that y-axis range ::laugh_cry_emoji::

let's see the relationship between tree detection (recall rate) and tree height within each flight. we'll look at true positive and omissions (FN) and use the field-measured tree height to quantify the proportion of trees within a height bin that were detected by the UAS-lidar

```{r, fig.height=10}
# make ggplot2::cut_* bins pretty
make_bins_pretty <- function(bin_factor, suffix = "") {
  # convert factor to character and remove brackets and parens
  clean_labels <- bin_factor %>%
    as.character() %>%
    stringr::str_replace_all(c("\\[" = "", "\\(" = "", "\\]" = "", "\\)" = "")) %>%
    stringr::str_replace(",", "-")
  
  # suffix
  if(stringr::str_squish(suffix) != "" && inherits(suffix,"character")){
    clean_labels <- stringr::str_c(clean_labels, " ", suffix) %>% stringr::str_squish()
  }
  
  # reorder based on the original factor levels 
  forcats::fct_reorder(factor(clean_labels), as.numeric(bin_factor))
}

dta_temp <-
  ground_truth_prediction_match_ans %>%
  purrr::list_rbind(names_to = "dataset_factor") %>%
  dplyr::mutate(
    dataset_factor = ordered(dataset_factor, levels(tracking_df$dataset_factor))
  ) %>% 
  sf::st_drop_geometry() %>% 
  dplyr::filter(match_grp %in% c("true positive","omission")) %>% 
  # dplyr::glimpse()
  # metric cuts/bins
  dplyr::mutate(
    tree_height_m_bin = ggplot2::cut_width(ref_tree_height_m, width = 5, boundary = 0, closed = "left") %>% 
      make_bins_pretty(suffix = "m") %>% 
      forcats::fct_relabel(~ stringr::str_replace(.x, "^0-", paste0(my_tree_ht_m,"-")) )
    , is_detected = match_grp %in% c("true positive")
    , is_detected = as.numeric(is_detected)
  )
dta_temp %>% 
  dplyr::count(dataset_factor, match_grp, tree_height_m_bin) %>% 
  dplyr::group_by(dataset_factor, tree_height_m_bin) %>% 
  dplyr::mutate(pct = n/sum(n,na.rm = T)) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(
    lab = ifelse(
      pct>0.095
      , paste0(
        scales::percent(pct,accuracy=1) 
        ,"\n("
        , scales::comma(n,accuracy=1) 
        , ")"
      )
      , ""
    )
  ) %>% 
  dplyr::ungroup() %>% 
  dataset_factor_relabel_fn() %>% 
  # dplyr::glimpse()
  ggplot2::ggplot(
    mapping = ggplot2::aes(y = pct, x = tree_height_m_bin)
  ) +
  ggplot2::geom_col(
    mapping = ggplot2::aes(fill = match_grp)
    , width = 0.6
    , color = NA, alpha = 0.9
  ) +
  ggplot2::geom_text(
    mapping = ggplot2::aes(label = lab, group = match_grp, fontface = "bold", color = match_grp)
    # , color = "black"
    , size = 2.5
    , position = ggplot2::position_stack(vjust = 0.5)
  ) +
  ggplot2::scale_fill_manual(values = pal_match_grp) +
  ggplot2::scale_color_manual(values = c("white","black"), guide = "none") +
  ggplot2::scale_x_discrete(drop = F) +
  ggplot2::scale_y_continuous(
    breaks = seq(0,1,by=0.2)
    , labels = scales::percent
    , expand = ggplot2::expansion(mult = c(0,0.02))
  ) +
  ggplot2::facet_wrap(
    facets = dplyr::vars(dataset_factor)
    # , rows = dplyr::vars(trtmnt_block)
    , ncol = 2
    , axes = "all"
    # , switch = "y"
  ) +
  ggplot2::labs(
    x = "tree height", y = "", fill = ""
    , subtitle = "Recall rate by tree height bin"
  ) +
  ggplot2::theme_light()+
  ggplot2::theme(
    legend.position = "top"
    , strip.text = ggplot2::element_text(size = 11, color = "black", face = "bold")
    , axis.text.x = ggplot2::element_text(size = 7, angle = 45, hjust = 1, vjust = 1)
    , axis.text.y = ggplot2::element_blank()
    , axis.ticks.y = ggplot2::element_blank()
  )
# ggplot2::ggsave("c:/users/georg/Downloads/n1_recall_by_htbin.jpg", height = 10, width = 8)
```

the trend of increasing detection rate with tree height demonstrated with this data aligns well with prior work using aerial point cloud data for single-tree forest inventories

let's check this for merchantable versus sub-merchantable trees. we'll use 10 cm (`r scales::comma(10/2.54,accuracy=0.1,suffix = " in")`) DBH as a threshold for classifying merchantable and sub-merchantable since it is a common value used in regional forest inventories.

```{r, fig.height=10, fig.width=8.3}
# 10 cm DBH, a common regional forest inventory breakpoint for describing merchantable and sub-merchantable forest densities
ostory_dbh <- 10 # = ~ 4 inches *2.54
# plt
dta_temp <- 
  ground_truth_prediction_match_ans %>%
  purrr::list_rbind(names_to = "dataset_factor") %>%
  dplyr::mutate(
    dataset_factor = ordered(dataset_factor, levels(tracking_df$dataset_factor))
  ) %>% 
  sf::st_drop_geometry() %>% 
  dplyr::filter(match_grp %in% c("true positive","omission")) %>% 
  # dplyr::glimpse()
  # metric cuts/bins
  dplyr::mutate(
    tree_dbh_cm_bin = 
      cut(ref_dbh_cm, breaks = c(0,ostory_dbh,ceiling(max(validation_trees$field_dbh_cm))), right = F) %>% 
      # ggplot2::cut_width(ref_dbh_cm, width = ostory_dbh, boundary = 0, closed = "left")
      make_bins_pretty(suffix = " cm") %>%
      forcats::fct_relabel(~ stringr::str_replace(.x, "^0-", paste0(floor(min(validation_trees$field_dbh_cm)),"-")) )
    , is_detected = match_grp %in% c("true positive")
    , is_detected = as.numeric(is_detected)
  ) %>% 
  dplyr::count(dataset_factor, match_grp, tree_dbh_cm_bin) %>% 
  dplyr::group_by(dataset_factor, tree_dbh_cm_bin) %>% 
  dplyr::mutate(pct = n/sum(n,na.rm = T)) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(
    lab = ifelse(
      pct>0.095
      , paste0(
        scales::percent(pct,accuracy=1) 
        ,"\n("
        , scales::comma(n,accuracy=1) 
        , ")"
      )
      , ""
    )
  ) %>% 
  dplyr::ungroup()
  
dta_temp %>% 
  dataset_factor_relabel_fn() %>% 
  # dplyr::glimpse()
  ggplot2::ggplot(
    mapping = ggplot2::aes(y = pct, x = tree_dbh_cm_bin)
  ) +
  ggplot2::geom_col(
    mapping = ggplot2::aes(fill = match_grp)
    , width = 0.55
    , color = NA, alpha = 0.9
  ) +
  ggplot2::geom_text(
    mapping = ggplot2::aes(label = lab, group = match_grp, fontface = "bold", color = match_grp)
    # , color = "black"
    , size = 3
    , position = ggplot2::position_stack(vjust = 0.5)
  ) +
  ggplot2::scale_fill_manual(values = pal_match_grp) +
  ggplot2::scale_color_manual(values = c("white","black"), guide = "none") +
  ggplot2::scale_x_discrete(drop = F) +
  ggplot2::scale_y_continuous(
    breaks = seq(0,1,by=0.2)
    , labels = scales::percent
    , expand = ggplot2::expansion(mult = c(0,0.02))
  ) +
  ggplot2::facet_wrap(
    facets = dplyr::vars(dataset_factor)
    # , rows = dplyr::vars(trtmnt_block)
    , ncol = 2
    , axes = "all"
    # , switch = "y"
  ) +
  ggplot2::labs(
    x = "reference tree DBH class"
    , y = "", fill = ""
    , subtitle = "Recall by reference tree DBH bin"
  ) +
  ggplot2::theme_light()+
  ggplot2::theme(
    legend.position = "top"
    , strip.text = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , axis.text.x = ggplot2::element_text(size = 8.5, angle = 0, face = "bold")
    , axis.text.y = ggplot2::element_blank()
    , axis.ticks.y = ggplot2::element_blank()
    , panel.grid.major.x = ggplot2::element_blank()
    , panel.grid.minor.x = ggplot2::element_blank()
    , plot.subtitle = ggplot2::element_blank()
  )
```

```{r, include=F, eval=T, results=F}
ggplot2::ggsave("../data/det_acc_by_dbh_bin.jpg", height = 10, width = 8.3)
```

table that

```{r}
dta_temp %>% 
  dplyr::select(-c(n,pct)) %>% 
  dplyr::mutate(
    match_grp = stringr::str_replace_all(match_grp," ","_")
    , lab = stringr::str_replace_all(lab, "\n","<br />")
  ) %>% 
  tidyr::pivot_wider(names_from = match_grp,values_from = lab) %>% 
  dplyr::inner_join(
    tracking_df %>% 
      dplyr::select(dataset_factor, tidyselect::starts_with("flight_"))
    , by = "dataset_factor"
  ) %>% 
  dplyr::arrange(dataset_factor, tree_dbh_cm_bin) %>% 
  dplyr::select(
    -c(
      dataset_factor
    )
  ) %>% 
  dplyr::mutate(
    dplyr::across(
      tidyselect::starts_with("flight_")
      , as.factor
    )
  ) %>% 
  dplyr::relocate(c(tree_dbh_cm_bin, true_positive)) %>% 
  dplyr::relocate(tidyselect::starts_with("flight_")) %>% 
  # ncol()
  kableExtra::kbl(
    caption = "Tree Detection by Tree Size"
    , col.names = c(
      "altitude (m)", "speed (m sec<sup>-1</sup>)"
      , "DBH bin"
      , "true positive (TP)", "omission (FN)"
    )
    , escape = F
    # , digits = 2
  ) %>% 
  kableExtra::kable_styling(font_size = 11.5) %>% 
  kableExtra::collapse_rows(columns = c(1:2), valign = "top")
```

let's look at the commission rate (commission error) by tree height using the UAS-detected tree heights to determine if our ITD variable window function could be improved

```{r, fig.height=10}
dta_temp <-
  ground_truth_prediction_match_ans %>%
  purrr::list_rbind(names_to = "dataset_factor") %>%
  dplyr::mutate(
    dataset_factor = ordered(dataset_factor, levels(tracking_df$dataset_factor))
  ) %>% 
  sf::st_drop_geometry() %>% 
  dplyr::filter(match_grp %in% c("true positive","commission")) %>% 
  # dplyr::glimpse()
  # metric cuts/bins
  dplyr::mutate(
    tree_height_m_bin = ggplot2::cut_width(pred_tree_height_m, width = 5, boundary = 0, closed = "left") %>% 
      make_bins_pretty(suffix = "m") %>% 
      forcats::fct_relabel(~ stringr::str_replace(.x, "^0-", paste0(my_tree_ht_m,"-")) )
    , is_detected = match_grp %in% c("true positive")
    , is_detected = as.numeric(is_detected)
  )
dta_temp %>% 
  dplyr::count(dataset_factor, match_grp, tree_height_m_bin) %>% 
  dplyr::group_by(dataset_factor, tree_height_m_bin) %>% 
  dplyr::mutate(pct = n/sum(n,na.rm = T)) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(
    lab = ifelse(
      pct>0.095
      , paste0(
        scales::percent(pct,accuracy=1) 
        ,"\n("
        , scales::comma(n,accuracy=1) 
        , ")"
      )
      , ""
    )
    , match_grp = forcats::fct_rev(match_grp)
  ) %>% 
  dplyr::ungroup() %>% 
  dataset_factor_relabel_fn() %>% 
  # dplyr::glimpse()
  ggplot2::ggplot(
    mapping = ggplot2::aes(y = pct, x = tree_height_m_bin)
  ) +
  ggplot2::geom_col(
    mapping = ggplot2::aes(fill = match_grp)
    , width = 0.6
    , color = NA, alpha = 0.9
  ) +
  ggplot2::geom_text(
    mapping = ggplot2::aes(label = lab, group = match_grp, fontface = "bold", color = match_grp)
    , color = "black"
    , size = 2.5
    , position = ggplot2::position_stack(vjust = 0.5)
  ) +
  ggplot2::scale_fill_manual(values = pal_match_grp) +
  # ggplot2::scale_color_manual(values = c("black","white"), guide = "none") +
  ggplot2::scale_x_discrete(drop = F) +
  ggplot2::scale_y_continuous(
    breaks = seq(0,1,by=0.2)
    , labels = scales::percent
    , expand = ggplot2::expansion(mult = c(0,0.02))
  ) +
  ggplot2::facet_wrap(
    facets = dplyr::vars(dataset_factor)
    # , rows = dplyr::vars(trtmnt_block)
    , ncol = 2
    , axes = "all"
    # , switch = "y"
  ) +
  ggplot2::labs(
    x = "tree height", y = "", fill = ""
    , subtitle = "Commission rate by tree height bin"
  ) +
  ggplot2::theme_light()+
  ggplot2::theme(
    legend.position = "top"
    , strip.text = ggplot2::element_text(size = 11, color = "black", face = "bold")
    , axis.text.x = ggplot2::element_text(size = 7, angle = 45, hjust = 1, vjust = 1)
    , axis.text.y = ggplot2::element_blank()
    , axis.ticks.y = ggplot2::element_blank()
  )
# ggplot2::ggsave("c:/users/georg/Downloads/n1_commission_by_htbin.jpg", height = 10, width = 8)
```

the commission error (percentage shown in the gray bar) is highest for the intermediate trees which suggests we could potentially improve our ITD window function to decrease the number of trees detected in this middle height range. However, going back to the detection rate above, this same height range also had the lowest detection rate which would suggest we increase the number of trees identified in this height range. This conflict suggests that the trees in this height range are likely difficult to detect for the study area using the data and ITD methods used in this study. For example, they may be subordinate to directly adjacent taller trees with the crowns interlocking, so they are often grouped with the taller tree during the CHM-based ITD.

## Quantification Accuracy Results{#quant_acc}

let's table the tree-level aggregated measurement error

```{r}
agg_ground_truth_match_ans %>% 
  dplyr::mutate(
    ref_trees = tp_n+fn_n
    , det_trees = tp_n+fp_n
  ) %>% 
  dplyr::select(
    dataset_factor, tidyselect::starts_with("flight_")
    # , site_area_m2
    # detection
    , ref_trees
    , det_trees
    , tp_n
    , omission_rate,commission_rate,recall,precision,f_score
    # quantification
    , tidyselect::ends_with("_mean")
    , tidyselect::ends_with("_rmse")
    , tidyselect::ends_with("_mape")
  ) %>% 
  # second select to arrange tree_metric
  dplyr::select(
    dataset_factor, tidyselect::starts_with("flight_")
    # , site_area_m2
    # detection
    , ref_trees
    , det_trees
    , tp_n
    , omission_rate,commission_rate,recall,precision,f_score
    # quantification
    # , c(tidyselect::contains("volume") & !tidyselect::contains("paraboloid"))
    , tidyselect::contains("area")
    , tidyselect::contains("height")
    , tidyselect::contains("diameter")
    , tidyselect::contains("dbh")
  ) %>% 
  # names()
  dplyr::mutate(
    dplyr::across(
      .cols = c(
        f_score, recall, precision, tidyselect::ends_with("_mape")
        , tidyselect::ends_with("_rate")
      )
      , .fn = ~ scales::percent(.x, accuracy = 1)
    )
    , dplyr::across(
      .cols = c(tidyselect::ends_with("_mean"))
      , .fn = ~ scales::comma(.x, accuracy = 0.01)
    )
    , dplyr::across(
      .cols = c(tidyselect::ends_with("_rmse"))
      , .fn = ~ scales::comma(.x, accuracy = 0.01)
    )
  ) %>% 
  dplyr::select(
    !tidyselect::contains("_area_")
    & !tidyselect::contains("diff_diameter_")
    & !tidyselect::ends_with("_trees")
    # & !tidyselect::ends_with("_n")
    & !tidyselect::ends_with("_rate")
  ) %>% 
  dplyr::arrange(dataset_factor) %>% 
  dplyr::select(
    -c(
      recall,precision,f_score
      , dataset_factor
    )
  ) %>% 
  dplyr::mutate(
    dplyr::across(
      tidyselect::starts_with("flight_")
      , as.factor
    )
  ) %>% 
  # dplyr::glimpse()
  
  kableExtra::kbl(
    caption = "Quantification Accuracy"
    , col.names = c(
      "altitude (m)", "speed (m sec<sup>-1</sup>)"
      , "TP predictions"
      , rep(c("ME","RMSE","MAPE"), times = 2)
    )
    , escape = F
    # , digits = 2
  ) %>% 
  kableExtra::kable_styling(font_size = 11.5) %>% 
  kableExtra::add_header_above(c(
    " "=3
    # , "Detection" = 3
    # , "Area" = 3
    , "Height (m)" = 3
    , "DBH (cm)" = 3
  )) %>% 
  kableExtra::column_spec(seq(3,8,by=3), border_right = TRUE, include_thead = TRUE) %>% 
  kableExtra::collapse_rows(columns = c(1), valign = "top")
```

and plot predicted versus reference values for correct predictions (TP)

```{r}
####################################
# area scatter
####################################
dta_temp <-
  ground_truth_prediction_match_ans %>% 
  purrr::compact() %>% 
  purrr::imap(
    \(x,nm) 
    dplyr::filter(x,match_grp=="true positive") %>% 
    sf::st_drop_geometry() %>% 
    dplyr::mutate(dataset_factor = nm)
  )%>% 
  dplyr::bind_rows() %>% 
  dplyr::left_join(
    agg_ground_truth_match_ans %>% 
      dplyr::select(dataset_factor, tidyselect::ends_with("_rmse") , tidyselect::ends_with("_mape"), tidyselect::ends_with("_mean") )
    , by = "dataset_factor"
  )
  # dplyr::glimpse()
# plt fn temp
plt_fn_temp <- function(
    x
    , which_col = "dbh_cm"
    , title_lab = "DBH"
    , x_lab = "reference DBH (cm)" # latex2exp::TeX("reference area ($m^2$)")
    , y_lab = "predicted DBH (cm)" # latex2exp::TeX("predicted area ($m^2$)")
){
    d <- dta_temp %>% 
      dplyr::filter(dataset_factor==x)
    # construct the formula: "y_var ~ x_var"
    y_var <- paste0("pred_",which_col)
    x_var <- paste0("ref_",which_col)
    formula_obj <- stats::as.formula(
      paste( 
        y_var
        , "~"
        , x_var 
      )
    )
    lm_temp <- lm(
      formula = formula_obj
      , data = d
    )
    # summary(lm_temp)
    # scales::percent(summary(lm_temp)$r.squared, accuracy = 0.1)
    lmf_temp <- paste0(
      "y = "
      , scales::number(summary(lm_temp)$coefficients[2,1], accuracy = 0.01) #slope
      , "x"
      , ifelse(summary(lm_temp)$coefficients[1,1]<0," - ", " + ")
      , scales::number(abs(summary(lm_temp)$coefficients[1,1]), accuracy = 0.01) #intercept
    )
    # scale limits
    max_val <- max( 
      max(dta_temp[[y_var]],na.rm=T)
      , max(dta_temp[[x_var]],na.rm=T)
    )
    
    # plot
    d %>% 
      dataset_factor_relabel_fn() %>% 
      ggplot2::ggplot(mapping = ggplot2::aes(y = .data[[y_var]], x = .data[[x_var]])) +
      ggplot2::geom_abline(color = "gray33", lwd = 1) +
      ggplot2::geom_point(color = "maroon4", alpha = 0.7, size = 0.9) +
      ggplot2::geom_smooth(lwd = 0.8, method = "lm", se=F, color = "gray", linetype = "dashed") +
      ggplot2::annotate(
        "text", 
        x = -Inf, y = Inf # Top Left
        , label = 
          # paste0(
          #   "R^2 == "
          #   , scales::number((summary(lm_temp)$r.squared)*100, accuracy = 0.1)
          #   , "*'%'"
          # )
          paste0(
            "RMSE = "
            , d %>% 
              dplyr::slice(1) %>% 
              dplyr::select(dplyr::all_of(paste0("diff_",which_col, "_rmse"))) %>% 
              dplyr::pull(1) %>% 
              unique() %>% 
              # scales::number(accuracy = 0.1, suffix = " (m²)")
              scales::number(
                accuracy = ifelse(which_col=="dbh_cm",0.1,0.01)
                , suffix = ifelse(which_col=="dbh_cm"," (cm)"," (m)") 
              )
            , "\nMAPE = "
            , d %>% 
              dplyr::select(dplyr::all_of(paste0("pct_diff_",which_col, "_mape"))) %>% 
              dplyr::pull(1) %>% 
              unique() %>% 
              scales::percent(accuracy = 0.1)
            , "\nME = "
            , d %>% 
              dplyr::select(dplyr::all_of(paste0("diff_",which_col, "_mean"))) %>% 
              dplyr::pull(1) %>% 
              unique() %>% 
              scales::number(
                accuracy = ifelse(which_col=="dbh_cm",0.1,0.01)
                , suffix = ifelse(which_col=="dbh_cm"," (cm)"," (m)") 
              )
            , "\n"
            , lmf_temp
            , "\n R² = "
            # , scales::number((summary(lm_temp)$r.squared)*100, accuracy = 0.1, suffix = "%")
            , scales::number((summary(lm_temp)$r.squared), accuracy = 0.01)
          )
        , hjust = -0.1, vjust = 1.1
        , parse = F
        , size = 3
      ) +
      ggplot2::facet_wrap(
        facets = dplyr::vars(dataset_factor), scales = "free", axis.labels = "all"
        # , labeller = ggplot2::labeller(class = ggplot2::label_wrap_gen(width = 10))
      ) +
      ggplot2::scale_x_continuous(limits = c(0, max_val ), expand = ggplot2::expansion(mult = c(0,0.1))) +
      ggplot2::scale_y_continuous(limits = c(0, max_val ), expand = ggplot2::expansion(mult = c(0,0.1))) +
      ggplot2::labs(
        x = x_lab
        , y = y_lab
        # , color = "image-field\ndiameter diff."
        # , subtitle = latex2exp::TeX("bulk volume ($m^3$) comparison")
        # , subtitle = paste("Predicted", title_lab,"versus Reference", title_lab)
      ) +
      ggplot2::theme_light() +
      ggplot2::theme(
        strip.text = ggplot2::element_text(size = 10, face = "bold", color = "black")
      ) 
  }
# get the plottttttttttttttttttt
dbh_plts_temp <- 
  unique(dta_temp$dataset_factor) %>% 
  purrr::map(
    \(z)
    plt_fn_temp(x=z, which_col = "dbh_cm", title_lab = "DBH", x_lab = "reference DBH (cm)", y_lab = "predicted DBH (cm)")
  )
# dbh_plts_temp[[1]]
dbh_plt_temp <- 
  patchwork::wrap_plots(dbh_plts_temp, ncol = 2) + 
  patchwork::plot_annotation(
    subtitle = "Predicted DBH versus Reference DBH"
    , theme = ggplot2::theme(plot.subtitle = ggplot2::element_blank()) # ggplot2::element_text(face = "bold"))
  ) &
  ggplot2::theme(
    axis.title = ggplot2::element_text(size = 6)
    , axis.text = ggplot2::element_text(size = 6)
  )
ggplot2::ggsave("../data/dbh_pred_v_ref.jpg", plot = dbh_plt_temp, height = 9.6, width = 8.7)
# dbh_plt_temp
# get the plottttttttttttttttttt
ht_plts_temp <- 
  unique(dta_temp$dataset_factor) %>% 
  purrr::map(
    \(z)
    plt_fn_temp(x=z, which_col = "tree_height_m", title_lab = "Height", x_lab = "reference Height (m)", y_lab = "predicted Height (m)")
  )
# ht_plts_temp[[1]]
ht_plt_temp <- 
  patchwork::wrap_plots(ht_plts_temp, ncol = 2) + 
  patchwork::plot_annotation(
    subtitle = "Predicted Height versus Reference Height"
    , theme = ggplot2::theme(plot.subtitle = ggplot2::element_blank()) # ggplot2::element_text(face = "bold"))
  ) &
  ggplot2::theme(
    axis.title = ggplot2::element_text(size = 6)
    , axis.text = ggplot2::element_text(size = 6)
  )
# ht_plt_temp
ggplot2::ggsave("../data/ht_pred_v_ref.jpg", plot = ht_plt_temp, height = 9.6, width = 8.7)
```

**DBH: predicted vs reference**

```{r, fig.height = 9.6, fig.width = 8.7}
dbh_plt_temp
```

the UAS-predicted DBH values are directionally correct as evidenced by the high correlation coefficient of approximately `r scales::comma(summary(lm(pred_dbh_cm~ref_dbh_cm , data = ground_truth_prediction_match_ans[[1]] %>% dplyr::filter(match_grp=="true positive")))$r.squared, accuracy=0.01)` across all tested flights. there is a size-dependent bias of DBH predictions with the UAS tending to overpredict the diameter of small and intermediate trees while underpredicting the diameter of the largest trees. the total measurement error remains reasonable for forest inventory with the RMSE of approximately `r scales::comma(median(agg_ground_truth_match_ans$diff_dbh_cm_rmse),accuracy = 0.01)` cm (`r scales::comma(median(agg_ground_truth_match_ans$diff_dbh_cm_rmse)/2.54,accuracy = 0.01)` in) and the average magnitude of the DBH prediction error quantified by MAPE is approximately `r scales::percent(median(agg_ground_truth_match_ans$pct_diff_dbh_cm_mape),accuracy = 1)` across the flights

**height: predicted vs reference**

```{r, fig.height = 9.6, fig.width = 8.7}
ht_plt_temp
```

UAS-predicted height is highly accurate compared to field-measured values, a finding supported by many prior studies. also, remember how we conditionally matched UAS-predicted trees with reference trees based on a maximum height difference threshold of `r my_tree_ht_m` m ?

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

full distribution of predicted and reference values irrespective of correct or incorrect prediction match

```{r, fig.height = 10, fig.width = 8.3}
ground_truth_prediction_match_ans %>%
  purrr::list_rbind(names_to = "dataset_factor") %>%
  dplyr::mutate(
    dataset_factor = ordered(dataset_factor, levels(tracking_df$dataset_factor))
  ) %>% 
  sf::st_drop_geometry() %>%
  dplyr::select(
    dataset_factor
    , ref_tree_height_m, pred_tree_height_m
    , ref_dbh_cm, pred_dbh_cm
  ) %>% 
  tidyr::pivot_longer(
    cols = -c(dataset_factor)
    , names_to = "metric"
    , values_to = "value"
    , values_drop_na = T
  ) %>% 
  dplyr::mutate(
    which_data = dplyr::case_when(
        stringr::str_starts(metric,"field_") | stringr::str_starts(metric,"ref_") ~ "field"
        , stringr::str_starts(metric,"pred_") ~ "prediction"
        , T ~ "error"
      ) %>% 
      ordered()
    , tree_metric = metric %>% 
      stringr::str_remove("(_rmse|_rrmse|_mean|_mape)$") %>% 
      stringr::str_extract("(height|dbh)") %>% 
      factor(
        ordered = T
        , levels = c(
          "height"
          , "dbh"
        )
        , labels = c(
          "Height (m)"
          , "DBH (cm)"
        )
      )
  ) %>% 
  dplyr::mutate(
    dataset_factor = dataset_factor %>% forcats::fct_relabel(~ str_replace_all(.x, "\\|", "\n"))
  ) %>% 
  dataset_factor_relabel_fn() %>% 
  # dplyr::glimpse()
# plot dist
  ggplot2::ggplot(mapping = ggplot2::aes(x = value, color = which_data, fill = which_data)) +
  ggplot2::geom_density(mapping = ggplot2::aes(y=ggplot2::after_stat(scaled)), alpha = 0.7) +
  ggplot2::facet_grid(
    rows = dplyr::vars(dataset_factor)
    , cols = dplyr::vars(tree_metric)
    , scales = "free_x", axes = "all_x"
    , switch = "y"
  ) +
  harrypotter::scale_color_hp_d(option = "hermionegranger") +
  harrypotter::scale_fill_hp_d(option = "hermionegranger") +
  ggplot2::scale_y_continuous(NULL,breaks=NULL) +
  ggplot2::scale_x_continuous(breaks=scales::breaks_extended(n=6)) +
  ggplot2::labs(
    color="",fill="",x=""
    # , subtitle = "comparison of height and DBH distributions"
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "top"
    , strip.text.x = ggplot2::element_text(size = 9, color = "black", face = "bold")
    , strip.text.y.left =  ggplot2::element_text(size = 8.5, angle = 0, color = "black", face = "bold")
    , axis.text.x = ggplot2::element_text(size = 6)
    , panel.grid = ggplot2::element_blank()
  )
# ggplot2::ggsave(filename = "../data/heydist.jpg", height = 10.5, width = 8, dpi = "print")
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
ggplot2::ggsave("../data/dists_pred_v_ref.jpg", height = 10, width = 8.3)
remove(list = ls()[grep("_temp",ls())])
gc()
```

let's dive deeper on the tree-level DBH accuracies. recall that over the entire range of tree sizes we found DBH RMSE values of ~`r agg_ground_truth_match_ans %>% dplyr::pull(diff_dbh_cm_rmse) %>% median() %>% scales::comma(accuracy = 0.1, suffix = " cm")`, ME of ~`r agg_ground_truth_match_ans %>% dplyr::pull(diff_dbh_cm_mean) %>% median() %>% scales::comma(accuracy = 0.1, suffix = " cm")`, with DBH MAPE of `r agg_ground_truth_match_ans %>% dplyr::pull(pct_diff_dbh_cm_mape) %>% median() %>% scales::percent(accuracy = 0.1)`. We also noticed that inaccuracies in DBH for small trees "pulled up" the linear regression fit line of predicted DBH on reference field DBH which diverged above the line of equality for stems of approximately <10 cm DBH.

let's shift our focus to relative error (MAPE) while providing absolute context (RMSE and ME) as text or secondary metrics to look into the penalty we face due to inaccuracies in DBH for small stems. for example, a 3 cm error on a 6 cm small tree is a 50% relative error while that same 3 cm error on a 30 cm merchantable tree is a minor 10% error. we'll use 10 cm (`r scales::comma(10/2.54,accuracy=0.1,suffix = " in")`) DBH as a threshold for classifying merchantable and sub-merchantable since it is a common value used in regional forest inventories.

```{r}
# do it by data set
dta_temp <- 
  ground_truth_prediction_match_ans %>% 
  purrr::map(
    \(x)
    x %>% 
      sf::st_drop_geometry() %>% 
      dplyr::filter(match_grp=="true positive") %>% 
      dplyr::mutate(
        tree_dbh_cm_bin = 
          cut(ref_dbh_cm, breaks = c(0,ostory_dbh,ceiling(max(validation_trees$field_dbh_cm))), right = F) %>% 
          # ggplot2::cut_width(ref_dbh_cm, width = ostory_dbh, boundary = 0, closed = "left")
          make_bins_pretty(suffix = " cm") %>%
          forcats::fct_relabel(~ stringr::str_replace(.x, "^0-", paste0(floor(min(validation_trees$field_dbh_cm)),"-")) )
      ) %>% 
      # dplyr::count(tree_dbh_cm_bin)
      dplyr::group_by(tree_dbh_cm_bin) %>% 
      dplyr::group_modify(~agg_ground_truth_match(.x)) %>% 
      dplyr::ungroup()
      # dplyr::glimpse()
  ) %>% 
  purrr::list_rbind(names_to = "dataset_factor") %>%
  dplyr::mutate(
    dataset_factor = ordered(dataset_factor, levels(tracking_df$dataset_factor))
  )
# dta_temp %>% dplyr::glimpse()
```

```{r, fig.height=10, fig.width=8.3}
dta_temp %>% 
  dplyr::rename_with(~stringr::str_remove(.x,"^diff_") %>% stringr::str_remove("^pct_diff_")) %>% 
  tidyr::pivot_longer(
    cols = tidyselect::starts_with(c("dbh_cm_", "tree_height_m_"))
    , names_to = "metric"
    , values_to = "value"
  ) %>% 
  dplyr::mutate(
    tree_metric = metric %>% 
      stringr::str_remove("(_rmse|_rrmse|_mean|_mape)$") %>% 
      stringr::str_extract("(height|dbh)") %>% 
      factor(
        ordered = T
        , levels = c(
          "height"
          , "dbh"
        )
        , labels = c(
          "Height (m)"
          , "DBH (cm)"
        )
      )
    , metric = stringr::str_extract(metric,"(_rmse|_rrmse|_mean|_mape)$") %>% 
      stringr::str_remove("^_")
  ) %>% 
  tidyr::pivot_wider(names_from = metric, values_from = value) %>% 
  dplyr::mutate(
    lab = 
      paste0(
        "MAPE = "
        , scales::percent(mape,accuracy = 0.1)
        , "\nRMSE = "
        , scales::number(
            rmse
            , accuracy = ifelse(stringr::str_starts(tolower(tree_metric),"dbh"), 0.1, 0.01)
            , suffix = paste0(" ",word(tree_metric,-1))
          )
        , "\nME = "
        , scales::number(
            mean
            , accuracy = ifelse(stringr::str_starts(tolower(tree_metric),"dbh"), 0.1, 0.01)
            , suffix = paste0(" ",word(tree_metric,-1))
          )
        # , "\nn ="
        # , scales::comma(tp_n, accuracy = 1)
      )
    , lab2 = 
      paste0(
        "RMSE = "
        , scales::number(
            rmse
            , accuracy = ifelse(stringr::str_starts(tolower(tree_metric),"dbh"), 0.1, 0.01)
            , suffix = paste0(" ",word(tree_metric,-1))
          )
        , "\nME = "
        , scales::number(
            mean
            , accuracy = ifelse(stringr::str_starts(tolower(tree_metric),"dbh"), 0.1, 0.01)
            , suffix = paste0(" ",word(tree_metric,-1))
          )
        # , "\nn ="
        # , scales::comma(tp_n, accuracy = 1)
      )
    , lab1 = 
      paste0(
        "MAPE = "
        , scales::percent(mape,accuracy = 0.1)
      )
    ## do something about the different factors to plot over
    , tree_metric = forcats::fct_relabel(tree_metric, ~ stringr::str_extract(.x, "^\\S+"))
    , xvar = forcats::fct_cross(tree_metric,tree_dbh_cm_bin)
  ) %>% 
  dataset_factor_relabel_fn() %>% 
  # dplyr::glimpse()
  # dplyr::filter(dataset_factor=="Altitude (m): 61 | Speed (m/s): 4.5") %>% 
ggplot2::ggplot(
  mapping = ggplot2::aes(
    y = mape
    , x = xvar #interaction(tree_metric,tree_dbh_cm_bin, lex.order = TRUE)
  )
) + 
  ggplot2::geom_vline(
    data = ~
      dplyr::distinct(.x, tree_metric, tree_dbh_cm_bin, xvar) %>%
      dplyr::mutate(xvar = as.numeric(xvar)) %>%
      dplyr::group_by(tree_metric) %>%
      dplyr::summarise(midx = mean(xvar),maxx = max(xvar)) %>% 
      dplyr::mutate(linex = maxx + 0.5) %>%
      dplyr::filter(dplyr::row_number() < dplyr::n())
    , ggplot2::aes(xintercept = linex)
    , color = "gray66", lwd = 0.8
  ) +
  ggplot2::geom_col(
    ggplot2::aes(fill = tree_dbh_cm_bin)
    , width = 0.55
    , color = NA, alpha = 0.95
    # , fill = "maroon4"
    , position = "dodge"
  ) +
  # add the parent factor labels
  ggplot2::geom_text(
    data = ~
      dplyr::distinct(.x, tree_metric, tree_dbh_cm_bin, xvar) %>%
      dplyr::mutate(xvar = as.numeric(xvar)) %>%
      dplyr::group_by(tree_metric) %>%
      dplyr::summarise(midx = mean(xvar))
    , ggplot2::aes(x = midx, y = Inf, label = tree_metric)
    , vjust = 1.5
    , fontface = "bold"
    , size = 3.4
    , inherit.aes = FALSE
  ) +
  ggplot2::geom_text(
    mapping = ggplot2::aes(label = lab1, fontface = "bold")
    # , color = "black"
     , color = "black", size = 2.6, vjust = -3.15
  ) +
  ggplot2::geom_text(
    mapping = ggplot2::aes(label = lab2)
    # , color = "black"
     , color = "black", size = 2, vjust = -0.2
  ) +
  ggplot2::facet_wrap(
    facets = dplyr::vars(dataset_factor)
    # , rows = dplyr::vars(trtmnt_block)
    , ncol = 2
    , axes = "all"
    # , axis.labels = "all_x"
    # , switch = "y"
  ) +
  
  ggplot2::scale_fill_manual(
    values = 
      c("darkslategray4","darkslategray")
      # c("deeppink3","maroon4")
  ) +
  ggplot2::scale_x_discrete(
    # get rid of parent factor label
    labels = ~stringr::str_remove(.x, "^.*:")
  ) + 
  ggplot2::scale_y_continuous(
    # breaks = seq(0,1,by=0.2)
    labels = scales::percent
    , breaks = scales::breaks_extended(n=4)
    , expand = ggplot2::expansion(mult = c(0,0.26))
  ) +
  ggplot2::labs(
    x = "reference tree DBH class"
    , y = "MAPE (relative error)"
    , subtitle = "DBH prediction accuracy by reference tree DBH bin"
  ) +
  ggplot2::theme_light()+
  ggplot2::theme(
    legend.position = "none"
    , strip.text = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , axis.text.x = ggplot2::element_text(size = 8.5, angle = 0, face = "bold")
    , axis.text.y = ggplot2::element_text(size = 5, angle = 0) #, hjust = 0.5)
    # , axis.text.y = ggplot2::element_blank()
    # , axis.ticks.y = ggplot2::element_blank()
    , panel.grid.major.x = ggplot2::element_blank()
    , panel.grid.minor.x = ggplot2::element_blank()
    , plot.subtitle = ggplot2::element_blank()
  )
ggplot2::ggsave("../data/dbh_acc_by_dbh_bin.jpg", height = 10, width = 8.3)
```

```{r, include=F, eval=F, fig.height=10, fig.width=8}
p_temp <- 
  dta_temp %>% 
  dplyr::mutate(
    lab = 
      paste0(
        "MAPE = "
        , scales::percent(pct_diff_dbh_cm_mape,accuracy = 0.1)
        , "\nRMSE = "
        , scales::number(
            diff_dbh_cm_rmse
            , accuracy = 0.1
            , suffix = " (cm)"
          )
        , "\nME = "
        , scales::number(
            diff_dbh_cm_mean
            , accuracy = 0.1
            , suffix = " (cm)"
          )
        # , "\nn ="
        # , scales::comma(tp_n, accuracy = 1)
      )
    , lab2 = 
      paste0(
        "RMSE = "
        , scales::number(
            diff_dbh_cm_rmse
            , accuracy = 0.1
            , suffix = " (cm)"
          )
        , "\nME = "
        , scales::number(
            diff_dbh_cm_mean
            , accuracy = 0.1
            , suffix = " (cm)"
          )
        # , "\nn ="
        # , scales::comma(tp_n, accuracy = 1)
      )
    , lab1 = 
      paste0(
        "MAPE = "
        , scales::percent(pct_diff_dbh_cm_mape,accuracy = 0.1)
      )
  ) %>% 
  # dplyr::glimpse()
  ggplot2::ggplot(
    mapping = ggplot2::aes(y = pct_diff_dbh_cm_mape, x = tree_dbh_cm_bin)
  ) +
  ggplot2::geom_col(
    width = 0.5
    , color = NA, alpha = 0.9, fill = "maroon4"
  ) +
  ggplot2::geom_text(
    mapping = ggplot2::aes(label = lab1, fontface = "bold")
    # , color = "black"
     , color = "black", size = 3, vjust = -3
  ) +
  ggplot2::geom_text(
    mapping = ggplot2::aes(label = lab2)
    # , color = "black"
     , color = "black", size = 2.1, vjust = -0.2
  ) +
  ggplot2::facet_wrap(
    facets = dplyr::vars(dataset_factor)
    # , rows = dplyr::vars(trtmnt_block)
    , ncol = 2
    , axes = "all"
    # , switch = "y"
  ) +
  ggplot2::scale_x_discrete(drop = F) +
  ggplot2::scale_y_continuous(
    # breaks = seq(0,1,by=0.2)
    labels = scales::percent
    , expand = ggplot2::expansion(mult = c(0,0.25))
  ) +
  ggplot2::labs(
    x = "", y = ""
    , subtitle = "DBH prediction accuracy by reference tree DBH bin"
  ) +
  ggplot2::theme_light()+
  ggplot2::theme(
    legend.position = "top"
    , strip.text = ggplot2::element_text(size = 11, color = "black", face = "bold")
    , axis.text.x = ggplot2::element_text(size = 9, angle = 45, hjust = 1, vjust = 1)
    , axis.text.y = ggplot2::element_blank()
    , axis.ticks.y = ggplot2::element_blank()
  )
p_temp
# ggplot2::ggsave("../data/dbh_acc_by_dbh_bin.jpg",p_temp+ggplot2::theme(plot.subtitle = ggplot2::element_blank()), height = 10, width = 8)
```

that looks like what we expected, with DBH MAPE of ~`r dta_temp %>% dplyr::filter(as.numeric(tree_dbh_cm_bin)==2) %>% dplyr::pull(pct_diff_dbh_cm_mape) %>% median() %>% scales::percent(accuracy = 0.1)` for merchantable trees and DBH MAPE of ~`r dta_temp %>% dplyr::filter(as.numeric(tree_dbh_cm_bin)==1) %>% dplyr::pull(pct_diff_dbh_cm_mape) %>% median() %>% scales::percent(accuracy = 0.1)` for sub-merchantable trees

table that

```{r}
dta_temp %>% 
  dplyr::inner_join(
    tracking_df %>% 
      dplyr::select(-c(path))
    , by = "dataset_factor"
  ) %>% 
  dplyr::select(
    dataset_factor, tidyselect::starts_with("flight_"), tree_dbh_cm_bin
    # , site_area_m2
    # detection
    , tp_n
    # quantification
    , tidyselect::ends_with("_mean")
    , tidyselect::ends_with("_rmse")
    , tidyselect::ends_with("_mape")
  ) %>% 
  # second select to arrange tree_metric
  dplyr::select(
    dataset_factor, tidyselect::starts_with("flight_"), tree_dbh_cm_bin
    # , site_area_m2
    # detection
    , tp_n
    # quantification
    # , c(tidyselect::contains("volume") & !tidyselect::contains("paraboloid"))
    , tidyselect::contains("area")
    , tidyselect::contains("height")
    , tidyselect::contains("diameter")
    , tidyselect::contains("dbh")
  ) %>% 
  # names()
  dplyr::mutate(
    dplyr::across(
      .cols = c(
        tidyselect::ends_with("_mape")
      )
      , .fn = ~ scales::percent(.x, accuracy = 1)
    )
    , dplyr::across(
      .cols = c(tidyselect::ends_with("_mean"))
      , .fn = ~ scales::comma(.x, accuracy = 0.01)
    )
    , dplyr::across(
      .cols = c(tidyselect::ends_with("_rmse"))
      , .fn = ~ scales::comma(.x, accuracy = 0.01)
    )
  ) %>% 
  dplyr::select(
    !tidyselect::contains("_area_")
    & !tidyselect::contains("diff_diameter_")
    & !tidyselect::ends_with("_trees")
    # & !tidyselect::ends_with("_n")
    & !tidyselect::ends_with("_rate")
  ) %>% 
  dplyr::arrange(dataset_factor, tree_dbh_cm_bin) %>% 
  dplyr::select(
    -c(
      dataset_factor
    )
  ) %>% 
  dplyr::mutate(
    dplyr::across(
      tidyselect::starts_with("flight_")
      , as.factor
    )
  ) %>% 
  # dplyr::glimpse()
  # ncol()
  kableExtra::kbl(
    caption = "Quantification Accuracy by Tree Size"
    , col.names = c(
      "altitude (m)", "speed (m sec<sup>-1</sup>)"
      , "DBH bin"
      , "TP predictions"
      , rep(c("ME","RMSE","MAPE"), times = 2)
    )
    , escape = F
    # , digits = 2
  ) %>% 
  kableExtra::kable_styling(font_size = 11.5) %>% 
  kableExtra::add_header_above(c(
    " "=4
    # , "Detection" = 3
    # , "Area" = 3
    , "Height (m)" = 3
    , "DBH (cm)" = 3
  )) %>% 
  kableExtra::column_spec(seq(4,10,by=3), border_right = TRUE, include_thead = TRUE) %>% 
  kableExtra::collapse_rows(columns = c(1:2), valign = "top")
```

that is interesting

let's look into the DBH prediction error over the range of tree sizes and use LOESS smoothing to highlight the trend in relative prediction error

```{r, fig.height = 10, fig.width = 8.3}
# Data trends (dashed) are visualized using a LOESS (locally weighted scatterplot smoothing) regression. 
# A span of 1.0 was utilized to include 100% of the observations in each local neighborhood, 
# and the local polynomial degree was restricted to 1 (local linear regression) to prevent overfitting and boundary inflation at the data extremes.

# The dashed line represents a locally estimated scatterplot smoothing (LOESS) trend line
# , calculated with a smoothing parameter (span) of 1.0 and a local linear polynomial fit (degree = 1).
p_temp <- 
  ground_truth_prediction_match_ans %>%
  # dplyr::mutate(pct_diff_dbh_cm = (pred_dbh_cm-ref_dbh_cm)/ref_dbh_cm) %>% 
  purrr::list_rbind(names_to = "dataset_factor") %>%
  dplyr::mutate(
    dataset_factor = ordered(dataset_factor, levels(tracking_df$dataset_factor))
    # swap the error calc
    , pct_diff_dbh_cm = (pred_dbh_cm-ref_dbh_cm)/ref_dbh_cm
  ) %>%
  sf::st_drop_geometry() %>%
  dplyr::filter(match_grp %in% c("true positive")) %>% 
  dataset_factor_relabel_fn() %>% 
  # dplyr::glimpse()
  ggplot2::ggplot(
    mapping = ggplot2::aes(y = pct_diff_dbh_cm, x = ref_dbh_cm)
  ) +
  ggplot2::geom_hline(yintercept = 0, color = "gray33", lwd = 1) +
  ggplot2::geom_point(color = "maroon4", alpha = 0.7, size = 0.7) +
  ggplot2::geom_smooth(
    method = "loess", se=F
    , span = 1 # 0.75 = default
    , method.args = list(degree = 1)
    , lwd = 1, color = "gray", linetype = "dashed"
  ) +
  ggplot2::annotate(
    "text"
    , x = Inf
    # , y = 0.045
    , y = Inf
    , label = "prediction > reference"
    , hjust = 1.1
    , vjust = 1.2
    , parse = F
    , size = 2.5
  ) +
  ggplot2::annotate(
    "text"
    , x = Inf
    # , y = -0.04
    , y = -Inf
    , label = "prediction < reference"
    , hjust = 1.1
    , vjust = -0.4
    , parse = F
    , size = 2.5
  ) +
  ggplot2::facet_wrap(
    facets = dplyr::vars(dataset_factor)
    # , rows = dplyr::vars(trtmnt_block)
    , ncol = 2
    , axes = "all"
    # , switch = "y"
  ) +
  ggplot2::scale_x_continuous(breaks = scales::breaks_extended(n=8)) +
  ggplot2::scale_y_continuous(
    limits = c(-1,1)
    # breaks = seq(-1,1,by=0.2)
    , breaks = scales::breaks_extended(n = 8)
    , labels = scales::percent
  ) +
  ggplot2::labs(
    x = "field reference DBH (cm)", y = "relative DBH error"
    , subtitle = "relative DBH error (%) by reference tree DBH"
  ) +
  ggplot2::theme_light()+
  ggplot2::theme(
    legend.position = "top"
    , strip.text = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , axis.text = ggplot2::element_text(size = 6)
  )
p_temp
ggplot2::ggsave("../data/dbh_acc_by_size.jpg",p_temp+ggplot2::theme(plot.subtitle = ggplot2::element_blank()), height = 10, width = 8.3)
```

that's neat

do it for height too

```{r, fig.height = 10, fig.width = 8.3}
# Data trends (dashed) are visualized using a LOESS (locally weighted scatterplot smoothing) regression. 
# A span of 1.0 was utilized to include 100% of the observations in each local neighborhood, 
# and the local polynomial degree was restricted to 1 (local linear regression) to prevent overfitting and boundary inflation at the data extremes.

# The dashed line represents a locally estimated scatterplot smoothing (LOESS) trend line
# , calculated with a smoothing parameter (span) of 1.0 and a local linear polynomial fit (degree = 1).
p_temp <- 
  ground_truth_prediction_match_ans %>%
  purrr::list_rbind(names_to = "dataset_factor") %>%
  dplyr::mutate(
    dataset_factor = ordered(dataset_factor, levels(tracking_df$dataset_factor))
    # swap the error calc
    , pct_diff_dbh_cm = (pred_tree_height_m-ref_tree_height_m)/ref_tree_height_m
  ) %>%
  sf::st_drop_geometry() %>%
  dplyr::filter(match_grp %in% c("true positive")) %>% 
  dataset_factor_relabel_fn() %>% 
  # dplyr::glimpse()
  ggplot2::ggplot(
    mapping = ggplot2::aes(y = pct_diff_tree_height_m, x = ref_tree_height_m)
  ) +
  ggplot2::geom_hline(yintercept = 0, color = "gray33", lwd = 1) +
  ggplot2::geom_point(color = "maroon4", alpha = 0.7, size = 0.7) +
  ggplot2::geom_smooth(
    method = "loess", se=F
    , span = 1 # 0.75 = default
    , method.args = list(degree = 1)
    , lwd = 1, color = "gray", linetype = "dashed"
  ) +
  ggplot2::annotate(
    "text"
    , x = Inf
    # , y = 0.045
    , y = Inf
    , label = "prediction > reference"
    , hjust = 1.1
    , vjust = 1.2
    , parse = F
    , size = 2.5
  ) +
  ggplot2::annotate(
    "text"
    , x = Inf
    # , y = -0.04
    , y = -Inf
    , label = "prediction < reference"
    , hjust = 1.1
    , vjust = -0.4
    , parse = F
    , size = 2.5
  ) +
  ggplot2::facet_wrap(
    facets = dplyr::vars(dataset_factor)
    # , rows = dplyr::vars(trtmnt_block)
    , ncol = 2
    , axes = "all"
    # , switch = "y"
  ) +
  ggplot2::scale_x_continuous(breaks = scales::breaks_extended(n=8)) +
  ggplot2::scale_y_continuous(
    limits = c(-1,1)
    # breaks = seq(-1,1,by=0.2)
    , breaks = scales::breaks_extended(n = 8)
    , labels = scales::percent
  ) +
  ggplot2::labs(
    x = "field reference height (m)", y = "relative height error"
    , subtitle = "relative tree height error (%) by reference tree height"
  ) +
  ggplot2::theme_light()+
  ggplot2::theme(
    legend.position = "top"
    , strip.text = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , axis.text = ggplot2::element_text(size = 6)
  )
p_temp
ggplot2::ggsave("../data/ht_acc_by_size.jpg",p_temp+ggplot2::theme(plot.subtitle = ggplot2::element_blank()), height = 10, width = 8.3)
```

that is also neat

## Stand-Level Accuracy Results{#stand_acc}

In addition to the tree-level accuracy metrics (e.g. height RMSE and MAPE), we can also calculate stand-level accuracy metrics by aggregating estimates from the reference (ground truth) and predicted trees irrespective of whether or not they were matched (i.e. true positive) during instance matching. 

We'll evaluate stand level aggregation accuracy for stand basal area and trees per hectare. For these stand metrics, we calculated the difference ($predicted - reference$) to measure the bias in the original units (e.g., $\frac{m^{2}}{ha}$ for basal area) which indicates the over- or under-estimation of the stand structure. We also calculated the percent difference ($\frac{predicted-reference}{reference}$) to measure the relative bias (scale-independent) to capture the magnitude of the error.

let's table the stand-level measurement error

```{r}
agg_ground_truth_match_ans %>% 
  dplyr::select(
    dataset_factor, tidyselect::starts_with("flight_")
    # , site_area_m2
    # detection
    , ref_n_trees
    , pred_n_trees
    # quantification
    , tidyselect::ends_with("_basal_area_m2_per_ha")
    , tidyselect::ends_with("_trees_per_ha")
  ) %>% 
  dplyr::select(!tidyselect::starts_with("abs_diff_")) %>% 
  dplyr::mutate(
    dplyr::across(
      tidyselect::starts_with("flight_")
      , as.factor
    )
    , dplyr::across(
      .cols = c(
        tidyselect::starts_with("pct_diff_")
        , tidyselect::ends_with("_rate")
      )
      , .fn = ~ scales::percent(.x, accuracy = .1)
    )
    , dplyr::across(
      .cols = c(tidyselect::ends_with("_trees"))
      , .fn = ~ scales::comma(.x, accuracy = 1)
    )
    , dplyr::across(
      .cols = dplyr::where(is.numeric)
      , .fn = ~ scales::comma(.x, accuracy = 0.1)
    )
    # # one column for diff/pct
    , diff_basal_area_m2_per_ha = paste0(diff_basal_area_m2_per_ha, "<br>(", pct_diff_basal_area_m2_per_ha, ")")
    , diff_trees_per_ha = paste0(diff_trees_per_ha, "<br>(", pct_diff_trees_per_ha, ")")
  ) %>% 
  dplyr::select(
    !tidyselect::starts_with("pct_diff_")
  ) %>% 
  dplyr::arrange(dataset_factor) %>% 
  dplyr::select(
    -c(
      dataset_factor
    )
  ) %>% 
  # dplyr::glimpse()
  
  kableExtra::kbl(
    caption = "Stand-Level Accuracy"
    , col.names = c(
      "altitude (m)", "speed (m sec<sup>-1</sup>)"
      , "reference", "predicted"
      , rep(c("reference","predicted","difference"), times = 2)
    )
    , escape = F
    # , digits = 2
  ) %>% 
  kableExtra::kable_styling(font_size = 11.5) %>% 
  kableExtra::add_header_above(
    c(
      " "=2
      , "# Trees" = 2
      # , "Area" = 3
      , "Basal Area<br />m<sup>2</sup> ha<sup>-1</sup>" = 3
      , "Trees<br />ha<sup>-1</sup>" = 3
    )
    , escape = F
  ) %>% 
  kableExtra::column_spec( c(2,seq(4,10,by=3) ), border_right = TRUE, include_thead = TRUE) %>% 
  kableExtra::collapse_rows(columns = c(1), valign = "top")
```

let's plot the percent difference in BA and TPH by flight

```{r, fig.height=8}
agg_ground_truth_match_ans %>% 
  dplyr::select(
    dataset_factor
    , pct_diff_basal_area_m2_per_ha
    , pct_diff_trees_per_ha
  ) %>% 
  dplyr::mutate(
    tot = -1* (abs(pct_diff_basal_area_m2_per_ha) + abs(pct_diff_trees_per_ha))
  ) %>% 
  tidyr::pivot_longer(cols = -c(dataset_factor,tot)) %>% 
  dplyr::mutate(
    name = dplyr::recode_values(
      name
      , "pct_diff_basal_area_m2_per_ha" ~ latex2exp::TeX("Basal Area $m^{2} \\cdot ha^{-1}$", output = "character")
      , "pct_diff_trees_per_ha" ~ latex2exp::TeX("Trees $ha^{-1}$", output = "character")
    )
    , value_lab = scales::percent(value,accuracy=0.1)
  ) %>% 
  dataset_factor_relabel_fn() %>% 
ggplot2::ggplot(
  mapping = ggplot2::aes(
    y = reorder(dataset_factor, tot)
    , x = value
    , group = name
  )
) +
  ggplot2::geom_vline(xintercept = 0, color = "gray95", lwd = 1.5, alpha = 0.7) +
  ggplot2::geom_vline(xintercept = c(0.05,-0.05), color = "gray77", lwd = 0.5, alpha = 0.7) +
  ggplot2::geom_vline(xintercept = c(0.1,-0.1), color = "gray55", lwd = 0.5, alpha = 0.7) +
  ggplot2::geom_vline(xintercept = c(0.15,-0.15), color = "gray33", lwd = 0.5, alpha = 0.7) +
  ggplot2::geom_vline(xintercept = c(0.2,-0.2), color = "gray11", lwd = 0.5) +
  ggplot2::geom_segment(
    mapping = ggplot2::aes(yend = dataset_factor, xend = 0,color = value > 0)
    # , color = "gray44"
  ) +
  ggplot2::geom_point(
    mapping = ggplot2::aes(color = value > 0)
    , size = 3
  ) +
  # lhs labs
  ggplot2::geom_text(
    data = function(x){dplyr::filter(x,value<0)}
    , mapping = ggplot2::aes(label = value_lab, fontface = "bold")
    # , color = "black"
    , size = 3
    , nudge_x = -0.006
    , hjust = 1
  ) +
  # rhs labs
  ggplot2::geom_text(
    data = function(x){dplyr::filter(x,value>=0)}
    , mapping = ggplot2::aes(label = value_lab, fontface = "bold")
    # , color = "black"
    , size = 3
    , nudge_x = 0.006
    , hjust = 0
  ) +
  ggplot2::facet_grid(
    cols = dplyr::vars(name)
    # , labeller = ggplot2::label_parsed
    , labeller = ggplot2::as_labeller(
      function(x) {
        paste0("atop(", x, ", 'percent error')")
      }
      , default = ggplot2::label_parsed
    )
  ) +
  ggplot2::scale_color_manual(values = c("firebrick","navy")) +
  ggplot2::scale_x_continuous(
    limits = c(-0.2,0.2)
    , labels = scales::percent
    # , breaks = scales::breaks_extended(n=10)
    # , breaks = function(x) seq(min(x), max(x), by = 0.05)
  ) +
  ggplot2::labs(
    y = "" # "comparison"
    , x = "percent error"
    , subtitle = "positive = UAS-measured higher than field-measured\nnegative = UAS-measured lower than field-measured"
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "none"
    , axis.text.y = ggplot2::element_text(size = 9, face = "bold")
    # , plot.subtitle = ggplot2::element_text(hjust = 1)
    , strip.text = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , panel.grid.major.x = element_blank()
    , panel.grid.minor.x = element_blank()
  )


```

let's take a slightly different look to explore the UAS bias in the original units for basal area which indicates the over- or under-estimation

```{r, fig.height=8}
agg_ground_truth_match_ans %>% 
  dplyr::mutate(
    dataset_factor = forcats::fct_rev(dataset_factor)
    , lab = paste0(
      ifelse(diff_basal_area_m2_per_ha>=0,"+","")
      , scales::comma(diff_basal_area_m2_per_ha,accuracy=0.1)
      , "\n("
      , scales::percent(pct_diff_basal_area_m2_per_ha,accuracy=0.1)
      , ")"
    )
  ) %>% 
  dataset_factor_relabel_fn() %>% 
  # dplyr::mutate(
  #   name = dplyr::recode_values(
  #     name
  #     , "diff_basal_area_m2_per_ha" ~ latex2exp::TeX("Basal Area $m^{2} \\cdot ha^{-1}$", output = "character")
  #     , "diff_trees_per_ha" ~ latex2exp::TeX("Trees $ha^{-1}$", output = "character")
  #   )
  #   , value_lab = scales::comma(value,accuracy=0.1)
  # ) %>% 
ggplot2::ggplot(
  mapping = ggplot2::aes(
    y = dataset_factor
  )
) +
  # ref
  ggplot2::geom_vline(
    mapping = ggplot2::aes(
      xintercept = ref_basal_area_m2_per_ha
      , color = "field"
    )
    , lwd = 1.5, alpha = 0.15
  ) +
  ggplot2::geom_segment(
    mapping = ggplot2::aes(
      x = ref_basal_area_m2_per_ha
      , xend = pred_basal_area_m2_per_ha
      , yend = dataset_factor
    )
    , color = "gray70", lwd = 1.5
  ) +
  # ref
  ggplot2::geom_point(
    mapping = ggplot2::aes(
      x = ref_basal_area_m2_per_ha
      , color = "field"
    )
    , size = 4
  ) +
  # pred
  ggplot2::geom_point(
    mapping = ggplot2::aes(
      x = pred_basal_area_m2_per_ha
      , color = "prediction"
    )
    , size = 4
  ) +
  # ref labs
  ggplot2::geom_text(
    mapping = ggplot2::aes(x = ref_basal_area_m2_per_ha, label = scales::comma(ref_basal_area_m2_per_ha, accuracy = 0.1))
    , size = 2.5
    , hjust = 0.5, vjust = 2.2
  ) +
  # pred labs
  ggplot2::geom_text(
    mapping = ggplot2::aes(x = pred_basal_area_m2_per_ha, label = scales::comma(pred_basal_area_m2_per_ha, accuracy = 0.1))
    , size = 2.5
    , hjust = 0.5, vjust = 2.2
  ) +
  # diff labs
  ggplot2::geom_text(
    mapping = ggplot2::aes(
      x = (ref_basal_area_m2_per_ha+pred_basal_area_m2_per_ha)/2
      , label = lab
    )
    , size = 2.1
    , hjust = 0.5, vjust = -0.5
    , color = "gray22"
  ) +
  harrypotter::scale_color_hp_d(option = "hermionegranger") +
  ggplot2::scale_x_continuous(
    limits = c(0,NA)
    , labels = scales::comma
    , breaks = scales::breaks_extended(n=10)
    , expand = ggplot2::expansion(mult = c(0.02,0.2))
    , sec.axis = ggplot2::dup_axis()
  ) +
  ggplot2::labs(
    y = "" # "comparison"
    , x = latex2exp::TeX("Basal Area $m^{2} \\cdot ha^{-1}$")
    , color = ""
    , subtitle = latex2exp::TeX("Predicted vs Reference Stand Basal Area ($m^{2} \\cdot ha^{-1}$)")
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "top"
    # , plot.subtitle = ggplot2::element_text(hjust = 1)
    , axis.text.y = ggplot2::element_text(size = 9, face = "bold")
    , axis.text.x = ggplot2::element_text(size = 7)
    , axis.title.x = ggplot2::element_text(size = 7)
    , panel.grid.major.x = element_blank()
    , panel.grid.minor.x = element_blank()
  ) +
  ggplot2::guides(
    color = ggplot2::guide_legend(
      override.aes = list(shape = 15, size = 6, linetype = 0)
    )
  )

```

above we saw that the UAS generally overpredicted the DBH and the figure above shows that this bias propagated through to the stand-level basal area. as a result, basal area was overestimated across all flights with overestimations ranging from `r paste(scales::percent(fivenum(abs(agg_ground_truth_match_ans$pct_diff_basal_area_m2_per_ha))[c(1,5)],accuracy=.1), collapse = " to ")` across the tested flights

now we'll look at the UAS bias in the original units for trees per hectare

```{r, fig.height=8}
agg_ground_truth_match_ans %>% 
  dplyr::mutate(
    dataset_factor = forcats::fct_rev(dataset_factor)
    , lab = paste0(
      ifelse(diff_trees_per_ha>=0,"+","")
      , scales::comma(diff_trees_per_ha,accuracy=0.1)
      , "\n("
      , scales::percent(pct_diff_trees_per_ha,accuracy=0.1)
      , ")"
    )
  ) %>% 
  dataset_factor_relabel_fn() %>% 
  # dplyr::mutate(
  #   name = dplyr::recode_values(
  #     name
  #     , "diff_trees_per_ha" ~ latex2exp::TeX("Basal Area $m^{2} \\cdot ha^{-1}$", output = "character")
  #     , "diff_trees_per_ha" ~ latex2exp::TeX("Trees $ha^{-1}$", output = "character")
  #   )
  #   , value_lab = scales::comma(value,accuracy=0.1)
  # ) %>% 
ggplot2::ggplot(
  mapping = ggplot2::aes(
    y = dataset_factor
  )
) +
  # ref
  ggplot2::geom_vline(
    mapping = ggplot2::aes(
      xintercept = ref_trees_per_ha
      , color = "field"
    )
    , lwd = 1.5, alpha = 0.15
  ) +
  ggplot2::geom_segment(
    mapping = ggplot2::aes(
      x = ref_trees_per_ha
      , xend = pred_trees_per_ha
      , yend = dataset_factor
    )
    , color = "gray70", lwd = 1.5
  ) +
  # ref
  ggplot2::geom_point(
    mapping = ggplot2::aes(
      x = ref_trees_per_ha
      , color = "field"
    )
    , size = 4
  ) +
  # pred
  ggplot2::geom_point(
    mapping = ggplot2::aes(
      x = pred_trees_per_ha
      , color = "prediction"
    )
    , size = 4
  ) +
  # ref labs
  ggplot2::geom_text(
    mapping = ggplot2::aes(x = ref_trees_per_ha, label = scales::comma(ref_trees_per_ha, accuracy = 0.1))
    , size = 2.5
    , hjust = 0.5, vjust = 2.2
  ) +
  # pred labs
  ggplot2::geom_text(
    mapping = ggplot2::aes(
      x = pred_trees_per_ha
      , label = ifelse(
        abs(pct_diff_trees_per_ha)<0.01
        , ""
        , scales::comma(pred_trees_per_ha, accuracy = 0.1)  
      )
    )
    , size = 2.5
    , hjust = 0.5, vjust = 2.2
  ) +
  # diff labs
  ggplot2::geom_text(
    mapping = ggplot2::aes(
      x = (ref_trees_per_ha+pred_trees_per_ha)/2
      , label = lab
    )
    , size = 2.1
    , hjust = 0.5, vjust = -0.5
    , color = "gray22"
  ) +
  harrypotter::scale_color_hp_d(option = "hermionegranger") +
  ggplot2::scale_x_continuous(
    labels = scales::comma
    , breaks = scales::breaks_extended(n=10)
    , expand = ggplot2::expansion(mult = c(0.5,0.2))
    , sec.axis = ggplot2::dup_axis()
  ) +
  ggplot2::labs(
    y = "" # "comparison"
    , x = latex2exp::TeX("Trees $ha^{-1}$")
    , color = ""
    , subtitle = latex2exp::TeX("Predicted vs Reference Stand Trees $ha^{-1}$")
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "top"
    # , plot.subtitle = ggplot2::element_text(hjust = 1)
    , axis.text.y = ggplot2::element_text(size = 9, face = "bold")
    , axis.text.x = ggplot2::element_text(size = 7)
    , axis.title.x = ggplot2::element_text(size = 7)
    , panel.grid.major.x = element_blank()
    , panel.grid.minor.x = element_blank()
  ) +
  ggplot2::guides(
    color = ggplot2::guide_legend(
      override.aes = list(shape = 15, size = 6, linetype = 0)
    )
  )
```

stand-level predictions for trees per hectare vary by flight setting but all flights resulted in TPH estimates near the field-measured data. the largest TPH error was a `r scales::percent(max(abs(agg_ground_truth_match_ans$pct_diff_trees_per_ha)),accuracy=.1)` underestimation while the best performing flight resulted in a `r scales::percent(min(abs(agg_ground_truth_match_ans$pct_diff_trees_per_ha)),accuracy=.1)` error (i.e. nearly identical to the reference value). The remaining flights achieved TPH error rates between `r paste(scales::percent(quantile(abs(agg_ground_truth_match_ans$pct_diff_trees_per_ha),c(.2,.8)),accuracy=.1), collapse = " and ")`.

the plots maybe don't help us see the impact of different flight settings, let's attempt that all on a single plot

**basal area**

```{r, fig.height=8.5}
agg_ground_truth_match_ans %>% 
  dplyr::select(
    dataset_factor, tidyselect::starts_with("flight_")
    # , diff_basal_area_m2_per_ha
    , tidyselect::ends_with("_basal_area_m2_per_ha")
  ) %>% 
  dplyr::select(!tidyselect::starts_with("abs_diff_")) %>% 
  tidyr::pivot_longer(cols = c(ref_basal_area_m2_per_ha, pred_basal_area_m2_per_ha)) %>% 
  dplyr::mutate(
    name = dplyr::recode_values(
      name
      , "ref_basal_area_m2_per_ha" ~ "field"
      , "pred_basal_area_m2_per_ha" ~ "prediction"
    )
    , value_lab = scales::comma(value,accuracy=0.1)
  ) %>% 
  dataset_factor_relabel_fn() %>% 
  ggplot2::ggplot(
    mapping = ggplot2::aes(x = name, y = value, group = dataset_factor)
  ) +
  ggplot2::geom_line(key_glyph = "point", alpha = 0.7, color = "gray", lwd = 1.1) +
  # ref labs
  ggplot2::geom_text(
    data = function(x){dplyr::filter(x,name=="field")}
    , mapping = ggplot2::aes(label = value_lab)
    , size = 3
    , hjust = 0.5, vjust = 2.2
  ) +
  # pred labs
  ggrepel::geom_text_repel(
    data = function(x){dplyr::filter(x,name=="prediction")}
    , mapping = ggplot2::aes(label = value_lab)
    , size = 2.3
    , hjust = 0.5, vjust = 2.2
  ) +
  ggplot2::geom_point(
    mapping = ggplot2::aes(color = name)
    # , alpha = 0.7
    , size = 4
  ) +
  ggplot2::facet_grid(
    cols = dplyr::vars(flight_altitude_m)
    , rows = dplyr::vars(flight_meters_per_second)
    , labeller = ggplot2::labeller( 
      flight_altitude_m = ~paste("Altitude:", .x, " (m)") 
      , flight_meters_per_second = ~paste("Speed:", .x, " (m/s)") 
    )
    , axes = "all_x"
  ) +
  harrypotter::scale_color_hp_d(option = "hermionegranger") +
  ggplot2::scale_y_continuous(
    limits = c(0,NA)
    , labels = scales::comma
    , breaks = scales::breaks_extended(n=10)
    , expand = ggplot2::expansion(mult = c(0.01,0.1))
  ) +
  ggplot2::labs(
    x = "" # "comparison"
    , y = latex2exp::TeX("Basal Area $m^{2} \\cdot ha^{-1}$")
    , color = ""
    , subtitle = latex2exp::TeX("Predicted vs Reference Stand Basal Area ($m^{2} \\cdot ha^{-1}$)")
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "top"
    , axis.text.x = ggplot2::element_text(size = 10, face = "bold", color = "black")
    , axis.text.y = ggplot2::element_text(size = 7)
    , axis.title.y = ggplot2::element_text(size = 7)
    , strip.text = ggplot2::element_text(size = 10, color = "black", face = "bold")
  ) +
  ggplot2::guides(
    color = ggplot2::guide_legend(
      override.aes = list(shape = 15, size = 6, linetype = 0)
    )
  )

```

**trees per ha**

```{r, fig.height=8.5}
agg_ground_truth_match_ans %>% 
  dplyr::select(
    dataset_factor, tidyselect::starts_with("flight_")
    # , diff_basal_area_m2_per_ha
    , tidyselect::ends_with("_trees_per_ha")
  ) %>% 
  dplyr::select(!tidyselect::starts_with("abs_diff_")) %>% 
  tidyr::pivot_longer(cols = c(ref_trees_per_ha, pred_trees_per_ha)) %>% 
  dplyr::mutate(
    name = dplyr::recode_values(
      name
      , "ref_trees_per_ha" ~ "field"
      , "pred_trees_per_ha" ~ "prediction"
    )
    , value_lab = scales::comma(value,accuracy=0.1)
  ) %>% 
  dataset_factor_relabel_fn() %>% 
  ggplot2::ggplot(
    mapping = ggplot2::aes(x = name, y = value, group = dataset_factor)
  ) +
  ggplot2::geom_line(key_glyph = "point", alpha = 0.7, color = "gray", lwd = 1.1) +
  # ref labs
  ggplot2::geom_text(
    data = function(x){dplyr::filter(x,name=="field")}
    , mapping = ggplot2::aes(label = value_lab)
    , size = 3
    , hjust = 0.5, vjust = 2.2
  ) +
  # pred labs
  ggrepel::geom_text_repel(
    data = function(x){dplyr::filter(x,name=="prediction")}
    , mapping = ggplot2::aes(label = value_lab)
    , size = 2.3
    , hjust = 0.5, vjust = 2.2
  ) +
  ggplot2::geom_point(
    mapping = ggplot2::aes(color = name)
    # , alpha = 0.7
    , size = 4
  ) +
  ggplot2::facet_grid(
    cols = dplyr::vars(flight_altitude_m)
    , rows = dplyr::vars(flight_meters_per_second)
    , labeller = ggplot2::labeller( 
      flight_altitude_m = ~paste("Altitude:", .x, " (m)") 
      , flight_meters_per_second = ~paste("Speed:", .x, " (m/s)") 
    )
    , axes = "all_x"
  ) +
  harrypotter::scale_color_hp_d(option = "hermionegranger") +
  ggplot2::scale_y_continuous(
    limits = c(0,NA)
    , labels = scales::comma
    , breaks = scales::breaks_extended(n=10)
    , expand = ggplot2::expansion(mult = c(0.01,0.1))
  ) +
  ggplot2::labs(
    x = "" # "comparison"
    , y = latex2exp::TeX("Trees $ha^{-1}$")
    , color = ""
    , subtitle = latex2exp::TeX("Predicted vs Reference Stand Trees $ha^{-1}$")
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "top"
    , axis.text.x = ggplot2::element_text(size = 10, face = "bold", color = "black")
    , axis.text.y = ggplot2::element_text(size = 7)
    , axis.title.y = ggplot2::element_text(size = 7)
    , strip.text = ggplot2::element_text(size = 10, color = "black", face = "bold")
  ) +
  ggplot2::guides(
    color = ggplot2::guide_legend(
      override.aes = list(shape = 15, size = 6, linetype = 0)
    )
  )

```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Statistical Analysis

### Aggregated Stand-Level

We'll quickly use a frequentist approach to test for differences in the detection (F-score) and quantification (e.g. DBH RMSE) accuracy at different levels of the flight setting variables. Due to the limited data, we'll test if the effect of speed changes with altitude (interaction) while controlling for terrain following (intercept) using multiple linear regression.

#### Tree Detection

influence of flight setting parameters on **F-score**

```{r}
# lm
lm_temp <- lm(
  # f_score ~ 0 + flight_altitude_m * flight_meters_per_second * flight_tf # full three and two-way interactions
  f_score ~ flight_altitude_m*flight_meters_per_second # metric interactions only
  , data = agg_ground_truth_match_ans
)
# summary(lm_temp)
broom::tidy(lm_temp) %>%
  dplyr::mutate(
    term = term %>% 
      stringr::str_replace_all("flight_tf","terrain follow") %>% 
      stringr::str_replace_all("flight_altitude_m","altitude") %>% 
      stringr::str_replace_all("flight_meters_per_second","speed") %>% 
      stringr::str_replace_all("FALSE"," = off") %>% 
      stringr::str_replace_all("TRUE"," = on")
  ) %>% 
  kableExtra::kbl(
    caption = "MLR: influence of flight setting parameters on F-score"
    , digits = 4
  ) %>%
  kableExtra::kable_styling()

```

I don't think we can make many inferences from this model other than "there is no impact (or no evidence of impact) of the various flight settings on overall tree detection accuracy."

```{r,include=FALSE, eval=F}
# verify if terrain follow is truly the only significant driver
stats::drop1(lm_temp, test = "F") %>%
  kableExtra::kbl(caption = "Marginal Significance Tests (drop1)") %>%
  kableExtra::kable_styling()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

#### Basal Area

influence of flight setting parameters on **basal area** based on the absolute difference (total error) from the field-measured data

```{r}
# lm
lm_temp <- lm(
  # f_score ~ 0 + flight_altitude_m * flight_meters_per_second * flight_tf # full three and two-way interactions
  abs_diff_basal_area_m2_per_ha ~ flight_altitude_m*flight_meters_per_second # metric interactions only
  , data = agg_ground_truth_match_ans
)
# summary(lm_temp)
broom::tidy(lm_temp) %>%
  dplyr::mutate(
    term = term %>% 
      stringr::str_replace_all("flight_tf","terrain follow") %>% 
      stringr::str_replace_all("flight_altitude_m","altitude") %>% 
      stringr::str_replace_all("flight_meters_per_second","speed") %>% 
      stringr::str_replace_all("FALSE"," = off") %>% 
      stringr::str_replace_all("TRUE"," = on")
  ) %>% 
  kableExtra::kbl(
    caption = "MLR: influence of flight setting parameters on basal area error"
    , digits = 4
  ) %>%
  kableExtra::kable_styling()

```

altitude alone is significant at the 0.05 level. this finding is evidence that the altitude flight setting influences basal area error in a meaningful way, but note that the estimated coefficient on the variable is very small.

let's make pairwise contrasts (`emmeans::contrast()`) based on the estimated marginal means (`emmeans::emmeans()`) over the range of possible flight settings for altitude and speed to test if the practical settings (i.e. we cannot fly at 1000 ft or 0.05 mph or 99 mph) are different. if any of these contrasts are statistically significant (i.e. there is not a zero difference in the basal area error), it will help us identify if there is a flight setting that we should use as an operational rule if our objective is to minimize basal area error at the stand level

```{r}
preds_temp <- emmeans::emmeans(
  lm_temp
  , ~ flight_altitude_m * flight_meters_per_second
  , at = list(
      flight_altitude_m = seq(
        floor( min(agg_ground_truth_match_ans$flight_altitude_m) )
        , floor( max(agg_ground_truth_match_ans$flight_altitude_m) )
        , by = round(100/3.281)
      )
      , flight_meters_per_second = seq(
        floor( min(agg_ground_truth_match_ans$flight_meters_per_second) )
        , floor( max(agg_ground_truth_match_ans$flight_meters_per_second) )
        , by = round(5/2.237,1)
      )
    )
)
# preds_temp
```

contrast of **altitude** levels at different flight speeds (see the table footnote for interpretation) for basal area per hectare error

```{r}
preds_temp %>% 
  emmeans::contrast(method = "pairwise", simple = "each", combine = TRUE) %>% 
  dplyr::as_tibble() %>%
  dplyr::filter(flight_meters_per_second!=".",!is.na(flight_meters_per_second)) %>% 
  dplyr::select(flight_meters_per_second, contrast,estimate,SE,t.ratio,p.value) %>%
  dplyr::arrange(flight_meters_per_second) %>% 
  dplyr::mutate(
    contrast = contrast %>% 
      # clean agl
      stringr::str_replace("flight_altitude_m", "xxx") %>% 
      stringr::str_remove_all("flight_altitude_m") %>% 
      stringr::str_replace("xxx", "altitude (m): ") %>% 
      # clean mph
      stringr::str_replace("flight_meters_per_second", "xxx") %>% 
      stringr::str_remove_all("flight_meters_per_second") %>% 
      stringr::str_replace("xxx", "speed (m sec<sup>-1</sup>): ")
      # rest
    , dplyr::across(
      c(estimate,SE,t.ratio,p.value)
      , ~scales::comma(.x, accuracy = 0.001)
    )
  ) %>% 
  kableExtra::kbl(
    caption = "pairwise tests for non-zero differences in basal area error (m2/ha)"
    , col.names = c(
      "speed (m sec<sup>-1</sup>)", "contrast"
      # , "speed (m sec<sup>-1</sup>)", "altitude (m)"
      , "estimate", "SE", "t.ratio", "p.value"
    )
    , digits = 3
    , escape = F
  ) %>%
  kableExtra::kable_styling() %>% 
  kableExtra::column_spec(c(2), border_right = TRUE, include_thead = TRUE) %>% 
  kableExtra::collapse_rows(columns = c(1), valign = "top") %>% 
  kableExtra::footnote(
    general_title = "interpretation:"
    , general = c(
      "`estimate`: difference in error; positive means the first setting has higher error."
      , "`SE`: measures the precision of the estimate (smaller values indicate higher precision)"
      , "`t.ratio`: The test statistic; indicates distance from the null (zero difference)"
      , "`p.value`: probability of observing this result by chance; values < 0.05 are considered significant"
    )
    , footnote_as_chunk = FALSE
    , escape = F
  )
```

contrast of **speed** levels at different flight speeds (see the table footnote for interpretation) for basal area per hectare error

```{r}
preds_temp %>% 
  emmeans::contrast(method = "pairwise", simple = "each", combine = TRUE) %>% 
  dplyr::as_tibble() %>%
  dplyr::filter(flight_altitude_m!=".",!is.na(flight_altitude_m)) %>% 
  dplyr::select(flight_altitude_m, contrast,estimate,SE,t.ratio,p.value) %>%
  dplyr::arrange(flight_altitude_m) %>% 
  dplyr::mutate(
    contrast = contrast %>% 
      # clean agl
      stringr::str_replace("flight_altitude_m", "xxx") %>% 
      stringr::str_remove_all("flight_altitude_m") %>% 
      stringr::str_replace("xxx", "altitude (m): ") %>% 
      # clean mph
      stringr::str_replace("flight_meters_per_second", "xxx") %>% 
      stringr::str_remove_all("flight_meters_per_second") %>% 
      stringr::str_replace("xxx", "speed (m sec<sup>-1</sup>): ")
      # rest
    , dplyr::across(
      c(estimate,SE,t.ratio,p.value)
      , ~scales::comma(.x, accuracy = 0.001)
    )
  ) %>% 
  kableExtra::kbl(
    caption = "pairwise tests for non-zero differences in basal area error (m2/ha)"
    , col.names = c(
      "altitude (m)", "contrast"
      # , "speed (m sec<sup>-1</sup>)", "altitude (m)"
      , "estimate", "SE", "t.ratio", "p.value"
    )
    , digits = 3
    , escape = F
  ) %>%
  kableExtra::kable_styling() %>% 
  kableExtra::column_spec(c(2), border_right = TRUE, include_thead = TRUE) %>% 
  kableExtra::collapse_rows(columns = c(1), valign = "top") %>% 
  kableExtra::footnote(
    general_title = "interpretation:"
    , general = c(
      "`estimate`: difference in error; positive means the first setting has higher error."
      , "`SE`: measures the precision of the estimate (smaller values indicate higher precision)"
      , "`t.ratio`: The test statistic; indicates distance from the null (zero difference)"
      , "`p.value`: probability of observing this result by chance; values < 0.05 are considered significant"
    )
    , footnote_as_chunk = FALSE
    , escape = F
  )
```

the significant linear trends from the MLR indicate that altitude influences basal area error, but the lack of significant pairwise contrasts suggests these settings are practically equivalent within the operationally feasible range of these flight settings. this result implies that the UAS-lidar and point cloud processing framework is not meaningfully influenced by changes in the flight settings over the range tested here. additionally, given the small sample size with no cross-site replication these findings should be viewed with caution as the lack of significance may reflect low statistical power rather than absence of effect.

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

#### Trees per hectare

influence of flight setting parameters on **trees per hectare** based on the absolute difference (total error) from the field-measured data

```{r}
# agg_ground_truth_match_ans$diff_trees_per_ha
# lm
lm_temp <- lm(
  # f_score ~ 0 + flight_altitude_m * flight_meters_per_second * flight_tf # full three and two-way interactions
  abs_diff_trees_per_ha ~ flight_altitude_m*flight_meters_per_second # metric interactions only
  , data = agg_ground_truth_match_ans
)
# summary(lm_temp)
broom::tidy(lm_temp) %>%
  dplyr::mutate(
    term = term %>% 
      stringr::str_replace_all("flight_tf","terrain follow") %>% 
      stringr::str_replace_all("flight_altitude_m","altitude") %>% 
      stringr::str_replace_all("flight_meters_per_second","speed") %>% 
      stringr::str_replace_all("FALSE"," = off") %>% 
      stringr::str_replace_all("TRUE"," = on")
  ) %>% 
  kableExtra::kbl(
    caption = "MLR: influence of flight setting parameters on trees per hectare error"
    , digits = 4
  ) %>%
  kableExtra::kable_styling()

```

speed alone is significant at the 0.1 level. this finding is weak evidence that the flight setting of speed influences tree density prediction error, but note that the estimated coefficient on those variables is small relative to the TPH of the site.

let's make pairwise contrasts (`emmeans::contrast()`) based on the estimated marginal means (`emmeans::emmeans()`) over the range of possible flight settings for altitude and speed to test if the practical settings (i.e. we cannot fly at 1000 ft or 0.05 mph or 99 mph) are different. if any of these contrasts are statistically significant (i.e. there is not a zero difference in the trees per hectare error), it will help us identify if there is a flight setting that we should use as an operational rule if our objective is to minimize trees per hectare error at the stand level

```{r}
preds_temp <- emmeans::emmeans(
  lm_temp
  , ~ flight_altitude_m * flight_meters_per_second
  , at = list(
      flight_altitude_m = seq(
        floor( min(agg_ground_truth_match_ans$flight_altitude_m) )
        , floor( max(agg_ground_truth_match_ans$flight_altitude_m) )
        , by = round(100/3.281)
      )
      , flight_meters_per_second = seq(
        floor( min(agg_ground_truth_match_ans$flight_meters_per_second) )
        , floor( max(agg_ground_truth_match_ans$flight_meters_per_second) )
        , by = round(5/2.237,1)
      )
    )
)
```

contrast of **altitude** levels at different flight speeds (see the table footnote for interpretation) for trees per hectare per hectare error

```{r}
preds_temp %>% 
  emmeans::contrast(method = "pairwise", simple = "each", combine = TRUE) %>% 
  dplyr::as_tibble() %>%
  dplyr::filter(flight_meters_per_second!=".",!is.na(flight_meters_per_second)) %>% 
  dplyr::select(flight_meters_per_second, contrast,estimate,SE,t.ratio,p.value) %>%
  dplyr::arrange(flight_meters_per_second) %>% 
  dplyr::mutate(
    contrast = contrast %>% 
      # clean agl
      stringr::str_replace("flight_altitude_m", "xxx") %>% 
      stringr::str_remove_all("flight_altitude_m") %>% 
      stringr::str_replace("xxx", "altitude (m): ") %>% 
      # clean mph
      stringr::str_replace("flight_meters_per_second", "xxx") %>% 
      stringr::str_remove_all("flight_meters_per_second") %>% 
      stringr::str_replace("xxx", "speed (m sec<sup>-1</sup>): ")
      # rest
    , dplyr::across(
      c(estimate,SE,t.ratio,p.value)
      , ~scales::comma(.x, accuracy = 0.001)
    )
  ) %>% 
  kableExtra::kbl(
    caption = "pairwise tests for non-zero differences in trees per hectare error"
    , col.names = c(
      "speed (m sec<sup>-1</sup>)", "contrast"
      # , "speed (m sec<sup>-1</sup>)", "altitude (m)"
      , "estimate", "SE", "t.ratio", "p.value"
    )
    , digits = 3
    , escape = F
  ) %>%
  kableExtra::kable_styling() %>% 
  kableExtra::column_spec(c(2), border_right = TRUE, include_thead = TRUE) %>% 
  kableExtra::collapse_rows(columns = c(1), valign = "top") %>% 
  kableExtra::footnote(
    general_title = "interpretation:"
    , general = c(
      "`estimate`: difference in error; positive means the first setting has higher error."
      , "`SE`: measures the precision of the estimate (smaller values indicate higher precision)"
      , "`t.ratio`: The test statistic; indicates distance from the null (zero difference)"
      , "`p.value`: probability of observing this result by chance; values < 0.05 are considered significant"
    )
    , footnote_as_chunk = FALSE
    , escape = F
  )
```

contrast of **speed** levels at different flight speeds (see the table footnote for interpretation) for trees per hectare per hectare error

```{r}
preds_temp %>% 
  emmeans::contrast(method = "pairwise", simple = "each", combine = TRUE) %>% 
  dplyr::as_tibble() %>%
  dplyr::filter(flight_altitude_m!=".",!is.na(flight_altitude_m)) %>% 
  dplyr::select(flight_altitude_m, contrast,estimate,SE,t.ratio,p.value) %>%
  dplyr::arrange(flight_altitude_m) %>% 
  dplyr::mutate(
    contrast = contrast %>% 
      # clean agl
      stringr::str_replace("flight_altitude_m", "xxx") %>% 
      stringr::str_remove_all("flight_altitude_m") %>% 
      stringr::str_replace("xxx", "altitude (m): ") %>% 
      # clean mph
      stringr::str_replace("flight_meters_per_second", "xxx") %>% 
      stringr::str_remove_all("flight_meters_per_second") %>% 
      stringr::str_replace("xxx", "speed (m sec<sup>-1</sup>): ")
      # rest
    , dplyr::across(
      c(estimate,SE,t.ratio,p.value)
      , ~scales::comma(.x, accuracy = 0.001)
    )
  ) %>% 
  kableExtra::kbl(
    caption = "pairwise tests for non-zero differences in trees per hectare error"
    , col.names = c(
      "altitude (m)", "contrast"
      # , "speed (m sec<sup>-1</sup>)", "altitude (m)"
      , "estimate", "SE", "t.ratio", "p.value"
    )
    , digits = 3
    , escape = F
  ) %>%
  kableExtra::kable_styling() %>% 
  kableExtra::column_spec(c(2), border_right = TRUE, include_thead = TRUE) %>% 
  kableExtra::collapse_rows(columns = c(1), valign = "top") %>% 
  kableExtra::footnote(
    general_title = "interpretation:"
    , general = c(
      "`estimate`: difference in error; positive means the first setting has higher error."
      , "`SE`: measures the precision of the estimate (smaller values indicate higher precision)"
      , "`t.ratio`: The test statistic; indicates distance from the null (zero difference)"
      , "`p.value`: probability of observing this result by chance; values < 0.05 are considered significant"
    )
    , footnote_as_chunk = FALSE
    , escape = F
  )
```

the significant linear trends from the MLR indicate that altitude and speed influence trees per hectare error, but the lack of significant pairwise contrasts suggests these settings are practically equivalent within the operationally feasible range of these flight settings. this result implies that the UAS-lidar and point cloud processing framework is not meaningfully influenced by changes in the flight settings over the range tested here. additionally, given the small sample size with no cross-site replication these findings should be viewed with caution as the lack of significance may reflect low statistical power rather than absence of effect.

### Individual Tree Level

Let's shift to the tree-level detection data to statistically analyze the influence of the flight settings (and tree characteristics) on the probability of individual tree detection. We'll use logistic multiple linear regression

here is our analysis data that includes only the reference trees (TP and FN) and whether or not they were detected during a flight

```{r}
# agg_ground_truth_match_ans %>% dplyr::glimpse()
# ground_truth_prediction_match_ans[[1]] %>% dplyr::glimpse()
tree_detect_df <- ground_truth_prediction_match_ans %>%
  purrr::list_rbind(names_to = "dataset_factor") %>%
  sf::st_drop_geometry() %>% 
  dplyr::filter(match_grp %in% c("true positive","omission")) %>% 
  dplyr::inner_join(
    tracking_df %>% 
      dplyr::select(-c(path))
    , by = "dataset_factor"
  ) %>% 
  dplyr::mutate(
    is_detected = match_grp=="true positive"
    , is_detected = as.numeric(is_detected)
    , dataset_factor = ordered(dataset_factor, levels(tracking_df$dataset_factor))
  )
# dplyr::glimpse(tree_detect_df)
tree_detect_df %>% 
  dplyr::group_by(dataset_factor) %>% 
  dplyr::summarise(
    n = dplyr::n()
    , n_detected = sum(is_detected)
    , prop_detected = mean(is_detected)
  ) %>% 
  dplyr::mutate(
    n = scales::comma(n,accuracy=1)
    , n_detected = scales::comma(n_detected,accuracy=1)
    , prop_detected = scales::percent(prop_detected,accuracy=0.1)
  ) %>% 
  kableExtra::kbl(
    caption = "summary of single tree dataset for Pr(detection) analysis"
    , col.names = c("flight", "reference trees", "detected trees","% detected<br>(recall)")
    , escape = F
  ) %>%
  kableExtra::kable_styling()
```

note, this aggregation is exactly how recall (detection rate) is calculated

fit the logistic model on this data without interacting the flight parameters of speed and altitude with terrain following since we don't have the data to estimate the influence of terrain following by speed or altitude

```{r}
# logistic
mod_glm_temp <- stats::glm(
  is_detected ~ 
    # main effects
    flight_altitude_m + flight_meters_per_second + ref_tree_height_m +     
    # two-way interactions
    flight_altitude_m:flight_meters_per_second +
    ref_tree_height_m:flight_altitude_m +
    ref_tree_height_m:flight_meters_per_second
    # ref_tree_height_m:flight_tf
  , data = tree_detect_df
  , family = stats::binomial(link = "logit")
)
# summary(mod_glm_temp)
```

the model form is: 

$$
\begin{split} \ln\left(\frac{\text{Pr(Tree Detected)}}{1 - \text{Pr(Tree Detected)}}\right) = & \beta_0 + \beta_1(\text{Tree Height}) + \beta_2(\text{Altitude}) + \beta_3(\text{Speed}) \\ & + \beta_4(\text{Altitude} \times \text{Speed}) \\ & + \beta_5(\text{Tree Height} \times \text{Altitude}) \\ & + \beta_6(\text{Tree Height} \times \text{Speed}) \end{split}
$$

logistic model of Pr(tree detected) coefficient table 

```{r}
broom::tidy(mod_glm_temp) %>%
  dplyr::select(term,estimate,std.error,p.value) %>% 
  dplyr::mutate(
    odds_ratio = exp(estimate)
    , term = term %>% 
      stringr::str_replace_all("flight_tf","terrain follow") %>% 
      stringr::str_replace_all("flight_altitude_m","altitude") %>% 
      stringr::str_replace_all("flight_meters_per_second","speed") %>% 
      stringr::str_replace_all("ref_tree_height_m","tree ht") %>% 
      stringr::str_replace_all("FALSE"," = off") %>% 
      stringr::str_replace_all("TRUE"," = on")
  ) %>%
  kableExtra::kbl(
    caption = "GLM coefficients and odds ratios"
    , digits = 3
  ) %>%
  kableExtra::kable_styling()
```

```{r, include = F, eval=TRUE, echo=F, warning=F, results=F}
huh_temp <- broom::tidy(mod_glm_temp) %>% dplyr::filter(term=="ref_tree_height_m") %>% dplyr::mutate(odds_ratio = exp(estimate))
# huh_temp
# ifelse(huh_temp$odds_ratio < 1, scales::percent(1-huh_temp$odds_ratio,accuracy=0.1), scales::percent(huh_temp$odds_ratio-1,accuracy=0.1))
```

the tree height is the only coefficient estimated to influence the probability of tree detection in a meaningful (and statistically significant with a p-value of `r ifelse(huh_temp$p.value < 0.001, "< 0.001",scales::comma(huh_temp$p.value, accuracy = 0.001))`) way. the odds ratio on reference tree height of `r scales::comma(huh_temp$odds_ratio,accuracy = 0.001)` means that for every one unit increase in tree height, the odds of UAS-lidar detection `r ifelse(huh_temp$odds_ratio < 1, "decrease", "increase")` by `r ifelse(huh_temp$odds_ratio < 1, scales::percent(1-huh_temp$odds_ratio,accuracy=0.1), scales::percent(huh_temp$odds_ratio-1,accuracy=0.1))`

let's look at the marginal effects at different levels of the metric variables in the model (tree height, flight altitude, flight speed) using the `emmeans` [package](https://rvlenth.github.io/emmeans/articles/interactions.html#covariates) to plot the predicted detection probabilities

we'll get the predictions over a range of each of the independent variables using `emmeans::emmeans()` with the range of the observed data for each metric variable (so we aren't making out of sample predictions). then, we'll look at the predictions at select tree heights

```{r}
# emmeans
preds_temp <-
  emmeans::emmeans(
    mod_glm_temp
    , ~ ref_tree_height_m + flight_altitude_m + flight_meters_per_second
    , at = list(
      ref_tree_height_m = 
        seq(
          floor( min(tree_detect_df$ref_tree_height_m) )
          , floor( max(tree_detect_df$ref_tree_height_m)*1.05 )
          , by = 1
        )
      , flight_altitude_m = 
        unique(tree_detect_df$flight_altitude_m)
        # seq(
        #   floor( min(tree_detect_df$flight_altitude_m) )
        #   , floor( max(tree_detect_df$flight_altitude_m) )
        #   , by = round(100/3.281)
        # )
      , flight_meters_per_second = 
        seq(
          floor( min(tree_detect_df$flight_meters_per_second) )
          , floor( max(tree_detect_df$flight_meters_per_second) )
          , by = round(5/2.237)
        )
    )
    , type = "response" # probabilities
  ) %>% 
 dplyr::as_tibble()
# huh?
# dplyr::glimpse(preds_temp)
# table
# fivenum(tree_detect_df$ref_tree_height_m)[c(1,3,5)]
preds_temp %>%
  dplyr::filter(
    ref_tree_height_m %in% 
      (quantile(tree_detect_df$ref_tree_height_m, probs = c(0.05,0.25,0.5,0.75,0.95)) %>% 
        round() %>% 
        unique())
  ) %>%
  dplyr::mutate(dplyr::across(
    c(prob, asymp.LCL, asymp.UCL)
    , ~scales::percent(.x,accuracy=0.1)
  )) %>% 
  dplyr::select(ref_tree_height_m, flight_altitude_m, flight_meters_per_second, prob, asymp.LCL, asymp.UCL) %>%
  dplyr::arrange(ref_tree_height_m, flight_altitude_m, flight_meters_per_second) %>%
  # dplyr::select(ref_tree_height_m, flight_meters_per_second, flight_altitude_m, prob, asymp.LCL, asymp.UCL) %>%
  # dplyr::arrange(ref_tree_height_m, flight_meters_per_second, flight_altitude_m) %>%
  kableExtra::kbl(
    caption = "tree detection probability at 5%, 25%, 50%, 75%, and 95% tree heights"
    , col.names = c(
      "tree ht (m)"
      , "altitude (m)", "speed (m sec<sup>-1</sup>)"
      # , "speed (m sec<sup>-1</sup>)", "altitude (m)"
      , "Pr(detection)", "lower bound", "upper bound"
    )
    , escape = F
    # , digits = 2
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::column_spec(c(3), border_right = TRUE, include_thead = TRUE) %>% 
  kableExtra::collapse_rows(columns = c(1:2), valign = "top") %>% 
  kableExtra::scroll_box(height = "7.5in", width = "100%")
```

let's check out the conditional influence of flight speed at different altitude levels. in these detection probability curves higher lines mean better detection and steeper lines mean higher sensitivity to tree height

```{r}
ggplot2::ggplot(
  preds_temp
  , mapping = ggplot2::aes(
    x = ref_tree_height_m
    , y = prob
    , color = as.factor(flight_meters_per_second )
  )
) +
  ggplot2::geom_ribbon(
    mapping = ggplot2::aes(
      ymin = asymp.LCL
      , ymax = asymp.UCL
      , fill = as.factor(flight_meters_per_second)
    )
    , alpha = 0.05, color = NA
  ) +
  ggplot2::geom_line(lwd = 1.05, alpha = 0.85) +
  ggplot2::facet_grid(
    cols = dplyr::vars(flight_altitude_m)
    , labeller = ggplot2::labeller( flight_altitude_m = ~paste("altitude:", .x, " (m)") )
  ) +
  ggplot2::scale_color_viridis_d(option = "rocket", begin = 0.9, end = 0.5) +
  ggplot2::scale_fill_viridis_d(option = "rocket", begin = 0.9, end = 0.5) +
  ggplot2::scale_y_continuous(
    limits = c(0,1)
    , labels=scales::percent
    , breaks = scales::breaks_extended(n=6)
    # , expand = ggplot2::expansion(mult = c(0,0.03))
  ) +
  ggplot2::scale_x_continuous(
    labels=scales::comma
    , breaks = scales::breaks_extended(n=8)
  ) +
  ggplot2::labs(
    color = latex2exp::TeX("speed  $m \\cdot sec^{-1}$ "), fill = latex2exp::TeX("speed  $m \\cdot sec^{-1}$ ")
    , x = "tree height (m)", y = "detection probability"
    # , title = "detection probability curves"
    # , subtitle = "conditional effect of flight speed"
    # , caption = "higher lines = better detection. steeper lines = higher sensitivity to tree height"
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "top"
    , strip.text.x = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , strip.text.y = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , plot.title = ggplot2::element_text(size = 10)
    , plot.subtitle = ggplot2::element_text(size = 10)
  ) +
  ggplot2::guides(
    color = ggplot2::guide_legend(
      override.aes = list(shape = 15, lwd = 8, fill = NA)
    )
    , fill = "none"
  )

```

```{r, include=F, eval=T, results=F}
ggplot2::ggsave("../data/pr_detect_by_speed.jpg", height = 7.2, width = 9)
```

The wide confidence bounds with all lines contained within the overlapping bounds indicate that at all flight altitudes, the flight speed does not impact tree detection. There is some evidence that the higher the flight altitude, the less impactful speed is on the resulting tree detection (but not statistically significant).

now the conditional influence of flight altitude at different speeds. in these detection probability curves higher lines mean better detection and steeper lines mean higher sensitivity to tree height

```{r}
ggplot2::ggplot(
  preds_temp
  , mapping = ggplot2::aes(
    x = ref_tree_height_m
    , y = prob
    , color = as.factor(flight_altitude_m)
  )
) +
  ggplot2::geom_ribbon(
    mapping = ggplot2::aes(
      ymin = asymp.LCL
      , ymax = asymp.UCL
      , fill = as.factor(flight_altitude_m)
    )
    , alpha = 0.05, color = NA
  ) +
  ggplot2::geom_line(lwd = 1.05, alpha = 0.85) +
  ggplot2::facet_grid(
    cols = dplyr::vars(flight_meters_per_second)
    , labeller = ggplot2::labeller( flight_meters_per_second = ~ paste("speed:", .x, "(m s\u207B\u00B9)") )
  ) +
  ggplot2::scale_color_viridis_d(option = "mako", begin = 0.5, end = 0.1) +
  ggplot2::scale_fill_viridis_d(option = "mako", begin = 0.5, end = 0.1) +
  ggplot2::scale_y_continuous(
    limits = c(0,1)
    , labels=scales::percent
    , breaks = scales::breaks_extended(n=6)
    # , expand = ggplot2::expansion(mult = c(0,0.03))
  ) +
  ggplot2::scale_x_continuous(
    labels=scales::comma
    , breaks = scales::breaks_extended(n=8)
  ) +
  ggplot2::labs(
    color = "altitude (m)", fill = "altitude (m)"
    , x = "tree height (m)", y = "detection probability"
    # , title = "detection probability curves"
    # , subtitle = "conditional effect of flight altitude"
    # , caption = "higher lines = better detection. steeper lines = higher sensitivity to tree height"
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "top"
    , strip.text.x = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , strip.text.y = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , plot.title = ggplot2::element_text(size = 10)
    , plot.subtitle = ggplot2::element_text(size = 10)
  ) +
  ggplot2::guides(
    color = ggplot2::guide_legend(override.aes = list(shape = 15, lwd = 8, fill = NA))
    , fill = "none"
  )
```

```{r, include=F, eval=T, results=F}
ggplot2::ggsave("../data/pr_detect_by_alt.jpg", height = 7.2, width = 9)
```

The wide confidence bounds with all lines contained within the overlapping bounds indicate that at all flight speeds, the flight altitude does not impact tree detection. There is some evidence that the higher the flight speed, the less impactful altitude is on the resulting tree detection (but not statistically significant).

let's do some contrasts of the probability of detection while holding the tree height constant at the median (`r scales::comma(median(tree_detect_df$ref_tree_height_m),accuracy=0.1)` m) to test an altitude level against every other altitude and each speed against every other speed.

```{r}
em_temp <- 
  emmeans::emmeans(
    mod_glm_temp
    , ~ flight_altitude_m + flight_meters_per_second
    , at = list(
      ref_tree_height_m = median(tree_detect_df$ref_tree_height_m)
      , flight_altitude_m = 
        unique(tree_detect_df$flight_altitude_m)
        # seq(
        #   floor( min(tree_detect_df$flight_altitude_m) )
        #   , floor( max(tree_detect_df$flight_altitude_m) )
        #   , by = round(100/3.281)
        # )
      , flight_meters_per_second = 
        seq(
          floor( min(tree_detect_df$flight_meters_per_second) )
          , floor( max(tree_detect_df$flight_meters_per_second) )
          , by = round(5/2.237)
        )
    )
    , type = "response" # probabilities
  )
```

contrast of altitude levels at different flight speeds (see the table footnote for interpretation) for the median (`r scales::comma(median(tree_detect_df$ref_tree_height_m),accuracy=0.1)` m) tree

```{r}
# em_temp
# pairwise test for non-zero differences
# simple = "each" tests altitudes within each speed and vice versa
em_temp %>% 
  emmeans::contrast(method = "pairwise", simple = "each", combine = TRUE) %>% 
  dplyr::as_tibble() %>%
  dplyr::filter(flight_meters_per_second!=".",!is.na(flight_meters_per_second)) %>% 
  dplyr::select(flight_meters_per_second, contrast,odds.ratio,SE,z.ratio,p.value) %>% 
  dplyr::arrange(flight_meters_per_second) %>% 
  dplyr::mutate(
    contrast = contrast %>% 
      # clean agl
      stringr::str_replace("flight_altitude_m", "xxx") %>% 
      stringr::str_remove_all("flight_altitude_m") %>% 
      stringr::str_replace("xxx", "altitude (m): ") %>% 
      # clean mph
      stringr::str_replace("flight_meters_per_second", "xxx") %>% 
      stringr::str_remove_all("flight_meters_per_second") %>% 
      stringr::str_replace("xxx", "speed (m sec<sup>-1</sup>): ")
      # rest
    , dplyr::across(
      c(odds.ratio,SE,z.ratio,p.value)
      , ~scales::comma(.x, accuracy = 0.001)
    )
  ) %>% 
  kableExtra::kbl(
    caption = "pairwise tests for non-zero differences (at median tree height)"
    , col.names = c(
      "speed (m sec<sup>-1</sup>)", "contrast"
      # , "speed (m sec<sup>-1</sup>)", "altitude (m)"
      , "odds.ratio", "SE", "z.ratio", "p.value"  
    )
    , digits = 3
    , escape = F
  ) %>%
  kableExtra::kable_styling() %>% 
  kableExtra::column_spec(c(2), border_right = TRUE, include_thead = TRUE) %>% 
  kableExtra::collapse_rows(columns = c(1), valign = "top") %>% 
  kableExtra::footnote(
    general_title = "interpretation:"
    , general = c(
      "`odds.ratio`: <br> values near 1 mean levels perform similarly <br> >1 means the first level is better <br> <1 means the second is better at tree detection"
      , "`SE`: measures the precision of the estimate (smaller values indicate higher precision)"
      # , "z.ratio: The test statistic; indicates how many standard deviations the result is from the null (0)."
      , "`z.ratio`: >0 means the first level detection is higher, <0 means the second level detection is higher"
      , "`p.value`: probability of observing this result by chance; values < 0.05 are considered significant"
    )
    , footnote_as_chunk = FALSE
    , escape = F
  )
```

contrast of speed levels at different flight altitudes (see the table footnote for interpretation) for the median (`r scales::comma(median(tree_detect_df$ref_tree_height_m),accuracy=0.1)` m) tree

```{r}
# em_temp
# pairwise test for non-zero differences
# simple = "each" tests altitudes within each speed and vice versa
em_temp %>% 
  emmeans::contrast(method = "pairwise", simple = "each", combine = TRUE) %>% 
  dplyr::as_tibble() %>%
  dplyr::filter(flight_altitude_m!=".",!is.na(flight_altitude_m)) %>% 
  dplyr::select(flight_altitude_m, contrast,odds.ratio,SE,z.ratio,p.value) %>% 
  dplyr::arrange(flight_altitude_m) %>% 
  dplyr::mutate(
    contrast = contrast %>% 
      # clean agl
      stringr::str_replace("flight_altitude_m", "xxx") %>% 
      stringr::str_remove_all("flight_altitude_m") %>% 
      stringr::str_replace("xxx", "altitude (m): ") %>% 
      # clean mph
      stringr::str_replace("flight_meters_per_second", "xxx") %>% 
      stringr::str_remove_all("flight_meters_per_second") %>% 
      stringr::str_replace("xxx", "speed (m sec<sup>-1</sup>): ")
      # rest
    , dplyr::across(
      c(odds.ratio,SE,z.ratio,p.value)
      , ~scales::comma(.x, accuracy = 0.001)
    )
  ) %>% 
  kableExtra::kbl(
    caption = "pairwise tests for non-zero differences (at median tree height)"
    , col.names = c(
      "altitude (m)", "contrast"
      # , "speed (m sec<sup>-1</sup>)", "altitude (m)"
      , "odds.ratio", "SE", "z.ratio", "p.value"  
    )
    , digits = 3
    , escape = F
  ) %>%
  kableExtra::kable_styling() %>% 
  kableExtra::column_spec(c(2), border_right = TRUE, include_thead = TRUE) %>% 
  kableExtra::collapse_rows(columns = c(1), valign = "top") %>% 
  kableExtra::footnote(
    general_title = "interpretation:"
    , general = c(
      "`odds.ratio`: <br> values near 1 mean levels perform similarly <br> >1 means the first level is better <br> <1 means the second is better at tree detection"
      , "`SE`: measures the precision of the estimate (smaller values indicate higher precision)"
      # , "z.ratio: The test statistic; indicates how many standard deviations the result is from the null (0)."
      , "`z.ratio`: >0 means the first level detection is higher, <0 means the second level detection is higher"
      , "`p.value`: probability of observing this result by chance; values < 0.05 are considered significant"
    )
    , footnote_as_chunk = FALSE
    , escape = F
  )
```

plot the probability of detection for the median (`r scales::comma(median(tree_detect_df$ref_tree_height_m),accuracy=0.1)` m) tree for all settings with 95% confidence intervals. if the error bars do not overlap, the difference is likely non-zero and vice-versa

the influence of speed at different altitudes

```{r}
em_temp %>% 
  as.data.frame() %>% 
  ggplot2::ggplot(
    mapping = ggplot2::aes(
      x = 0
        # as.factor(flight_altitude_m)
      , y = prob
      , color = as.factor(flight_meters_per_second )
    )
  ) +
  ggplot2::geom_point(
    size = 4
    , position = ggplot2::position_dodge(width = 0.5)
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(ymin = asymp.LCL, ymax = asymp.UCL)
    , width = 0.2
    , position = ggplot2::position_dodge(width = 0.5)
  ) +
  ggplot2::facet_grid(
    cols = dplyr::vars(flight_altitude_m)
    , labeller = ggplot2::labeller( flight_altitude_m = ~paste("altitude:", .x, " (m)") )
  ) +
  ggplot2::scale_color_viridis_d(option = "rocket", begin = 0.9, end = 0.5) +
  ggplot2::scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  ggplot2::scale_x_continuous(NULL, breaks = NULL) +
  ggplot2::labs(
    title = "comparison of detection probabilities at median tree height"
    , y = "detection probability"
    , x = "altitude (m)"
    , color = latex2exp::TeX("Speed  $m \\cdot sec^{-1}$ ")
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "top"
    , strip.text.x = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , strip.text.y = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , plot.title = ggplot2::element_text(size = 10)
    , plot.subtitle = ggplot2::element_text(size = 10)
  ) +
  ggplot2::guides(
    color = ggplot2::guide_legend(override.aes = list(shape = 15, lwd = 8, fill = NA))
    , fill = "none"
  )


```

the influence of altitude at different speeds

```{r}
em_temp %>% 
  as.data.frame() %>% 
  ggplot2::ggplot(
    mapping = ggplot2::aes(
      x = 0
      , y = prob
      , color = as.factor(flight_altitude_m)
    )
  ) +
  ggplot2::geom_point(
    size = 4
    , position = ggplot2::position_dodge(width = 0.5)
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(ymin = asymp.LCL, ymax = asymp.UCL)
    , width = 0.2
    , position = ggplot2::position_dodge(width = 0.5)
  ) +
  ggplot2::facet_grid(
    cols = dplyr::vars(flight_meters_per_second)
    , labeller = ggplot2::labeller( flight_meters_per_second = ~ paste("speed:", .x, "(m s\u207B\u00B9)") )
  ) +
  ggplot2::scale_color_viridis_d(option = "mako", begin = 0.5, end = 0.1) +
  ggplot2::scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  ggplot2::scale_x_continuous(NULL, breaks = NULL) +
  ggplot2::labs(
    title = "comparison of detection probabilities at median tree height"
    , y = "detection probability"
    , x = "speed (m sec<sup>-1</sup>)"
    , color = "altitude (m)"
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "top"
    , strip.text.x = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , strip.text.y = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , plot.title = ggplot2::element_text(size = 10)
    , plot.subtitle = ggplot2::element_text(size = 10)
  ) +
  ggplot2::guides(
    color = ggplot2::guide_legend(override.aes = list(shape = 15, lwd = 8, fill = NA))
    , fill = "none"
  )


```

yeah, no difference in the detection accuracy across the flight settings tested for this forest stand

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```
