---
title: "influence of UAS-lidar flight parameters on tree detection at N1 study site"
subtitle: ""
author: "George Woolsey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding){ 
    out_dir <- '../';
    rmarkdown::render(inputFile, encoding = encoding, output_file=file.path(dirname(inputFile), out_dir, 'index.html'));
    file.copy(from = "..//index.html", to = '..///data//uas_lidar_n1.html', overwrite = TRUE)
  })
---

```{r, include=FALSE, warning=F, message=F}
################################################
##```{r,eval=T
################################################
# clean session
remove(list = ls())
gc()
# knit options
knitr::opts_chunk$set(
  echo = TRUE
  , warning = FALSE
  , message = FALSE
  # , results = 'hide'
  , fig.width = 10.5
  , fig.height = 7
)
# option to put satellite imagery as base layer of mapview maps
  mapview::mapviewOptions(
    homebutton = FALSE
    # , basemaps = c("Esri.WorldImagery","OpenStreetMap")
    , basemaps = c("OpenStreetMap", "Esri.WorldImagery")
  )
```

# Data

load the standard libraries we use to do work 

```{r}
# bread-and-butter
library(tidyverse) # the tidyverse
library(viridis) # viridis colors
library(harrypotter) # hp colors
library(RColorBrewer) # brewer colors
library(scales) # work with number and plot scales
library(latex2exp)

# visualization
library(mapview) # interactive html maps
library(kableExtra) # tables
library(patchwork) # combine plots
library(ggnewscale) # new scale
library(ggrepel) # repel labels

# spatial analysis
library(terra) # raster
library(sf) # simple features
library(lidR) # lidR
library(cloud2trees) # cloud2trees
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Data Location

Let's check out the location of the data we got

```{r,eval=T,purl=FALSE}
# directory with the downloaded .las|.laz files
f <- "../data/N1_400AGL_20MPH_TFOFF"
# f <- "d:/BLM_CO_SWDF_DawsonFuelsTreatment/Final/PointCloud/Full/"
# is there data?
list.files(f, pattern = ".*\\.(laz|las)$") %>% length()
# what files are in here?
list.files(f, pattern = ".*\\.(laz|las)$")[1]
```

what information does `lidR` read from the catalog?

```{r,eval=T,purl=FALSE}
ctg_temp <- lidR::readLAScatalog(f)
ctg_temp
```

that's a lot of points...can an ordinary laptop handle it? we'll find out.

We'll plot our point cloud data tiles real quick

```{r,eval=T,purl=FALSE}
ctg_temp@data %>%
  dplyr::mutate(
    nm =
      basename(filename) %>%
      stringr::str_remove("\\.[^.]+$")
  ) %>%
  ggplot2::ggplot() + 
  ggplot2::geom_sf(fill=NA, color = "red") +
  ggplot2::theme_light()
```

let's look at this on a map to orient ourselves

```{r,eval=T,purl=FALSE}
ctg_temp %>% 
  cloud2trees:::check_las_ctg_empty() %>% 
  purrr::pluck("data") %>% 
  mapview::mapview(popup = F, layer.name = "point cloud tile")
```

read RGB

```{r}
#### read RGB data keep only RGB
rgb_rast_fnm <- "../data/dom/dom.tif"
```

```{r,eval=T,purl=FALSE}
rgb_rast <- terra::rast(rgb_rast_fnm) 
# rgb_rast %>% 
# (NIR-R-G)
rgb_rast %>% 
  terra::subset(c(1,2,3)) %>% 
  terra::plotRGB()
# rgb_rast %>% 
#   terra::subset(c(4,1,2)) %>% 
#   terra::plotRGB()
```

pc extent vs rgb rast

```{r}
rgb_rast %>% 
  terra::subset(c(1,2,3)) %>% 
  terra::plotRGB()
ctg_temp %>% 
  cloud2trees:::check_las_ctg_empty() %>% 
  purrr::pluck("data") %>% 
  sf::st_transform(terra::crs(rgb_rast)) %>% 
  terra::vect() %>% 
  terra::plot(add = T, col = NA, border = "gold", lwd = 3)
```

## ITD Tuning

```{r}
itd_tuning_ans <- 
  cloud2trees::itd_tuning(
    input_las_dir = "../data/N1_400AGL_20MPH_TFOFF/"
    , n_samples = 2
    , min_height = 1.37
    , chm_res_m = 0.25
  )
names(itd_tuning_ans)
itd_tuning_ans$plot_samples
# ggplot2::ggsave(filename = "../data/itd_tuning_plot_samples1.jpg", height = 11, width = 8, dpi = "print")
itd_tuning_ans$plot_sample_summary
# ggplot2::ggsave(filename = "../data/itd_tuning_plot_sample_summary1.jpg", height = 11, width = 8, dpi = "print")
```

look at the default window functions from `cloud2trees`

```{r}
itd_tuning_ans$ws_fn_list$lin_fn
# plot the ws fn
ggplot2::ggplot() +
  ggplot2::geom_function(fun = itd_tuning_ans$ws_fn_list$lin_fn, aes(color = "lin_fn"), lwd = 1) +
  ggplot2::geom_function(fun = itd_tuning_ans$ws_fn_list$exp_fn, aes(color = "exp_fn"), lwd = 1) +
  ggplot2::xlim(-5,60) +
  ggplot2::labs(x = "heights", y = "ws", color = "") +
  ggplot2::theme_light()

# define a custom function
my_linear <- function(x){
    y <- dplyr::case_when(
      is.na(x) ~ 0.001
      , x < 0 ~ 0.001
      , x < 2 ~ 1
      , x < 10 ~ 0.75 + (x * 0.2)
      , x == 10 ~ 0.75 + (10 * 0.14)
      , x > 30 ~ 5
      , TRUE ~ 0.75 + (x * 0.14)
    )
    return(y)
}
ggplot2::ggplot() +
  ggplot2::geom_function(fun = itd_tuning_ans$ws_fn_list$log_fn, aes(color = "log_fn"), lwd = 1) +
  ggplot2::geom_function(fun = itd_tuning_ans$ws_fn_list$lin_fn, aes(color = "lin_fn"), lwd = 1) +
  ggplot2::geom_function(fun = itd_tuning_ans$ws_fn_list$exp_fn, aes(color = "exp_fn"), lwd = 1) +
  ggplot2::geom_function(fun = my_linear, aes(color = "my_linear"), lwd = 1) +
  ggplot2::xlim(-5,60) +
  ggplot2::labs(x = "heights", y = "ws", color = "") +
  ggplot2::theme_light()
```

re-run tuning with new function

```{r}
# let's put these in a list to test with the best default function we saved from above
my_fn_list <- list(
  lin_fn = cloud2trees::itd_ws_functions()[["lin_fn"]]
  , exp_fn = cloud2trees::itd_ws_functions()[["exp_fn"]]
  , my_linear = my_linear
)
itd_tuning_ans2 <- 
  cloud2trees::itd_tuning(
    input_las_dir = "../data/N1_400AGL_20MPH_TFOFF/"
    , ws_fn_list = my_fn_list
    , n_samples = 4
    , min_height = 1.37
    , chm_res_m = 0.25
  )
itd_tuning_ans2$plot_samples
ggplot2::ggsave(filename = "../data/itd_tuning_plot_samples2.jpg", height = 11, width = 8, dpi = "print")
itd_tuning_ans$plot_sample_summary
# ggplot2::ggsave(filename = "../data/itd_tuning_plot_sample_summary1.jpg", height = 11, width = 8, dpi = "print")
itd_tuning_ans2$crowns %>% dplyr::glimpse()
itd_tuning_ans2$crowns %>% 
  sf::st_drop_geometry() %>% 
  dplyr::count(sample_number,ws_fn)
```

with the crowns data we can explore alternative visualizations including overlaying the detected trees on the RGB data if available

plot it

```{r,eval=T,purl=FALSE}
# make a function to plot these detected crowns with rgb data
plt_rgb_rast_itd_crowns <- function(sample_nmbr = 1, rgb_rast, itd_crowns, plt_lwd = 0.7) {
  # crop
  crp_rgb_rast_temp <- rgb_rast %>% 
    terra::crop(
      itd_crowns %>% 
        dplyr::ungroup() %>% 
        dplyr::filter(sample_number == sample_nmbr) %>% 
        sf::st_union() %>% 
        sf::st_bbox() %>% 
        sf::st_as_sfc() %>% 
        sf::st_buffer(0.2) %>% 
        sf::st_transform(terra::crs(rgb_rast)) %>% 
        terra::vect()
    )
  # convert raster to a data frame and create hex colors
  # ?grDevices::rgb
  rgb_df_temp <-
    crp_rgb_rast_temp %>% 
    terra::as.data.frame(xy = TRUE) %>%
    dplyr::rename(
      red = 3, green = 4, blue = 5
    ) %>% 
    dplyr::mutate(
      hex_col = grDevices::rgb(
        red
        , green
        , blue
        , maxColorValue = 255
      )
    )
    # dplyr::glimpse()
  
  # plt
  plt <- ggplot2::ggplot() +
    # add rgb base map
    ggplot2::geom_raster(data = rgb_df_temp, mapping = ggplot2::aes(x = x, y = y, fill = hex_col)) +
    # use identity scale so the hex codes are used directly
    ggplot2::scale_fill_identity() +
    # overlay polygons
    # ggplot2::geom_sf(data = polys, fill = NA, color = "red", linewidth = 0.5) +
    ggplot2::geom_sf(
      data = itd_crowns %>% 
        dplyr::filter(sample_number==sample_nmbr) %>% 
        cloud2trees::simplify_multipolygon_crowns() %>% 
        sf::st_make_valid() %>% 
        dplyr::filter(sf::st_is_valid(.)) %>% 
        sf::st_transform(terra::crs(crp_rgb_rast_temp))
      , mapping = ggplot2::aes(color = ws_fn)
      , fill = NA
      , lwd = plt_lwd
      , inherit.aes = F
    ) +
    ggplot2::facet_grid(cols = dplyr::vars(ws_fn)) +
    ggplot2::scale_color_viridis_d(name = "") +
    ggplot2::coord_sf(expand = F) +
    ggplot2::theme_void() +
    ggplot2::theme(
      legend.position = "none"
      , strip.text = ggplot2::element_text(face = "bold", color = "black", margin = ggplot2::margin(t = 4, b = 4))
      , strip.background = ggplot2::element_rect(fill = "gray88", color = "gray88")
      , panel.spacing = ggplot2::unit(1,"lines")
    )
  return(plt)
}
```

plot the trees detected in sample one on the RGB

```{r,eval=T,purl=FALSE, fig.height=5.5, fig.width=7.2}
plt_rgb_temp <- plt_rgb_rast_itd_crowns(
  sample_nmbr = 3
  , rgb_rast = rgb_rast %>% terra::subset(c(1:3))
  , itd_crowns = itd_tuning_ans2$crowns
)

rgb_rast %>% 
  terra::crop(
    itd_tuning_ans2$crowns %>% 
    dplyr::filter(sample_number==2, ws_fn == "lin_fn") %>% 
    sf::st_union() %>% 
    sf::st_bbox() %>% 
    sf::st_as_sfc() %>% 
    sf::st_buffer(5) %>% 
    sf::st_transform(terra::crs(rgb_rast)) %>% 
    terra::vect()
  ) %>% 
  terra::plotRGB()
itd_tuning_ans2$crowns %>% 
  dplyr::filter(sample_number==2, ws_fn == "lin_fn") %>% 
  sf::st_transform(terra::crs(rgb_rast)) %>% 
  terra::vect() %>% 
  terra::plot(add=T, col = NA, border = "cyan", lwd = 2.5)
itd_tuning_ans2$crowns %>% 
  dplyr::filter(sample_number==2, ws_fn == "exp_fn") %>% 
  sf::st_transform(terra::crs(rgb_rast)) %>% 
  terra::vect() %>% 
  terra::plot(add=T, col = NA, border = "gold", lwd = 1.5)
  

```



## Point Cloud to Raster: `cloud2raster()`

Although the `cloud2trees::cloud2trees()` function combines methods in the `cloud2trees` package for an all-in-one approach, we'll instead use the `cloud2trees::cloud2raster()` function to generate a CHM from the point cloud that we can work with to perform both individual tree detection and slash pile identification (discussed later).

We'll set the options in the function to generate a CHM which represents a DSM with the ground removed and no other filtering. This high resolution (i.e. fine-grain) CHM will serve as the foundation for tree detection and slash pile detection as we can manipulate it to optimize the processing for both methods.

```{r}
# outdir
c2t_output_dir <- "../data"
c2t_process_dir <- file.path(c2t_output_dir, "point_cloud_processing_delivery")
c2t_tracking_fnm <- file.path(c2t_process_dir, "processed_tracking_data.csv")
##############################################################
# cloud2trees::cloud2raster
##############################################################
if(
  !file.exists( file.path(c2t_process_dir, "chm_0.1m.tif") )
  || !file.exists( file.path(c2t_process_dir, "dtm_0.25m.tif") )
){
  # time it
  st_temp <- Sys.time()
  # run it
  # cloud2trees
  cloud2raster_ans <- cloud2trees::cloud2raster(
    output_dir = c2t_output_dir
    , input_las_dir = f
    , accuracy_level = 2
    , keep_intrmdt = T
    , dtm_res_m = 0.25
    , chm_res_m = 0.1
    , min_height = 0 # effectively generates a DSM based on non-ground points
  )
  
  # timer
  mins_temp <- difftime(Sys.time(),st_temp,units = "mins") %>% as.numeric()
  # save tracking
  dplyr::tibble(
    timer_cloud2raster_mins = mins_temp
  ) %>% 
    write.csv(
      file = c2t_tracking_fnm
      , row.names = F, append = F
    )
}else{
  dtm_temp <- terra::rast( file.path(c2t_process_dir, "dtm_0.25m.tif") )
  chm_temp <- terra::rast( file.path(c2t_process_dir, "chm_0.1m.tif") )
  
  cloud2raster_ans <- list(
    "dtm_rast" = dtm_temp
    , "chm_rast" = chm_temp
  )
}
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
remove(f)
gc()
```

let's see what we got from `cloud2trees::cloud2raster()`

```{r}
cloud2raster_ans %>% names()
```

there's a DTM

```{r}
# plot to check out the fine-resolution DTM raster
cloud2raster_ans$dtm_rast %>% 
  terra::plot(col = harrypotter::hp(n=100, option = "mischief"), main = "DTM (m)")
```

there's a CHM

```{r}
# plot to check out the fine-resolution CHM raster
cloud2raster_ans$chm_rast %>% 
  terra::plot(col = viridis::plasma(n=100), main = "CHM (m)")
```

let's see some details about the CHM

```{r}
# what chm?
cloud2raster_ans$chm_rast
# what data?
cloud2raster_ans$chm_rast %>% terra::summary()
```

for tree detection, we'll aggregate the CHM to a lower resolution (i.e. coarser) to smooth out some of the fine detail that can increase the noise in the tree detection processing. This aggregation process also will speed up the processing and reduce the chances of computational memory issues.

```{r}
##################################################################
# aggregate to make raster more coarse for tree detection
##################################################################
# first, we'll borrow from the `cloud2trees` codebase to get a function to change the resolution of a raster exactly
###___________________________________________###
# adjust the resolution of a raster to be in exactly the target resolution
###___________________________________________###
if(
  !file.exists( file.path(c2t_process_dir, "chm_0.25m.tif") )
){
  agg_chm_rast <- cloud2trees:::adjust_raster_resolution(
    cloud2raster_ans$chm_rast
    , target_resolution = 0.25
    , fun = max
    , resample_method = "max"
    , ofile = file.path(c2t_process_dir, "chm_0.25m.tif")
  )
}else{
  agg_chm_rast <- terra::rast( file.path(c2t_process_dir, "chm_0.25m.tif") )
}
# what chm?
agg_chm_rast
# what data?
agg_chm_rast %>% terra::summary()
```

# Individial Tree Detection Tuning: `itd_tuning()`{#itd}

The `cloud2trees` package performs individual tree detection using `lidR::locate_trees()` with the `lidR::lmf()` algorithm. The local maximum filter algorithm allows for a constant window size or a variable window size defined by a function. See the `lidR` [package book section](https://r-lidar.github.io/lidRbook/itd.html) by point cloud processing expert [Jean-Romain Roussel](https://github.com/Jean-Romain) for excellent detail on ITD and defining window size.

The `itd_tuning()` function is used to visually assess tree crown delineation results from different window size functions used for the detection of individual trees. `itd_tuning()` allows users to test different window size functions on a sample of data to determine which function is most suitable for the area being analyzed. The preferred function can then be used in the `ws` parameter in `raster2trees()` and `cloud2trees()`.

Let's run `itd_tuning()` on our CHM data starting with default window size functions

```{r,eval=T,purl=FALSE}
###___________________________________________###
# itd_tuning
###___________________________________________###
# run it
set.seed(99)
itd_tuning_ans <- cloud2trees::itd_tuning(
  input_chm_rast = agg_chm_rast
  , min_height = 1.37
  , n_samples = 2
)
# huh?
itd_tuning_ans %>% names()
```

plot the detected trees on the CHM

```{r,eval=T,purl=FALSE, fig.height=6, fig.width=7}
itd_tuning_ans$plot_samples
ggplot2::ggsave(
  filename = file.path(c2t_process_dir, "itd_tuning_plot_samples.jpg")
  , plot = itd_tuning_ans$plot_samples
  , height = 7, width = 7
  , dpi = "print"
)
```

plot the summary of the detected trees

```{r,eval=T,purl=FALSE, fig.height=7, fig.width=8.3}
itd_tuning_ans$plot_sample_summary
ggplot2::ggsave(
  filename = file.path(c2t_process_dir, "itd_tuning_plot_sample_summary.jpg")
  , plot = itd_tuning_ans$plot_sample_summary
  , height = 7, width = 9.5
  , dpi = "print"
)  
```

the `exp_fn` results in too few trees with crown diameters generally similar to the total tree height which is not expected

check the `ws_fn_list` return which includes the different window size functions tested

```{r,eval=T,purl=FALSE}
# what ws_fn_list
itd_tuning_ans$ws_fn_list %>% str()
```

let's look at the function definition for the linear function (`lin_fn`)

```{r,eval=T,purl=FALSE}
# the linear function
itd_tuning_ans$ws_fn_list$lin_fn
```

let's plot all of the functions we tested with our call to `itd_tuning()` using the defaults

```{r,eval=T,purl=FALSE}
# shape of the ws functions
plt_ws_fn <- ggplot2::ggplot() +
  ggplot2::geom_function(mapping=ggplot2::aes(color = "lin_fn"),fun=itd_tuning_ans$ws_fn_list$lin_fn, lwd=1.2) +
  ggplot2::geom_function(mapping=ggplot2::aes(color = "log_fn"),fun=itd_tuning_ans$ws_fn_list$log_fn, lwd=1.2) +
  ggplot2::geom_function(mapping=ggplot2::aes(color = "exp_fn"),fun=itd_tuning_ans$ws_fn_list$exp_fn, lwd=1.2) +
  ggplot2::xlim(-5,35) +
  ggplot2::scale_color_manual( values = c(viridis::viridis(3), "gray") ) +
  ggplot2::labs(x = "heights", y = "ws", color = "ws function") +
  ggplot2::theme_light()
plt_ws_fn
```

Let's define our own custom linear function that slightly decreases the window size for the shortest trees but increases the window for taller trees compared to the default linear function.

```{r}
my_lin_fn <- function (x) {
      y <- dplyr::case_when(
        is.na(x) ~ 0.001
        , x < 0 ~ 0.001
        , x > (5.3-0.51)/0.21 ~ 5.3
        , TRUE ~ 0.51 + (x * 0.21)
      )
      return(y)
}
```

add to the plot

```{r,eval=T,purl=FALSE}
# plt
plt_ws_fn +
  ggplot2::geom_function(
    fun = my_lin_fn
    , mapping = ggplot2::aes(color = "my_lin_fn")
    , lwd = 2
  )
```

We'll run another sample test using `itd_tuning()`with our new function

```{r,eval=T,purl=FALSE}
itd_tuning_ans2 <- cloud2trees::itd_tuning(
  input_chm_rast = agg_chm_rast
  , min_height = 1.37
  , ws_fn_list = list(
    lin_fn = itd_tuning_ans$ws_fn_list$lin_fn
    , log_fn = itd_tuning_ans$ws_fn_list$log_fn
    , my_lin_fn = my_lin_fn
  )
)
```

plot the detected trees on the CHM

```{r,eval=T,purl=FALSE, fig.height=6, fig.width=7}
itd_tuning_ans2$plot_samples
ggplot2::ggsave(
  filename = file.path(c2t_process_dir, "itd_tuning_plot_samples2.jpg")
  , plot = itd_tuning_ans2$plot_samples
  , height = 7, width = 7
  , dpi = "print"
)
```

plot the summary of the detected trees

```{r,eval=T,purl=FALSE, fig.height=7, fig.width=8.3}
itd_tuning_ans2$plot_sample_summary
ggplot2::ggsave(
  filename = file.path(c2t_process_dir, "itd_tuning_plot_sample_summary2.jpg")
  , plot = itd_tuning_ans2$plot_sample_summary
  , height = 7, width = 9.5
  , dpi = "print"
)
```

notice, there is also a tree crown data frame in the return

```{r,eval=T,purl=FALSE}
# what?
itd_tuning_ans2$crowns %>% dplyr::glimpse()
# count trees by sample
itd_tuning_ans2$crowns %>% sf::st_drop_geometry() %>% dplyr::count(sample_number, ws_fn)
```

with the crowns data we can explore alternative visualizations including overlaying the detected trees on the RGB data if available

plot it

```{r,eval=T,purl=FALSE}
# make a function to plot these detected crowns with rgb data
plt_rgb_rast_itd_crowns <- function(sample_nmbr = 1, rgb_rast, itd_crowns, plt_lwd = 0.7) {
  # crop
  crp_rgb_rast_temp <- rgb_rast %>% 
    terra::crop(
      itd_crowns %>% 
        dplyr::ungroup() %>% 
        dplyr::filter(sample_number == sample_nmbr) %>% 
        sf::st_union() %>% 
        sf::st_bbox() %>% 
        sf::st_as_sfc() %>% 
        sf::st_buffer(0.2) %>% 
        sf::st_transform(terra::crs(rgb_rast)) %>% 
        terra::vect()
    )
  # convert raster to a data frame and create hex colors
  # ?grDevices::rgb
  rgb_df_temp <-
    crp_rgb_rast_temp %>% 
    terra::as.data.frame(xy = TRUE) %>%
    dplyr::rename(
      red = 3, green = 4, blue = 5
    ) %>% 
    dplyr::mutate(
      hex_col = grDevices::rgb(
        red
        , green
        , blue
        , maxColorValue = 255
      )
    )
    # dplyr::glimpse()
  
  # plt
  plt <- ggplot2::ggplot() +
    # add rgb base map
    ggplot2::geom_raster(data = rgb_df_temp, mapping = ggplot2::aes(x = x, y = y, fill = hex_col)) +
    # use identity scale so the hex codes are used directly
    ggplot2::scale_fill_identity() +
    # overlay polygons
    # ggplot2::geom_sf(data = polys, fill = NA, color = "red", linewidth = 0.5) +
    ggplot2::geom_sf(
      data = itd_crowns %>% 
        dplyr::filter(sample_number==sample_nmbr) %>% 
        cloud2trees::simplify_multipolygon_crowns() %>% 
        sf::st_make_valid() %>% 
        dplyr::filter(sf::st_is_valid(.)) %>% 
        sf::st_transform(terra::crs(crp_rgb_rast_temp))
      , mapping = ggplot2::aes(color = ws_fn)
      , fill = NA
      , lwd = plt_lwd
      , inherit.aes = F
    ) +
    ggplot2::facet_grid(cols = dplyr::vars(ws_fn)) +
    ggplot2::scale_color_viridis_d(name = "") +
    ggplot2::coord_sf(expand = F) +
    ggplot2::theme_void() +
    ggplot2::theme(
      legend.position = "none"
      , strip.text = ggplot2::element_text(face = "bold", color = "black", margin = ggplot2::margin(t = 4, b = 4))
      , strip.background = ggplot2::element_rect(fill = "gray88", color = "gray88")
      , panel.spacing = ggplot2::unit(1,"lines")
    )
  return(plt)
}
```

plot the trees detected in sample one on the RGB

```{r,eval=T,purl=FALSE, fig.height=5.5, fig.width=7.2}
plt_rgb_rast_itd_crowns(sample_nmbr = 1, rgb_rast = pj_rgb_rast, itd_crowns = itd_tuning_ans2$crowns)
```

```{r, include=FALSE, eval=FALSE}
### building the plt_rgb_rast_itd_crowns fn
  crp_rgb_rast_temp <- pj_rgb_rast %>% 
    terra::crop(
      itd_tuning_ans2$crowns %>% 
        dplyr::ungroup() %>% 
        dplyr::filter(sample_number == 1) %>% 
        sf::st_union() %>% 
        sf::st_bbox() %>% 
        sf::st_as_sfc() %>% 
        sf::st_buffer(0.5) %>% 
        sf::st_transform(terra::crs(pj_rgb_rast)) %>% 
        terra::vect()
    )
  crp_chm_rast_temp <- agg_chm_rast %>% 
    terra::crop(
      itd_tuning_ans2$crowns %>% 
        dplyr::ungroup() %>% 
        dplyr::filter(sample_number == 1) %>% 
        sf::st_union() %>% 
        sf::st_bbox() %>% 
        sf::st_as_sfc() %>% 
        sf::st_buffer(0.5) %>% 
        sf::st_transform(terra::crs(agg_chm_rast)) %>% 
        terra::vect()
    )
  # ?terra::plotRGB
  # ?terra::ifel
  terra::plotRGB(crp_rgb_rast_temp, stretch = "lin", axes = F)
  terra::plot(
    crp_chm_rast_temp
    # terra::ifel(crp_chm_rast_temp<1.37,NA,crp_chm_rast_temp)
    , add = T
    , col = viridis::plasma(100), alpha = 0.3
    , legend = F
    , axes = F
  )
  terra::plot(
    itd_tuning_ans2$crowns %>%
      cloud2trees::simplify_multipolygon_crowns() %>%
      dplyr::ungroup() %>%
      dplyr::filter(sample_number == 1 & ws_fn == "my_lin_fn") %>%
      sf::st_transform(terra::crs(rgb_rast_temp)) %>%
      terra::vect()
    , add = T, border = "brown", col = NA, lwd = 1.6
  )
  terra::plot(
    itd_tuning_ans2$crowns %>%
      cloud2trees::simplify_multipolygon_crowns() %>%
      dplyr::ungroup() %>%
      dplyr::filter(sample_number == 1 & ws_fn == "lin_fn") %>%
      sf::st_transform(terra::crs(rgb_rast_temp)) %>%
      terra::vect()
    , add = T, border = "gold", col = NA, lwd = 1.2
  )
  terra::plot(
    itd_tuning_ans2$crowns %>%
      cloud2trees::simplify_multipolygon_crowns() %>%
      dplyr::ungroup() %>%
      dplyr::filter(sample_number == 1 & ws_fn == "log_fn") %>%
      sf::st_transform(terra::crs(rgb_rast_temp)) %>%
      terra::vect()
    , add = T, border = "navy", col = NA, lwd = 0.8
  )
  
  xxx <-
    itd_tuning_ans2$crowns %>%
      dplyr::ungroup() %>%
      cloud2trees::simplify_multipolygon_crowns() %>%
      sf::st_make_valid() %>%
      dplyr::filter(sample_number == 1 & ws_fn == "my_lin_fn")

  mapview::mapview(xxx, zcol = "tree_height_m")
  remove(xxx)
  gc()
  
  ctg_temp <- lidR::readLAScatalog(
    file.path(c2t_output_dir, "point_cloud_processing_temp", "02_normalize")
  )
  las_temp <- lidR::clip_roi(
    ctg_temp
    , itd_tuning_ans2$crowns %>%
      dplyr::ungroup() %>%
      cloud2trees::simplify_multipolygon_crowns() %>%
      sf::st_make_valid() %>%
      dplyr::filter(sample_number == 1 & ws_fn == "my_lin_fn") %>%
      dplyr::filter(
        # treeID %in% c("68_708803.7_4193239.3", "67_708804.9_4193239.9" , "69_708804.9_4193237.9", "66_708803.9_4193240.3")
        # treeID %in% c("6_708811.3_4193258.3")
        treeID %in% c("71_708809.1_4193235.3", "76_708807.3_4193233.9", "72_708805.5_4193234.7")
      ) %>%
      sf::st_union() %>%
      sf::st_buffer(1) %>%
      sf::st_transform(lidR::st_crs(ctg_temp))
  )
  las_temp
  lidR::plot(
    las_temp, color = "RGB", bg = "white", legend = F
  )
  lidR::plot(
    las_temp, color = "Z", bg = "white", legend = T
  )
  
  # convert raster to a data frame and create hex colors
  # ?grDevices::rgb
  rgb_df_temp <-
    crp_rgb_rast_temp %>% 
    terra::as.data.frame(xy = TRUE) %>%
    dplyr::rename(
      red = 3, green = 4, blue = 5
    ) %>% 
    dplyr::mutate(
      hex_col = grDevices::rgb(
        red
        , green
        , blue
        , maxColorValue = 255
      )
    )
    # dplyr::glimpse()
  
  # plt
  ggplot2::ggplot() +
    # add rgb base map
    ggplot2::geom_raster(data = rgb_df_temp, mapping = ggplot2::aes(x = x, y = y, fill = hex_col)) +
    # use identity scale so the hex codes are used directly
    ggplot2::scale_fill_identity() +
    # overlay polygons
    # ggplot2::geom_sf(data = polys, fill = NA, color = "red", linewidth = 0.5) +
    ggplot2::geom_sf(
      data = itd_tuning_ans2$crowns %>% 
        dplyr::filter(sample_number==1) %>% 
        sf::st_transform(terra::crs(crp_rgb_rast_temp))
      , mapping = ggplot2::aes(color = ws_fn)
      , fill = NA
      , lwd = 0.8
      , inherit.aes = F
    ) +
    ggplot2::facet_grid(cols = dplyr::vars(ws_fn)) +
    ggplot2::scale_color_viridis_d(name = "") +
    ggplot2::coord_sf(expand = F) +
    ggplot2::theme_void() +
    ggplot2::theme(
      legend.position = "none"
      , strip.text = ggplot2::element_text(face = "bold", color = "black", margin = ggplot2::margin(t = 4, b = 4))
      , strip.background = ggplot2::element_rect(fill = "gray88", color = "gray88")
      , panel.spacing = ggplot2::unit(1,"lines")
    )
  
  
  remove(list = ls()[grep("_temp",ls())])
  gc()
```

this is only moderately helpful as the tree density in tree clumps is difficult to infer from the 2D aerial imagery alone. let's move forward with the default linear function in the `raster2trees()` function

```{r}
my_ws_fn <- cloud2trees::itd_ws_functions()[["lin_fn"]]
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
remove(itd_tuning_ans, itd_tuning_ans2)
gc()
```